### 2020-05-21 write-up/scifile.tex
Update scifile.tex

% Use only LaTeX2e, calling the article.cls class and 12-point type.

\documentclass[12pt]{article}

% Users of the {thebibliography} environment or BibTeX should use the
% scicite.sty package, downloadable from *Science* at
% http://www.sciencemag.org/authors/preparing-manuscripts-using-latex
% This package should properly format in-text
% reference calls and reference-list numbers.

\usepackage{scicite}

\usepackage{times}
\usepackage{amsmath,amsthm,amsfonts}

% The preamble here sets up a lot of new/revised commands and
% environments.  It's annoying, but please do *not* try to strip these
% out into a separate .sty file (which could lead to the loss of some
% information when we convert the file to other formats).  Instead, keep
% them in the preamble of your main LaTeX source file.

% The following parameters seem to provide a reasonable page setup.

\topmargin 0.0cm
\oddsidemargin 0.2cm
\textwidth 16cm
\textheight 21cm
\footskip 1.0cm

%The next command sets up an environment for the abstract to your paper.

\newenvironment{sciabstract}{%
\begin{quote} \bf}
{\end{quote}}

\def\I{{\bf I}}
\def\bary{{\bar y}}
\def\pr{\text{pr}}
\def\sgn{\text{sgn}}

% Include your paper's title here

\title{The Hypothesis of Testing: Paradoxes arising out of reported coronavirus case-counts}

% Place the author information here.  Please hand-code the contact
% information and notecalls; do *not* use \footnote commands.  Let the
% author contact information appear immediately below the author names
% as shown.  We would also prefer that you don't change the type-size
% settings shown here.

\author
{Walter Dempsey$^{1\ast}$\\
\\
\normalsize{$^{1}$Department of Biostatistics, University of Michigan,}\\
\normalsize{1415 Washington Heights, Ann Arbor, MI 48109, USA}\\
\\
\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  wdem@umich.edu.}
}

% Include the date command, but leave its argument blank.

\date{}

%%%%%%%%%%%%%%%%% END OF PREAMBLE %%%%%%%%%%%%%%%%

\begin{document}

% Double-space the manuscript.

\baselineskip24pt

% Make the title.

\maketitle

% Place your abstract within the special {sciabstract} environment.

\begin{sciabstract}
  Many statisticians, epidemiologists, economists and data scientists have registered serious reservations regarding the reported coronavirus case-counts. Limited testing capacity across the country has been widely identified as a key driver of suppressed coronavirus case-counts.  The calls to increase testing capacity are well-justified as they become a more frequent point of discussion in the public sphere.  While expanded testing is a laudable goal, selection bias will impact estimates of disease prevalence and the effective reproduction number until the entire population is sampled.  Moreover, tests are imperfect as false positive/negative rates interact in complex ways with selection bias.  In this paper, we attempt to clarify this interaction.  Through simple calculations, we demonstrate pitfalls and paradoxes that can arise when considering case-count data in the presence of selection bias and measurement error. The discussion guides several suggestions on how to improve current case-count reporting.
\end{sciabstract}

\paragraph*{Introduction}

The World Health Organization has declared the coronavirus disease 2019 (COVID-19) a public health emergency.  As of May 21st, 2020, over 5 million cases have been confirmed worldwide with 1.5 million cases and over 90 thousand confirmed deaths across the United States. This pandemic
has become the focal point of every day life; yet the data landscape for understanding COVID-19 remains limited.  Public databases~\cite{JHU_Lancet,NYT} provide incoming county-level information of confirmed cases and deaths.  Statisticians, epidemiologists, economists, and data scientists have used this granular data to forecast COVID-19 case-counts, deaths, and hospitalizations~\cite{Song2020,Ray2020,2020.IHME}.

The goal of this paper is to express reservations at the use of case-counts as a proxy for disease prevalence and in estimation of standard epidemiological models for inference and forecasting.  The reason is straightforward: observed case-count data is plagued by selection bias and measurement error. We will demonstrate that the information gained from increasing testing capacity is limited when selection bias and testing inaccuracies persist.

\paragraph*{Analysis of imperfect case-count data}

Let $N$ denote the population size.  At a fixed time, let $Y_j$ denote COVID-19 status for the $j$th individual in the population, $j=1,\ldots, N$. To start, for simplicity, we ignore the dynamic nature of the outbreak and the fact that individuals can recover.  As in survey methodology~\cite{Cochran77}, we treat COVID-19 status as a fixed but unknown quantity of interest. Let $I_j \in \{0,1\}$ be an indicator that the individual was selected for testing ($I = 1$) or not ($I=0$). Suppose we are interested in COVID-19 prevalence, i.e., the population average $\bar Y = Y/N$, and that $n$ tests are performed and we observe the values $y_1, \ldots, y_n \in \{0,1\}$.  The natural estimator is the proportion of positive tests $\bar y = n^{-1} \sum_{i=1}^n y_i$.

Tests are imperfect.  COVID-19 testing is no exception which implies a selected individual's true status is not observed. Here we assume the false positive rate (FP) and false negative rate (FN) are known.  Here we consider the estimator $\tilde y_n = \bar y_n + (1-\bar y_n) FP + FN \bar y_n$ which is unbiased under simple random sampling (SRS). To understand the impact of selection bias and imperfect testing, we provide a statistical decomposition of the error between $\tilde y_n$ and $\bar Y$
\begin{equation}
\label{eq:error}
\rho_{I,Y} \times \sqrt{\frac{1-f}{f}} \times \sigma_{Y}
\times \underbrace{\left[ 1 + FP + FN - \Delta \times \frac{\bar Y}{1-\bar Y} \times \frac{FP(1-\bar Y) + FN \cdot \bar Y}{f_0 (1-\bar Y) + f_1 \bar Y} \right]}_{D_M},
\end{equation}
where $f_1 := \pr (I_J = 1 \mid Y_J = 1)$ and $f_0 := \pr(I_J = 1 \mid Y_J = 0)$ are sampling rates for COVID-19 positive and negative individuals respectively, and $\Delta = f_1 - f_0$ is the sampling rate differential.  This extends work by \cite{Meng2018} to account for imperfect testing. The first term represents \emph{data quality}, the second \emph{data quantity}, and the third \emph{problem difficulty}.  The final term represents the \emph{imperfect testing adjustment} which is a complex function of the sampling rate differential, the odds ratio, and the ratio of measurement error interaction with prevalence and sampling rates interaction with prevalence. Figure~2 presents $D_M$ as a function of the relative frequency and odds ratio can be both positive and negative as well as a range of magnitudes~\cite{Beesley2020,Beesley2019,Smeden2019}.

Comparing the mean-squared error (MSE) under a selection mechanism $\I$ with imperfect testing and SRS with perfect testing, we see that
$$
\frac{E_{\I} \left[ (\bar y_n - \bar Y)^2 \right]}{\sqrt{V_{SRS} (\bar Y)}} = (N-1) E_{\I} \left[ \rho_{I,Y}^2 D_M^2 \right].
$$
This points to a paradoxical situation: the error relative to SRS increases as a function of population size.  This was termed the ``Law of Large Populations'' (LLP)~\cite{Meng2018} and continues to hold in our imperfect testing context.  This points to a critical issue in current media practices in communication of case-count data: two countries with the same testing strategy can yield very different estimates due to population size.  When comparing countries like the US to the UK, the MSE increasea by a factor of almost five.  Thus conclusions drawn from observed case-count records may be not just wrong, but very wrong.

An important follow-up question is ``What is the sample size from a SRS with perfect testing that would yield equivalent MSE to the current testing strategy?'' We show this \emph{effective} sample size can be bounded by
$$
n_{eff} \le \frac{f}{1-f} \times \frac{1}{E_{I} \left[ \rho_{I,Y}^2 D_M^2 \right]}.
$$
As of May 8th, the US has performed $8,412,095$ tests.  The US population is roughly $328$ million, so the fraction sampled is $f = 0.026$.  The empirical prevalence is $9.1\%$ on May 8th. A recent study estimates a false positive and negative rate for antibody testing using a lateral flow immunoassay are $0.5$\% and $17.2$\% respectively \cite{Bendavid2020}. Supposing COVID-19 positive individuals are $2$ times more likely to get tested, then using these parameters the effective sample size is $10$.  Recent proposals~\cite{Siddarth2020} have argued for increased testing capacity, which will reduce the relative sampling rate.  Supposing it drops to $1.2$ then the effective sample size will increase to $216$.  This points to a sad state of affairs, saying that the effective sample size even in optimistic scenarios is not better than a small random sample from the population.

Increased testing capacity may also come at the cost of increased in false positive and negative rates.  In terms of effective sample size, if the data quality term remains constant, then testing increases may not significantly improve the effective sample size.  Consider the simplest case where $f$ increases from $0.05$ to $0.1$; suppose this increase was associated with  false positive rate increasing from $0.5$\% to $5$\% and the false negative rate rising from $5\%$ to $20$\%.  Then the standard analysis would suggest a reduction in the MSE by 47\% but the actual reduction is 26\%.  This implies that doubling the sampling rate will only increase the effective sample size by a factor $2.9$ rather than the expected $4$ fold increase that is expected.

\paragraph*{Regrettable rates.}

Another important question is whether selection bias and imperfect testing impact the ability to estimate rates of growth in case-count.  Let $\bar Y_{t-1}$ and $\bar Y_{t}$ denote the prevalence on two consecutive times $t-1$ and $t$.  We consider the ratio estimator $r = \bar y_t / \bar y_{t-1}$.  Using a Taylor series approximation, the error between ${\bar y_t}/{\bar y_{t-1}}$ and the true rate ${\bar Y_{t}}/{\bar Y_{t-1}}$ can be expressed approximately as
$$
\begin{aligned}
\frac{\bar Y_t}{\bar Y_{t-1}} &\times \bigg[ \rho_{I_t,Y_t} D_{M_t} \sqrt{\frac{1-f_t}{f_t}} CV (Y_t)  -\rho_{I_{t-1},Y_{t-1}} D_{M_{t-1}} \sqrt{\frac{1-f_{t-1}}{f_{t-1}}} CV (Y_{t-1}) \bigg] \\
&\times \left[ 1 - \rho_{I_{t-1},Y_{t-1}} D_{M_{t-1}} \sqrt{\frac{1-f_{t-1}}{f_{t-1}}} CV (Y_{t-1}) \right]
\end{aligned}
$$
where $\rho_{I_j, Y_j}$ is the data quality, $f_j$ is the sampling fraction, $D_{M_j}$ is the measurement error adjustment, and $CV(Y_j) = \sigma_{Y_j}/\bar Y_j$ is the coefficient of variation at time step $j$.  The error magnitude depends on the true rate $\bar Y_{t} / \bar Y_{t-1}$ so a large decrease will have a relatively small error compared to a large increase. The second term represents potential \emph{cancellation} which may occurs when data quality, sampling fraction, measurement error, and prevalence are all constant across time.

Consider the state of New York.  Figure~5 shows the number of tests and positive tests per day in New York state.  Figure~5b presents the fraction of tests that were positive per day in New York state.  For cancellation, the expected data qualities need to be uncorrelated and equal. Testing per day suggests it is plausible to assume $\Delta_t$ varies slowly over time.  Note, however, that constant $\Delta_t$ does not imply constant data quality.  Figure~5a shows the number of tests has remained approximately constant in the last few weeks relative to the total NY population.  Therefore, the assumption that $f_t$ also varies slowly in time is reasonable. Finally, the assumption that the imperfect testing adjustment is constant is reasonable as no new testing procedures have been reported.  This would not be true in states like Michigan, where testing criteria have been expanding every few weeks.

Figure~3a shows a trajectory of the true ratio and the potential biased estimators under an SIR model~\cite{Pastor2001,Newman2002,Parshani2010} for the epidemic dynamics, with state evolution given by the following ODE
$$
\begin{aligned}
\frac{\partial S_t}{\partial t} &= - \beta S_t I_t N^{-1}; \quad
\frac{\partial I_t}{\partial t} = \beta S_t I_t N^{-1} - \gamma I_t; \quad
\frac{\partial R_t}{\partial t} = - \gamma I_t \\
\end{aligned}
$$
where $S_t$ is the number of susceptible individuals at time $t$, $I_t$ is the number of infected individuals at time $t$, $N$ is the total population size, and $R_t$ is the number of removed (recovered or deceased) individuals at time $t$.   Figure~4a presents the fraction of population newly infected individuals at each time step for $\beta = 1.4$ and $\gamma = 0.2$ (black line).  In terms of bias, the rate is likely to be overestimated prior to the peak and then underestimated afterwards.  From a decision making perspective, such biases have potential impact.  First, the over-estimation may give lawmakers and governors more leverage in proposing aggressive actions that reduce prevalence.  Of course, the analysis supports the argument that the estimates based on available data are overly pessimistic. What both sides miss is that the direction of bias is non-constant over time.  After the peak, the estimator likely under-estimates the rate of change.  Such bias puts pressure on lawmakers and governors to prematurely relax social distancing measures.  Moreover, it appears that the peak time is the easiest to estimate in terms of having minimal error using availability data.  Note, however, that the standard errors on such estimates will be wide due to relatively low effective sample size.

\paragraph*{Estimation of effective reproduction number.}
Many epidemiologists argue that tracking the effective reproduction number $R_t$ is the only way to manage through the crisis~\cite{Gabriel2020}.  Under a Poisson likelihood, a simple relation between the trajectory of new cases and the effective reproduction number can be derived \cite{Bettencourt2008}.  In particular, under an SIR model the number of case counts on day $t$ is Poisson distributed with rate $K_{t-1} \exp \left( \frac{1}{\gamma} (R_t - 1) \right)$ where $K_{t-1} = Y_{t-1}-Y_{t-2}$ is the number of new cases on day $t-1$ and $\gamma$ is the serial interval which is approximately $7$ days for COVID-19~\cite{Sanche2020}.  Using this, a simple estimator can be given by
$$
R_t = 1 + \frac{1}{\gamma} \log \left( \frac{K_t}{K_{t-1}} \right).
$$
Of course, we do not observe $K_t$ and $K_{t-1}$.  Under SRS of new cases among those susceptible on day $t$, the natural estimator is $S_t \bar y_t$.  Unfortunately the number of susceptible individuals on day $t$ is unknown.  Here, we study the estimator $\hat R_t = 1 + \frac{1}{\gamma} \log \left( \bar y_t / \bar y_{t-1} \right)$. We can again express the statistical error of $\hat R_t - R_t$ in useful terms as follows
$$
\begin{aligned}
\frac{1}{\gamma}\log &\bigg( 1 + \bigg[ \rho_{I_t,K_t} D_{M_t} \sqrt{\frac{1-f_t}{f_t}} CV (K_t)  -\rho_{I_{t-1},K_{t-1}} D_{M_{t-1}} \sqrt{\frac{1-f_{t-1}}{f_{t-1}}} CV (K_{t-1}) \bigg] \\
&\times \left[ 1 - \rho_{I_{t-1},K_{t-1}} D_{M_{t-1}} \sqrt{\frac{1-f_{t-1}}{f_{t-1}}} CV (K_{t-1}) \right] \bigg) - \frac{1}{\gamma} \log \left( \frac{S_t}{S_{t-1}} \right).
\end{aligned}
$$
This demonstrates the exact same trade-off as with the rates but on the logarithmic scale.  For small values of $x$, we have $\log(1+x) = x$ and so the biases are similar.  However, the error is no longer scaled by the rate of change in prevalence $\bar Y_2/\bar Y_1$. Instead the error is scaled by the serial interval $\gamma^{-1} = 1/7$.  Moreover, the error depends not on aggregate prevalence $\bar Y_t$ but fraction of new cases $\bar K_t$. The final term is the error due to varying population sizes.  Since $S_t \leq S_{t-1}$, the final term $- \log (S_t/S_{t-1}) \geq 0$. Figure~4b displays the bias for the effective reproductive number and how it is impacted by the relative sampling fraction.  Here for simplicity, we assume the fraction of the population that is susceptible remains large and near the total population size. Changes to this assumption can impact the potential bias.  The key difference between Figure~4a and Figure~4b is that prevalence depends on the fraction of the population infected at time $t$ while $\hat R_t$ depends on the fraction of new cases in the population. This leads to differences in when the bias is most pronounced.

\paragraph*{Cross-population rate comparisons.}
So far we have focused on understanding the limitations of using case-count data to understand population quantities of interest for a \emph{single} population.  Many are interested in cross-population comparisons.  Many pieces in the media as well as academic articles have plots of case-count over time aligned by time of first known case.  Others have claimed such comparisons unfair due to unequal population sizes and instead plotted case-count per million.  Here, for simplicity, we focus on comparing the estimated effective reproductive rate.  We assume the two time-series are aligned so that $t=0$ is the time of first case.

We focus on the interaction of biases in estimation of the two trajectories when the peak infection times differ slightly.  Recall that the error allowed for over-estimation prior to the peak and then under-estimation post-peak.  Here, these errors can mingle in interesting ways.  Consider two countries (A and B) in which the peak occurs 2 weeks prior for country A then country B but other.  Figure~4b presents such a comparison where each country's disease trajectory follows an SIR model as described above with different parameters.  Figure~4 shows how the errors interact in complex ways.  First, the gap is correctly estimated; then the gap is over-estimated as country A sees a rapid rise in cases; then the gap is even more over-estimated as country A improves and the rate is under-estimated while country B sees a rapid increase in their case-count and their rate is over-estimated; then country A's rate is correctly estimated while country B's rate is under-estimated as it sees improvement in infection rates; finally, the gap disappears.  We are not claiming this will always be the case, and indeed for small enough selection bias and/or large gaps in case-counts such complex behaviour will not occur; however, this is simply to point out how observed information can tell a more complex story than the truth (i.e., country A recovers 1 month prior to country B).

\paragraph*{Estimation of Data Quality.} Here, we try and address the empirical question of how to estimate data quality $E_{\I} \left[ \rho_{I,Y} \right]$ in the current context. In \cite{Meng2018}, estimation relied on observing the true outcome.  In election data, for example, this is possible as election vote totals can be used as ground truth.  This is not possible in the current crisis.  Here, we propose a simple procedure that uses survey samples as noisy but unbiased estimates of the true prevalence. Our goal is less inferential and more to provide a sensitivity analysis that can aid in understanding the amount of information in observational COVID case-count data.

On April 23rd, Andrew Cuomo announced results from a study in New York state.  It found 13.9\% of 3,000 people tested across the state had signs of the virus.  The study did not report sensitivity and specificity; therefore, we take reported measures from the Santa Clara study~\cite{Bendavid2020}.  The reported specificity is $82.8\%$ (95\% CI: 76.0-88.4\%) and sensitivity is $99.5\%$ (95\% CI: 99.2-99.7\%).  This corresponds to a false negative rate of $17.2\%$ (11.6-24.0\%) and false positive rate of $0.5\%$ (0.3 - 0.8\%).  Correcting for these rates we have an estimated prevalence of $15.9\%$ with a range of 14.8\% to 17.0\% using the ranges of the false positive and negative rates.

Given the report was on the 23rd, we suppose for simplicity that the study was run several days prior on the 20th of April. On the 20th, there were $19,654$ tests performed and $617,555$ tests had been performed up to to and including the 19th. The population of New York state is $19,378,102$.   Subtracting off the number of individuals who had already been tested yields a sampling fraction of $f = 0.001$ on the 20th.  The smoothed prevalence estimate on the 20th is 28.3\%.  Adjusting for measurement error yields an estimate of $32.5\%$.

Here, we assume the New York survey is a simple random sample from the state's population. We recognize that the survey is not such a sample, but techniques like poststratification and raking can be applied to calibrate estimates. Then the error is $16.6$\%.  Using~\eqref{eq:error}, we can use the error to construct an estimate of the relative sampling rate:
\[
\rho D_M = \sqrt{\frac{f}{1-f}} \frac{\text{0.166}}{\sigma_Y} = 1.43 \times 10^{-2} \Rightarrow \Delta = 1.06 \times 10^{-3} \Rightarrow M = 2.29.
\]
The range for the CIs is (2.05, 2.54).  We can also perform sensitivity as a function of the false negative and positive rates.  Under these measurement error ranges, we saw that the impact on the effective reproductive rate estimator was much smaller.  However, for $FP = 0.05$ and $FN = 0.005$ we have $M = 4.31$.  While nothing above is definitive, it points to the interplay of selection bias and measurement error.  This calculation gives us simple tools to understand how sensitivity and specificity impact our estimators and therefore how much we should trust conclusions based on these assumptions.  Under the Santa Clara study estimates of these quantities, while we expect impact, the bias is much smaller than under other estimates.

\paragraph*{Discussion}
\label{section:discussion}

There is nothing routine about COVID-19, including the corresponding statistical questions.  The goal of this paper was to point out questionable statistical routines.  Indeed, case-count trajectories are being reported daily by governments and broadcast by the media.  The precision in the reported numbers gives the illusion of accuracy when what needs to be understood is the level of uncertainty. Selection bias not only means uncertainty, it also implies analysts feel certain about incorrect conclusions.  We end this paper with a brief discussion of several important topics related to selection bias and measurement error.

\paragraph*{Stratified sampling: improving precision in low-prevalence environments}

This article argues why random sampling is a powerful tool and the limitations of data obtained through a self-selection process.  As COVID-19 survey studies become more readily accessible, we encourage researchers to focus on building representative samples.  One difficulty in the current crisis is the low prevalence relative to the population size in certain geographies.  Take the recent Santa Clara study, for example, that estimated less than 2\% prevalence in the area.  The key issue with low prevalence is that if the test's specificity is not lower than the prevalence, then the observed data is consistent with zero recovered individuals.

In such settings, simple random sampling is likely to fail; however, there are potential solutions.  For example, one can stratify the population into risk categories (based on population density, occupation, age, and other important factors) and perform Neyman allocation~\cite{Cochran77}.  One could go attempt to go further and incorporate contact network information.  However, such data is difficult to obtain in practice.  Indeed, geographic location, occupation, age and basic demographic information are decent good proxies for latent network connectivity.

\paragraph*{Effect modifiers in COVID-19 clinical trials}

Here we show how selection bias and measurement error can creep into clinical trial analysis. The key concern is whether clinical trials on COVID-19 recruit from the pool of individuals who have tested positive for COVID-19 or whether they sample randomly from the population, test, and then recruit from this subset of tested individuals.  Randomization of treatment assignment in clinical trials negates unobserved confounders.  This, however, does not negate effect modifiers.  Therefore, for marginal treatment effects to be interpretable, there must be a well-defined population.  Most often, our main interest is in causal effects on the population of COVID-19 positive.  Randomization within the clinical trial design yields internal validity, but we also need external validity to estimate the correct marginal effect of interest~\cite{Keiding2016}.

\paragraph*{Model-based solutions.}

One alternative approach to what is presented in this article is a model-based extension of SIR models to try and capture both selection bias and measurement error; however, without strong assumptions on the selection mechanism, the estimates are often not nonparametrically identifiable.  When an issue ``cannot be resolved nonparametrically then it is usually dangerous to resolve it parametrically''\cite{CoxHink74}. In the absence of randomized experiments, the best route forward using any approach is the use of detailed sensitivity analyses and humility in conclusions drawn from such data sources.

% \bibliography{scibib}
\bibliographystyle{Science}
\bibliography{covid-refs}

\section*{Acknowledgments}
Include acknowledgments of funding, any patents pending, where raw data for the paper are deposited, etc.

%Here you should list the contents of your Supplementary Materials -- below is an example.
%You should include a list of Supplementary figures, Tables, and any references that appear only in the SM.
%Note that the reference numbering continues from the main text to the SM.
% In the example below, Refs. 4-10 were cited only in the SM.
\section*{Supplementary materials}

\subsection*{Prevalence comparisons}

At a fixed time, suppose prevalence estimates $\bar y_1$ and $\bar y_2$ are observed for two populations. We can express $\bar y_1 - \bar y_2$ in statistical terms as
$$
(\bar Y_1 - \bar Y_2) + \left[ \rho_{I_1, Y_1} \times D_{M_1} \times \sqrt{\frac{1-f_1}{f_1}} \times \sigma_{Y_1}  - \rho_{I_2, Y_2} \times D_{M_2} \times \sqrt{\frac{1-f_2}{f_2}} \times \sigma_{Y_2} \right].
$$
The first term is the difference in population prevalence (i.e., the quantity of interest).  If random sampling were performed in each population, $E_{\I} [ \rho_{I_j, Y_j}] = 0$ and so the estimate would be unbiased.  However, the second term represents the complex error that results from selection bias.
The particular test of interest is $H_0: \bar Y_1 = \bar Y_2$.  Under the null, $\sigma_{Y_1} = \sigma_{Y_2}$.  In classical statistics, we would compute a Z-score:
$$
\begin{aligned}
\frac{\bar y_1 - \bar y_2}{ \sqrt{V_{SRS} (\bar y_1 - \bar y_2)} } &=
\frac{\rho_{I_1, Y_1} \times D_{M_1} \times \sqrt{\frac{1-f_1}{f_1}}  - \rho_{I_2, Y_2} \times D_{M_2} \times \sqrt{\frac{1-f_2}{f_2}} }{ \sqrt{\frac{1}{N_1 -1} \times \frac{1-f_1}{f_1} + \frac{1}{N_2 - 1} \times \frac{1-f_2}{f_2}} }  \\
&\overset{\star}{=}
\sqrt{\frac{(N_1-1)(N_2 -1)}{N_1 + N_2}} \left[ \rho_{I_1, Y_1} - \rho_{I_2, Y_2} \right]
\end{aligned}
$$
Where the last equality holds under $f_1 = f_2$, i.e., same sampling rate relative to population size. This tells us that the exact error in the comparison of sample means, as an estimate of the difference in population means, scales as a function of the difference in data quality $\rho_{I_1, Y_1} - \rho_{I_2, Y_2}$ times a square root of a population adjustment.  For highly unequal population sizes (e.g., $N_1 \gg N_2$), the adjustment is approximated well by the square root of the smaller population size minus one (e.g. $\sqrt{N_2 -1}$).

Take, for example, a comparison of the US and Canada.  The two population sizes are approximately 328 million and 38 million people.  Then the population adjustment is $5.84$ million. As of May 5th, Canada has performed a total of 24.92 tests per 1,000 people while the US has performed a total of 22.01 tests per 1,000 people~\cite{worlddata} so the data quantities are approximately equal to $f = 23/1000 = 0.023$.  Then, under a prevalence of 10\%, one would need the differential $\Delta_1-\Delta_2$ to be much smaller than $11.68 \times 10^{-6}$ in order for the selection bias to not impact.  So only under the very strong assumptions of equal data quantity and quality will the comparisons be valid and the $Z$-score can be treated as in the classical setting.  A simple alternative is to use the effective sample sizes to build more appropriate $Z$-score.  That is, setting
$$
Z_n = \frac{\bar y_1 - \bar y_2}{ \frac{1-f}{f} \sigma_y \sqrt{ \frac{1}{n_{1,eff} - 1} + \frac{1}{n_{2,eff} - 1}}}
$$
allows us to build Z-scores that account for the true sample size.  While still not perfect, this allows us to build Z-scores that are not overly confident due to selection bias and measurement error.

\subsection*{Technical details}

Let $P_j$ be an indicator of measurement error, equal to $1$ when we incorrectly measure the outcome and $0$ when we observe the true outcome. We suppose this is a stochastic variable where $\pr(P_j = 1 \mid Y_j = 1) =: FN$ is the false-negative rate and $\pr(P_j = 1 \mid Y_j = 0) =: FP$ is the false-positive rate.  If individual $j$ is selected (i.e., $I_j = 1$) then the observed outcome can be written as $Y_j^{\star} = Y_j(1-P_j) + (1-Y_j) P_j$.

\subsection*{Imperfect testing: derivation}
\label{app:imperfect}
We considered the mean estimator
$$
\bar y_n^\star = \frac{\sum_{j=1}^N Y_j^\star I_j}{\sum_{j=1}^N I_j} = \frac{\sum_{i=1}^N  I_j Y_j^\star }{\sum_{j=1}^N  I_j } = \frac{\sum_{i=1}^N  I_j \left[ Y_j (1-P_j) + (1-Y_j) P_j \right]}{\sum_{j=1}^N  I_j }
$$
For any set of numbers $\{ A_1, \ldots, A_N \}$ we can view it as the support of a random variable $A_J$ induced by the random index $J$ defined on $\{1,\ldots, N\}$.  When $J$ is uniformly distributed $E_J (A_J) = \sum_{j=1}^N A_j / N \equiv \bar A_N$. Then
$$
\begin{aligned}
\bar y_n^\star  - \bar Y_N &= \frac{E_J \left[ I_J \left[ Y_J (1-P_J) + (1-Y_J) P_J \right] \right]}{E_J [ I_J ] } - E_J[Y_J] \\
&= \frac{E_J \left[ I_J P_J (1-2Y_J) \right]}{E_J [ I_J ] } + \left( \frac{E_J [I_J Y_J]}{E_J [ I_J ] } - \frac{E_J[Y_J] E_J[I_J]}{E_J[I_J]} \right) \\
\end{aligned}
$$
The term in parentheses can be re-written as
$$
\begin{aligned}
\frac{E_J [I_J Y_J]- E_J[Y_J] E_J[I_J]}{E_J[I_J]} &=  \frac{E_J [I_J Y_J]- E_J[Y_J] E_J[I_J]}{\sqrt{V_J(I_J) V_J(Y_J)}} \frac{\sqrt{V_J(I_J)}}{E_J[I_J]} \times \sqrt{V_J(Y_J)} \\
&= \rho_{I,Y} \times \sqrt{\frac{(1-f)}{f}} \times \sigma_Y
\end{aligned}
$$
which agrees with Meng's (2019) decomposition. For the other term, first we define $Z_j := 1 - 2 Y_j $. Then $Z_j = 1$ if $Y_j = 0$ and $Z_j = -1$ if $Y_j = 1$. Then the term can be re-written as
$$
\begin{aligned}
\frac{E_J \left[ I_J P_J (1-2Y_J) \right]}{E_J [ I_J ] } &= \left( \frac{E_J \left[ I_J P_J Z_J \right]}{E_J [ I_J ] } -  \frac{E_J \left[ P_J Z_J \right] E_J[ I_J]}{E_J [ I_J ] } \right) +  \frac{E_J \left[ P_J Z_J \right] E_J[ I_J]}{E_J [ I_J ] } \\
\end{aligned}
$$
The term in parentheses can be re-expressed using the previous technique as:
$$
\rho_{I, PZ} \times \sqrt{\frac{1-f}{f}} \times \sigma_{PZ}
$$
where now the ``data defect'' and ``problem difficulty'' are with respect to $PZ$ rather than $Y$. The final term is equal to
$$
\begin{aligned}
E_J [P_J Z_J ] &= E_J [ E_J [ P_J Z_J \mid Y_J ] ] \\
&= \pr (P = 1 \mid Y = 0) (1-\bar Y) - \pr(P=1 \mid Y = 1) \bar Y \\
&= FP - (FP + FN) \cdot \bar Y
\end{aligned}
$$
Combining these yields:
$$
\bar y_n^\star - \bar Y = \sqrt{\frac{1-f}{f}} \left[\rho_{I,Y} \sigma_Y + \rho_{I, PZ} \sigma_{PZ} + \sqrt{\frac{f}{1-f}} \left( FP - (FP+FN) \bar Y \right) \right]
$$
For the binary outcome $Y$, we have $\sigma_Y = \sqrt{\bar Y (1-\bar Y)}$. Moreover,
$$
\begin{aligned}
V_J(P_J Z_J) &= E_J[(P_J Z_J)^2] - E[P] E[Z] \\
&= E[P] - E[P] (1 - 2 \bar Y) = 2 \bar Y E_J [ P_J ] \\
&= 2 \bar Y \left( FP (1-\bar Y) + FN \bar Y \right) \\
\Rightarrow \sigma_{PZ} &= \sqrt{ 2 \bar Y \left( FP (1-\bar Y) + FN \cdot  \bar Y \right) }
\end{aligned}
$$
Then the formula for the error is given by:
$$
\sqrt{\frac{1-f}{f}} \left[\rho_{I,Y} \sqrt{\bar Y (1-\bar Y)} + \rho_{I, PZ} \sqrt{ 2 \bar Y \left( FP (1-\bar Y) + FN \cdot \bar Y \right )} + \sqrt{\frac{f}{1-f}} \left( FP - (FP+FN) \bar Y \right) \right]
$$

\subsubsection*{Estimate of observed data quality}

By definition, we have
$$
\begin{aligned}
\rho_{I,PZ} &= \frac{C(I, PZ)}{\sqrt{V(PZ) V(I)}} \\
&= \frac{C(I, PZ)}{\sqrt{V(Y) V(I)}} \sqrt{\frac{V(Y)}{V(PZ)}} \\
&= \rho_{I,Y} \frac{C(I,PZ)}{C(I,Y)} \sqrt{ \frac{(1-\bar Y)}{2 ( FP (1-\bar Y) + FN \cdot \bar Y)} }
\end{aligned}
$$

$$
\begin{aligned}
C(I, PZ) &= E[ I P Z ] - E[I] E[PZ] \\
&=  [FP f_0 - (FP f_0 + FN f_1) \bar Y] - f [ FP - (FP+FN) \bar Y ] \\
&=  - FP \Delta \bar Y + FP \bar Y^2 \Delta - FN \bar Y^2 \Delta \\
&=  - \Delta \bar Y (FP \cdot (1-\bar Y) + FN \cdot \bar Y) \\
\end{aligned}
$$
where $f = f_1 \bar Y + f_0 (1-\bar Y)$ so $f_0 - f = -\Delta \bar Y$ and $f_1 - f = \Delta (1-\bar Y)$.
$$
\begin{aligned}
C(I, Y) &= E[ I Y ] - f \bar Y \\
&=  f_1 \bar Y + f_0 (1-\bar Y) - f \bar Y \\
&=  f_0 (1-\bar Y) + \Delta (1-\bar Y) \bar Y \\
&= (1-\bar Y) (f_0 + \Delta \bar Y)
\end{aligned}
$$
Combining yields
$$
\begin{aligned}
\rho_{I,PZ} &= \rho_{I,Y} \times \frac{- \Delta \bar Y (FP \cdot (1-\bar Y) + FN \cdot \bar Y) }{(1-\bar Y) (f_0 + \Delta \bar Y)} \times \sqrt{ \frac{(1-\bar Y)}{2 ( FP (1-\bar Y) + FN \cdot \bar Y)} } \\
&= - \rho_{I, Y} \times \Delta \times \sqrt{\frac{\bar Y}{1-\bar Y}} \frac{\sqrt{FP(1-\bar Y) + FN \cdot \bar Y}}{f_0 (1-\bar Y) + f_1 \bar Y} \times \sqrt{\frac{\bar Y}{2}}
\end{aligned}
$$
We can then re-write $\rho_{I,Y} \sigma_Y + \rho_{I,PZ} \sigma_{PZ}$ as
$$
\rho_{I,Y} \sigma_Y \left( 1 - \Delta \times \frac{\bar Y}{1-\bar Y} \times \frac{FP(1-\bar Y) + FN \cdot \bar Y}{f_0 (1-\bar Y) + f_1 \bar Y} \right)
$$
Suppose $f_1 = M \cdot f_0$.  Then $\Delta = f_1 - f_0 = f_0 (M-1)$ and $f_0 (1-\bar Y) + f_1 \bar Y = f_0 ( (1-\bar Y) + M \bar Y )$ and we can re-write above as
$$
\rho_{I,Y} \sigma_Y \left( 1 - (M-1) \times \frac{\bar Y}{1-\bar Y} \times \frac{FP(1-\bar Y) + FN \cdot \bar Y}{(1-\bar Y) + M \bar Y} \right).
$$

\subsection*{Ratio estimator}
\label{app:ratio}

Let ${\bf u} = (u_{t-1},u_t) \in \mathbb{R}^2$ and $g({\bf u}) = \frac{u_t}{u_{t-1}}$, i.e., a differentiable function $g:\mathbb{R}^2 \to \mathbb{R}$. Centering a Taylor series expansion of second-order around coordinates $(U_2, U_1) \in \mathbb{R}^2$ yields
$$
\begin{aligned}
g({\bf u}) =& g(U_{t-1}, U_t) - \frac{U_t}{U_{t-1}^2} (u_{t-1} - U_{t-1}) + \frac{1}{U_{t-1}} (u_t - U_t) \\
&+ \frac{1}{2} \left[ \frac{2 U_t}{U_{t-1}^3} (u_{t-1} - U_{t-1})^2 + 0 \times (u_t - U_t)^2 - 2 \times (u_{t-1} - U_{t-1}) (u_t - U_t) \frac{1}{U_t^2} \right]
\end{aligned}
$$
Plugging in $(\bar y_{t-1}, \bar y_t)$ for $(u_{t-1}, u_t)$ and $(\bar Y_{t-1}, \bar Y_t)$ for $(U_{t-1}, U_t)$ yields the $\frac{\bar y_t}{\bar y_{t-1}} - \frac{\bar Y_t}{\bar Y_{t-1}} $ is equal to
$$
\begin{aligned}
=&
- \frac{\bar Y_t}{\bar Y_{t-1}^2} (\bar y_{t-1} - \bar Y_{t-1}) + \frac{1}{\bar Y_{t-1}} (\bar y_t - \bar Y_t) \\
&+ \frac{\bar Y_t}{\bar Y_{t-1}^3} (\bar y_{t-1} - \bar Y_{t-1})^2 -  (\bar y_{t-1} - \bar Y_{t-1}) (\bar y_t - \bar Y_t) \frac{1}{\bar Y_{t-1}^2} \\
&= \frac{\bar Y_t}{\bar Y_{t-1}} \bigg[  \rho_{I_t,Y_t} \sqrt{\frac{1-f_t}{f_t}} CV (Y_t)  -\rho_{I_{t-1},Y_{t-1}} \sqrt{\frac{1-f_{t-1}}{f_{t-1}}} CV (Y_{t-1}) \\
&+ \rho^2_{I_{t-1},Y_{t-1}} \frac{1-f_{t-1}}{f_{t-1}} CV^2 (Y_{t-1}) -  \rho_{I_{t-1},Y_{t-1}} \sqrt{\frac{1-f_{t-1}}{f_{t-1}}} CV (Y_{t-1}) \times
\rho_{I_t,Y_t} \sqrt{\frac{1-f_t}{f_t}} CV (Y_t)   \bigg] \\
&= \frac{\bar Y_t}{\bar Y_{t-1}} \bigg[ \rho_{I_t,Y_t} \sqrt{\frac{1-f_t}{f_t}} CV (Y_t)  -\rho_{I_{t-1},Y_{t-1}} \sqrt{\frac{1-f_{t-1}}{f_{t-1}}} CV (Y_{t-1}) \bigg] \left[ 1 - \rho_{I_{t-1},Y_{t-1}} \sqrt{\frac{1-f_{t-1}}{f_{t-1}}} CV (Y_{t-1}) \right]
\end{aligned}
$$
where the second equality is obtained by plugging in the statistical decomposition of the error for both time points and the coefficient of variation being defined as $CV(Y) := \sigma_Y/\mu_Y$.  Under measurement error, the extra terms $D_{2}$ and $D_1$ can be inserted in the correct locations.

\subsubsection*{Estimation of effective reproduction number}

Let
$$
\delta_t := \bigg[ \rho_{I_t,K_t} D_{M_t} \sqrt{\frac{1-f_t}{f_t}} CV (K_t)  -\rho_{I_{t-1},K_{t-1}} D_{M_{t-1}} \sqrt{\frac{1-f_t}{f_t}} CV (K_{t-1}) \bigg].
$$
Then results from Section~\ref{app:ratio} show that the estimate of the number of new cases on day t is given by
$$
\frac{S_t \cdot \bar y_t}{S_{t-1} \cdot \bar y_{t-1}} =
\frac{K_t}{K_{t-1}} \left( 1 + \delta_t \times \left[ 1 - \rho_{I_{t-1},K_{t-1}} D_{M_{t-1}} \sqrt{\frac{1-f_t}{f_t}} CV (K_{t-1}) \right] \right)
$$
Then setting $e_t = \delta_t \times [1 - \rho_{I_{t-1},K_{t-1}} D_{M_{t-1}} \sqrt{\frac{1-f_t}{f_t}} CV (K_{t-1}) ]$, we have
$$
\begin{aligned}
\log \left( \frac{S_t \bar y_t}{S_{t-1} \bar y_{t-1}} \right) - \log \left( \frac{K_t}{K_{t-1}} \right) &= \log (1 + e_t) \\
\log \left( \frac{\bar y_t}{\bar y_{t-1}} \right) - \log \left( \frac{K_t}{K_{t-1}} \right) &= 1 + e_t - \log \left( \frac{S_t}{S_{t-1}} \right) \\
1 + \frac{1}{\gamma} \log \left( \frac{\bar y_t}{\bar y_{t-1}} \right) - \left[ 1 + \frac{1}{\gamma} \log \left( \frac{K_t}{K_{t-1}} \right) \right] &= \frac{1}{\gamma} \left[ \log \left( 1 + e_t \right) - \log \left( \frac{S_{t}}{S_{t-1}} \right) \right] \\
\Rightarrow \hat R_t - R_t &= \frac{1}{\gamma} \left[ \log \left( 1 + e_t \right) - \log \left( \frac{S_{t}}{S_{t-1}} \right) \right]
\end{aligned}
$$

\subsection*{Derivation for the effective sample size}
\label{section:effss}

For binary outcomes, we have
\begin{equation} \label{eq:binaryrho}
\rho_{I,Y} = \Delta \sqrt{\frac{\bar Y (1 - \bar Y)}{f (1-f)} }
\end{equation}
where $\Delta = P_J (I_J = 1 \mid Y_J = 1) - P(I_J = 1 \mid Y_J = 0) = f_1 - f_0$.  Suppose that $M = f_1/f_0$; then
$f_0 = f / (\bar Y \cdot (M-1) + 1)$.  For $\bar Y = 0.091$, $M = 2$ and $f = 0.026$ we have $f_0 = 0.024$ and $f_1 = 0.048$ which implies $\Delta = 0.024$.
Then $\rho_{I,Y} = 0.043$.  Then $E_{\I} [ \rho_{I,Y}^2 ] \leq (E_{\I} [\rho_{I,Y} ]^2 = 0.0019$.  Then
$$
n_{eff} \leq \frac{f}{1-f} \frac{1}{0.0019} = 14.39.
$$
Using this computation, we compute effective sample size under a range of values $\bary$, $M$ with $f = 0.026$ (i.e., current sampling fraction) and $f =1/7$ (sample individuals roughly every 7 days).

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
& \multicolumn{5}{c}{$M$} \\ \cline{2-6}
$\bar y$ & 1.2 & 1.4 & 1.6 & 1.8 & 2 \\
  \hline
0.016 & 1598.00 & 402.00 & 180.00 & 102.00 & 66.00 \\
  0.036 & 731.00 & 185.00 & 84.00 & 48.00 & 31.00 \\
  0.056 & 484.00 & 124.00 & 56.00 & 32.00 & 21.00 \\
  0.076 & 367.00 & 94.00 & 43.00 & 25.00 & 16.00 \\
  0.096 & 299.00 & 78.00 & 36.00 & 21.00 & 14.00 \\
   \hline
\end{tabular}
\end{table}

We also present the same plot under $FP = 0.005$ and $FN = 0.172$ to show the impact of measurement error on effective sample size.

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
  & \multicolumn{5}{c}{$M$} \\ \cline{2-6}
$\bar y$ & 1.2 & 1.4 & 1.6 & 1.8 & 2 \\
  \hline
0.016 & 1154.00 & 290.00 & 130.00 & 73.00 & 47.00 \\
  0.036 & 528.00 & 134.00 & 60.00 & 34.00 & 22.00 \\
  0.056 & 349.00 & 89.00 & 41.00 & 23.00 & 15.00 \\
  0.076 & 265.00 & 68.00 & 31.00 & 18.00 & 12.00 \\
  0.096 & 216.00 & 56.00 & 26.00 & 15.00 & 10.00 \\
   \hline
\end{tabular}
\end{table}

Materials and Methods\\
Supplementary Text\\

Figs. S1 to S3\\
Tables S1 to S4\\
References \textit{(4-10)}

% For your review copy (i.e., the file you initially send in for
% evaluation), you can use the {figure} environment and the
% \includegraphics command to stream your figures into the text, placing
% all figures at the end.  For the final, revised manuscript for
% acceptance and production, however, PostScript or other graphics
% should not be streamed into your compliled file.  Instead, set
% captions as simple paragraphs (with a \noindent tag), setting them
% off from the rest of the text with a \clearpage as shown  below, and
% submit figures as separate files according to the Art Department's
% instructions.

\clearpage

\noindent {\bf Fig. 1.} US total tests and total positive tests per day\\
\noindent {\bf Fig. 2.} US total tests and total positive tests per day\\
\noindent {\bf Fig. 3.} US total tests and total positive tests per day\\
\noindent {\bf Fig. 4.} US total tests and total positive tests per day\\
\noindent {\bf Fig. 5.} US total tests and total positive tests per day

% Please do not use figure environments to set
% up your figures in the final (post-peer-review) draft, do not include graphics in your
% source code, and do not cite figures in the text using \LaTeX\
% \verb+\ref+ commands.  Instead, simply refer to the figure numbers in
% the text per {\it Science\/} style, and include the list of captions at
% the end of the document, coded as ordinary paragraphs as shown in the
% \texttt{scifile.tex} template file.  Your actual figure files should
% be submitted separately.

\end{document}