### 2021-09-01 write-up/covid-writeup.tex
Updated SEIR one last time

\documentclass[11pt]{amsart}

\RequirePackage[OT1]{fontenc}

\usepackage[foot]{amsaddr}
% \usepackage{biblatex}
\usepackage{natbib}
%\setcitestyle{author-year,open={((},close={))}} %Citation-related commands
%\bibliographystyle{apalike}
\usepackage{multirow}
\usepackage{enumitem}
\usepackage{caption,subcaption}

\usepackage{xr}
\makeatletter
\newcommand*{\addFileDependency}[1]{% argument=file name and extension
  \typeout{(#1)}% latexmk will find this if $recorder=0 (however, in that case, it will ignore #1 if it is a .aux or .pdf file etc and it exists! if it doesn't exist, it will appear in the list of dependents regardless)
  \@addtofilelist{#1}% if you want it to appear in \listfiles, not really necessary and latexmk doesn't use this
  \IfFileExists{#1}{}{\typeout{No file #1.}}% latexmk will find this message if #1 doesn't exist (yet)
}
\makeatother

\newcommand*{\myexternaldocument}[1]{%
    \externaldocument{#1}%
    \addFileDependency{#1.tex}%
    \addFileDependency{#1.aux}%
}
%%% END HELPER CODE

% put all the external documents here!
\myexternaldocument{covid-supplement-writeup}

\usepackage{tchdr}
\boldshortcuts

\usepackage{url}

\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{setspace}
\DeclareGraphicsExtensions{.eps, .ps}
\usepackage{amsmath, amsthm, amsfonts}
% \usepackage[margin=1.5in]{geometry}

\numberwithin{equation}{section}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
% \newtheorem{example}{Example}
 \newtheorem{remark}{Remark}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{criterion}[theorem]{Criterion}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}

\setlength{\textwidth}{6.5in}
\setlength{\textheight}{8.5in}
\setlength{\topmargin}{-0.25in}
\setlength{\evensidemargin}{0in}
\setlength{\oddsidemargin}{0in}
%\setlength{\evensidemargin}{.25in}
%\setlength{\oddsidemargin}{.25in}

\usepackage{setspace}
% \onehalfspacing
\doublespacing

\def\pr{\text{pr}}
\def\sgn{\text{sgn}}
\def\I{\bf I}
\def\code#1{\texttt{#1}}

\begin{document}

\title[Statistical paradoxes in coronavirus case-counts]{Statistical paradoxes in coronavirus case-counts: selection bias, measurement error, and the COVID-19 pandemic} %

\author{Walter Dempsey}
\address{Department of Biostatistics, University of Michigan, Ann Arbor, MI 48109}
% \email{wdem@umich.edu}

\begin{abstract}
  Coronavirus case-count data has influenced government policies and drives most epidemiological forecasts. Limited testing is cited as the key driver behind minimal information on the COVID-19 pandemic. While expanded testing is laudable, measurement error and selection bias are the two greatest problems limiting our understanding of the COVID-19 pandemic; neither can be fully addressed by increased testing capacity. In this paper, we demonstrate their impact on estimation of point prevalence and the effective reproduction number. We show that estimates based on the millions of molecular tests in the US has the same mean square error as a small simple random sample.  To address this, a procedure is presented that combines case-count data and random samples over time to estimate selection propensities based on key covariate information. We then combine these selection propensities with epidemiological forecast models to construct a \emph{doubly robust} estimation method that accounts for both measurement-error and selection bias.  This method is then applied to estimate Indiana's prevalence using case-count, hospitalization, and death data with demographic information, a statewide random molecular sample collected from April 25--29th, and Facebook's COVID-19 symptom survey.  We end with a series of recommendations based on the proposed methodology.
\end{abstract}

\maketitle

\newpage

\section{Introduction}
The World Health Organization has declared the coronavirus disease 2019 (COVID-19) a public health emergency.  As of July 29th, 2021, over 196 million cases have been confirmed worldwide with 34.8 million cases and over 612 thousand confirmed deaths across the United States. This pandemic
has become the focal point of everyday life; yet the data landscape for understanding COVID-19 remains limited.  Public databases~\citep{JHU_Lancet,NYT} provide incoming county-level information of confirmed cases and deaths.  Statisticians, epidemiologists, economists, and data scientists have used this granular data to forecast COVID-19 case-counts, deaths, and hospitalizations~\citep{Giordano2020,Song2020,Ray2020,2020.IHME,Wang2020.03,JTD36385}.

This paper has two main objectives.  The first objective is to express reservations at the use of observed case-counts as a proxy for disease prevalence and in estimation of standard epidemiological models for inference and forecasting.  The reason is straightforward: observed case-count data is plagued by selection bias and measurement error. Through a series of calculations, we will demonstrate that the information gained from increasing testing capacity is limited in the presence of selection bias and when testing inaccuracies persist.  In particular, the millions of tests in the US have a small effective sample size when compared to random sampling. These calculations demonstrate the importance of probabilistic sampling designs over time for estimation of point prevalence and effective reproduction number.

Due to monetary and time constraints, however, random testing is performed infrequently.  As of June 2021, Indiana and Ohio are the only states to conduct statewide random sample testing\footnote{These are is the only random samples to collect both seroprevalence and diagnostic testing results. The CDC and other states have conducted seroprevalence-only studies.}. Indiana's sample was collected from April 25--29, 2020~\citep{Yiannoutsos2021}.  Such infrequent random testing is likely to provide insufficient information to help researchers and policy makers better understand the disease trajectory which can change rapidly over time.  Therefore, while random testing may be preferable in theory, in practice governments, researchers, and policy makers continue to use coronavirus case-counts to understand the impact of COVID-19 on the population and make data-informed decisions.

The second objective is to demonstrate how random samples provide necessary \emph{auxiliary information} to address selection bias in coronavirus case-count data.  Random samples provide the necessary covariate information from a representative sample from the population to estimate selection propensities. These propensities can then used in an inverse-probability weighting scheme to construct estimators of disease prevalence that attempt to control for selection bias.  A doubly robust extension allows researchers to combine these estimates with epidemiological forecasts based on compartmental models that are common in the study of infectious diseases~\citep{Hao2020,Song2020,Ray2020,Johndrow2020}.

The proposed approach requires covariate information to be collected on individuals who receive a COVID-19 test.  Unfortunately, many states do not require or report auxiliary covariate information beyond basic demographic information (e.g., gender, age, race, and ethnicity).  We end with a brief list of suggestions of minor changes to current practice based on the proposed methodology. While we demonstrate empirical improvements over simple disease prevalence estimates, we also highlight how selection bias may persist and impact uncertainty quantification.

 \subsection{Related work}

 This article discusses the relationship between three statistical concepts: selection bias, measurement error, and population size.  While the impact of measurement error~\citep{Smeden2019} and selection bias~\citep{Keiding2016} on estimation are both well-studied topics, here we provide a new perspective by building on the work of \cite{Meng2018} who studied an error decomposition to understand the relationship between selection bias and population size.  Specifically, we quantify the interaction between measurement-error and selection bias on statistical error, showing how the sign and magnitude can change drastically.
 % the impact on the effective sample size can therefore be quite large.

 After demonstrating the limitations of case-count analysis when compared to random sampling, we then assess whether there is potential for combining the \emph{nonprobability samples} with \emph{probability samples} to improve point prevalence estimation. For any probability sampling design, the Horvitz-Thompson estimator~\citep{HT1952} incorporates design information via inverse-probability weights (IPW).  For nonprobability samples, the IPW estimator requires modelling the propensity scores.  Its use in the survey context is also referred to as quasi-randomization~\citep{Elliott2017}. \cite{Valliant2011} consider a weighted logistic regression procedure using the pooled probability and nonprobability samples.  \cite{Chen2019} consider a pseudo-log-likelihood approach that uses the random samples as a proxy for a term in the log-likelihood.  Here, we extend this approach to account for measurement-error as well as observing random samples at multiple times. We then provide an extension of the statistical error decomposition and discuss the trade-offs inherent in such a weighting approach.

 One core component of coronavirus research is epidemiological compartmental modelling of case-count and death data. These models can be used to answer a variety of research questions including case-count forecasting, estimation of the effective reproduction number, and estimation of quarantine and other health policies on infectious disease dynamics.  The basic approach is a deterministic compartmental model called the susceptible-infectious-recovered (SIR) model.  A probabilistic extension was proposed by~\cite{Osthus2017} to model one-dimensional time series of infected proportions. \cite{Song2020} extended this approach to incorporate interventions and assess interventions on COVID-19 epidemic in China. \cite{Hao2020} extends this work further to account for various presymptomatic infectiousness, time-varying ascertainment rates, transmission rates and population movements.

Given a probability sampling design, individual predictions can be leveraged to improve estimation via model-assisted approaches~\citep{Breidt2017}.  For nonprobability samples, \cite{Chen2019} derive a doubly robust approach that uses outcome predictions given covariates on the nonprobability and probability samples.  Here, we combine the compartmental model of \cite{Song2020} but instead, as in \cite{Johndrow2020}, focus on COVID-19 confirmed death count data.  We generate epidemiological forecasts for point prevalence of each population strata.  We then demonstrate how to combine these forecasts with the IPW approach to construct doubly-robust estimates of point prevalence.  A derived statistical error decomposition guides this discussion.

Recent work by \cite{Zhao2021} pointed out that estimation of key epidemiological parameters such as the incubation time using standard epidemiological models can suffer from severe bias due to issues beyond selection bias and measurement error.  Namely, right truncation and epidemic growth leading to patients ``being more likely to be infected towards the end of their exposure period''~\citep[pp. 3]{Zhao2021}.  Their approach carefully constructs the study sample and statistical model to account for these issues.  In this paper, rather than focusing on construction of a study sample, we ask whether one can collect auxiliary information to address selection bias in the observed case count data directly.

 % \subsection{Outline}

 % In Section~\ref{section:data}, we summarize the publicly available data on COVID-19.  We highlight how testing information is collected as well as demographic and comorbidity information.  In Section~\ref{section:testinginfo}, we briefly summarize COVID-19 testing.  We clarify the difference between viral and antibody testing.  For viral testing, we briefly discuss the reasons behind measurement-error in RT-PCR tests including test timing and sample collection techniques.

 % In Section~\ref{section:casecount}, we present simple mathematical arguments to show that unadjusted prevalence rates are (unsurprisingly) biased when tests are imperfect. What is surprising is that the direction and magnitude of the bias can vary substantially and interacts with the sampling fraction.  We show that the rate of change in COVID-19 observed case-counts overestimates the true rate of change prior to the peak time and underestimates it immediately after.  This implies a data scientist analyzing observed rates will be (on average) overly pessimistic in the early stages of the pandemic, and overly optimistic subsequent to the peak times.  We show that estimates of the effective reproduction number may be similarly impacted.

 % In Section~\ref{section:improvedcasecount}, we present a doubly robust estimation method for combining epidemiological forecasts with inverse-probability weighting methods to estimate prevalence.  Weight estimation requires auxiliary information provided by random samples.  A statistical error decomposition illustrates a trade-off in use of weights: a potential increase in data quality but a reduction in effective data quantity.

 % In Section~\ref{section:applications}, the IPW and doubly-robust estimation methods are applied to the case study of Indiana statewide prevalence from March 2020 until June 2021.  We estimate data quality for IPW estimates using the random sample on April 28th.  This demonstrates the proposed approach is an improvement over simple prevalence estimation but that data quality is not of the same order as random sampling.

 % In Section~\ref{section:discussion}, a series of recommendations are provided that aim to help improve statistical inference in settings with large nonprobability samples from a stochastic process where small random sampling over time is possible.

 \section{COVID-19 testing and data}
 \label{section:data}

 Here we provide the necessary background to understand COVID-19 diagnostic testing, its scientific use in managing the pandemic, and the data streams considered in this manuscript.

 \subsection{Diagnostic testing}
 \label{section:testinginfo}

 Upon infection with SARS-CoV-2, an incubation period starts and lasts approximately five days~\citep{Lauer2020} before the viral load is high enough to be detected.  A \emph{molecular test} refers to diagnostic tests that aim to detect increased viral load (e.g., RT-PCR or antigen tests).  If someone has an active infection then after the incubation period, a molecular test with perfect sensitivity will yield a positive result for the next few weeks.  After that, the viral load will decrease and a molecular test with perfect specificity will come back negative. While an infected individual may yield a positive molecular test for several weeks, they will likely stop transmitting the disease within a few days of infection, meaning a positive molecular test does not imply transmissibility.

 A molecular test conducted on an infected individual may return a negative result.  Such false negatives are very strongly associated with when the test is conducted.  In the incubation phase, most molecular tests will return a false negative result.  Molecular tests are most sensitive when the viral loads are highest which occurs during the first few days of transmissibility~\citep{Mina2020}.   Moreover, most molecular tests are performed via nasopharyngeal swab.  Specimen collection by swab is known to impact false negative/positive rates regardless of test timing.   Systematic reviews suggest that 87\% sensitivity and 97.6\% specificity are reasonable estimates for current RT-PCR tests~\citep{Arevalo2020, Woloshin2020,Cohen2020}.

 The primary goal of molecular tests is diagnosis of active infections in the population.  Such diagnostic tests generate important information about the presence of SARS-CoV-2 in the population, and help scientists and policy-makers understand patterns of transmission and propagation. Rapid and frequent molecular testing is cited as a key component~\citep{OECD2021} in effective strategies to identify active infections and prevent systemic outbreaks.

 \subsection{Publicly available data on COVID-19}
 \label{subsection:testinginfo}

 The primary goal of this manuscript is to produce accurate estimates of the population-level active infection rates over time using \emph{publicly available} viral testing data.  Secondary goals include estimation of rates of change and the effective reproduction number which characterize disease trajectory.
 These quantities are fundamental to public health policy and provide critical information on the presence and transmission of SARS-CoV-2.

 % {\bf key point: WHAT IS THE GOAL OF CASE COUNT DATA? WHAT IS THE GOAL OF SEROPREVALENCE STUDIES?  are we using it for that reason. cse count data is colelcted as part of surveillance programs wehre teh goal is to identify and isolate active infections.  However, many people are trying to use these to udnerstand the disease.  This is our issue.}

 Coronavirus case-count data refers to the number of positive molecular tests performed on each day.  Figure~\ref{fig:in-cases} plots the number of reported confirmed COVID-19 cases per day in the state of Indiana.  Figure~\ref{fig:in-tests} plots the total number of COVID-19 molecular tests performed per day. Figure~\ref{fig:in-deaths} plots the total number of COVID-19 related reported deaths per day. Public databases maintained by \href{https://bit.ly/2UqFSuA}{Johns Hopkins University} and the \href{https://bit.ly/2vUHfrK}{New York Times} provide accessible incoming county-level information of confirmed cases and deaths.

 Public databases most often, however, only contain aggregate information.  The Johns Hopkins dashboard, for example, provides demographic breakdown of case counts as well as the total confirmed cases and deaths by county.  Aggregate time series of case count and deaths can also be extracted.  Unfortunately, most dashboards do not provide demographic information on who requested a test nor on who tested positive for SARS-CoV-2 \emph{over time}. Working closely with the State of Indiana, we were able to obtain access to COVID-19 total tests, positive tests, and related deaths per day broken out marginally by age, gender, ethnicity, and race~\citep{IndianaData2021}.  These granular datasets are now publicly available and motivate the proposed approach.
 % https://www.regenstrief.org/covid-dashboard/
 %

\begin{remark}[Reporting Delays]
Figure~\ref{fig:in-data} shows clear reduced testing and case count reporting on weekends compared to weekdays. Moreover, COVID-19 tests are reported on the day they were administered, while case counts are reported based on the date the positive test was reported to and confirmed by the Indiana Department of Health system.  To minimize the impact of reporting on analysis, testing and case-count data are aggregated at the weekly level for analysis.
 \end{remark}

 %https://www.npr.org/sections/goatsandsoda/2020/05/31/865932474/should-i-get-tested-for-coronavirus-just-for-the-heck-of-it

 \subsubsection{Testing restrictions and public health policy in Indiana}
 \label{section:publicpolicyindiana}

 Due to limited testing capacity, many US states instituted testing restrictions early on in the pandemic.
 % on who can request a molecular test at a given time.
 %https://www.coronavirus.in.gov/
 % Early on, limited testing meant that only those at high risk were allowed to be tested (i.e., symptomatic or high risk populations); as testing capacity expanded, these restrictions were gradually lifted.
 Here, we reconstruct the testing restriction history for the state of Indiana. As of May 12th, 2020, the Indiana State Department of Health (ISDH) laboratory was testing high-risk individuals for COVID-19. Any symptomatic individual could receive a test.  Moreover there were key factors for people who can be tested regardless of symptom status: over the age of 65, diabetic, high-blood pressure, obesity, pregnancy, and minority populations at greater risk of severe illness~\citep{indystar2020}.
 % https://www.indystar.com/story/news/health/2020/05/12/coronavirus-testing-indiana-who-should-get-tested/3110592001/
 On April 28th, 2020, the criteria expanded to include any Indiana resident with virus symptoms, people in close contact with those who have tested positive and residents of congregant communities~\citep{wishtv2020}.
 % https://www.wishtv.com/news/medical/indiana-opens-up-covid-19-testing-to-more-hoosiers/
 On June 15th, 2020, ISDH lifted all restrictions and testing expanded to include anyone in Indiana who wants to be tested for coronavirus~\citep{indystar2020v2}.
 % https://www.indystar.com/story/news/health/2020/06/12/indiana-says-anyone-who-wants-coronavirus-test-can-get-one/3179151001/
 Testing restrictions impact the propensity of an individual to receive a test and therefore, if COVID-19 positive, contribute to the case count.  In this paper, testing restrictions are addressed by fitting time-varying testing propensities that depend on relevant covariate information.

 On March 23rd, 2020, Indiana's governor issued a \emph{stay at home} order effective March 26th through April 5th.  The order was extended until April 30th. On May 1st, 2020, a five-stage plan for gradual reopening was announced by Governon Holcomb~\citep{fivestageplan}.  Such policies target reduction in active infection rates. In this paper, Indiana's public health policy is incorporated in our construction of epidemiological forecasts in Section~\ref{section:modelbased}.

 \subsection{Probabilistic samples in Indiana}

 Due to testing restrictions and other potential selection biases, publicly reported COVID-19 case count data may not be sufficient to understand active infection rates or the disease trajectory.  Here, we discuss two random samples that provide auxiliary information and may help address selection bias.

 \subsubsection{Random statewide testing}

 Between April 25--29, 2020, Indiana conducted statewide random molecular testing of persons ages $\geq 12$ years to assess prevalence of active infection to SARS-CoV-2~\citep{Yiannoutsos2021}. A stratified random sampling design was conducted using Indiana’s 10 public health preparedness districts as sampling strata. 15,495 participants were contacted resulting in a final sample size of 3,658. Demographic data was collected (e.g., summary statistics on age, sex, and race) as well as data on whether they experienced any COVID-19 compatible symptoms during the past 2 weeks or had shared a household with someone who had a positive test result for SARS-CoV-2. During May 2--3, 2020, an additional non-random sample of 898 individuals was also collected.
 Table~\ref{tab:indiana} summarizes the data.
 % In this paper, the nonrandom sample is used to assess whether the cumulative summary statistics are appropriate in constructing the inverse-probability weights as well as in constructing appropriate weights using symptom information.

 \subsubsection{Facebook/CMU symptom survey.}
 \label{subsection:fbsymptom}
 Since April 2020, Facebook has started randomly showing an opt-in symptom survey to its users~\citep{delphisurvey}.  Data collected includes basic demographic information and if the respondent has symptoms such as fever, coughing, shortness of breath, or loss of smell which are associated with COVID-19.  The survey defines an individual as displaying \emph{COVID-like symptoms} if they exhibit a fever along with a cough, or shortness of breath, or difficulty breathing.  Figure~\ref{fig:fbsymptoms} displays smoothed estimates of the fraction of individuals within age and gender strata who report a fever within the past 24-hours.  This paper uses the Facebook symptom survey as the main source of auxiliary information on time-varying characteristics of the population of Indiana, e.g., fraction of population displaying symptoms or in contact with COVID-19 positive individuals.

 \begin{figure}[!th]
 \centering
 \includegraphics[width = 0.9\textwidth]{../figs/fbcovid19symptoms.png}
 \caption{Rate of reported COVID-19 symptoms per strata.  Daily rates were estimated using weighted method suggested by~\citep{delphisurvey} on each day separately and then smoothed over time using local-linear nonparametric regression.}
 \label{fig:fbsymptoms}
 \vspace{-0.3cm}
 \end{figure}

 %% GOOD THROUGH HERE

 \section{Analysis of case-count data}
 \label{section:casecount}

 We start with some simple notation.  Let $N$ denote the population size.  At a given time, let $Y_j$ denote COVID-19 status for the $j$th individual in the population, $j=1,\ldots, N$. Here, like in survey methodology~\citep{Cochran77}, we treat COVID-19 status as a fixed but unknown quantity of interest. For simplicity, we start by ignoring the dynamic nature of the outbreak and recoverability of individuals. We assume either individual $j$ is COVID-19 positive and $Y_j=1$ or is COVID-19 negative and $Y_j=0$. We also let $I_j \in \{0,1\}$ be an indicator that the individual was tested ($I_j = 1$) or not ($I_j=0$).

 To start, we assume the overall number of active COVID-19 cases and/or active infection rate (AIR) are of primary interest. That is, we are interested in either the population total $Y = \sum_{j=1}^N Y_j$ or the population average $\bar Y = Y/N$. Suppose that $n$ tests are performed and we observe the values $y_1, \ldots, y_n \in \{0,1\}$.  Then a natural candidate for AIR is the proportion of positive tests $\bar y = \frac{1}{n} \sum_{i=1}^n y_i$, and a natural candidate for overall active cases is $N \times \bar y$.
 Under simple random sampling (SRS) or any other epsem\footnote{equal probability of selection method} design, the above are unbiased estimators of the population-level quantities of interest.  Under SRS, the variance of the estimator can be expressed as $\frac{1}{N-1} \times \frac{1-f}{f} \times \sigma_Y^2$ where $f = n/N$ is the sampling fraction and $\sigma_Y^2 = \frac{1}{N} \sum_{j=1}^N (Y_j - \bar Y)^2 = \bar Y (1- \bar Y)$.

 These random selection mechanisms are independent of the outcome of interest. When this is not the case, selection effects may cause bias. To better understand this issue, \cite{Meng2018} recently provided the following intuitive and powerful statistical decomposition of the error between $\bar y$ and the true proportion $\bar Y$
 $$
 \bar y_n - \bar Y =  \rho_{I, Y} \times \sqrt{\frac{1-f}{f}} \times \sigma_Y.
 $$
 The first term represents \emph{data quality}, the second \emph{data quantity}, and the third \emph{problem difficulty}. The term $\rho_{I,Y}$ is the empirical correlation between the population values~$\{ Y_j \}_{j=1}^N$ and the selection values $\{ I_j \}_{j=1}^N$.  Under simple random sampling, $E_{\I} [ \rho_{I,Y} ] = 0$, where the expectation is with respect to the selection mechanism~$\I$, so there is no bias.  The SRS variance formula above shows that $E_{\I} [ \rho_{I,Y}^2 ]  = 1/(N-1)$.

 % The key issue with selective testing is that $E_{\I} [ \rho_{I,Y} ] \neq 0$.  Meng identified this as the fundamental issue that can lead to paradoxes in the analysis of big data. The remainder of this paper aims to build upon this fundamental work by extending the decomposition in two directions: accounting for measurement error and the temporal nature of the pandemic.  We then discuss effective sampling methods and their importance in the current pandemic.

 \subsection{Imperfect testing}
 \label{section:imperfecttesting}

 Tests are imperfect.  As discussed in Section~\ref{section:testinginfo}, COVID-19 testing is no exception. Here we investigate the interplay between imperfect testing and selection bias.  Researchers most often assume measurement error will lead to parameter attenuation.  When paired with selection bias, however, the two sources of error become entangled, and resulting errors can become magnified, muted, or even switch signs.

 First we require some additional notation.  Let $P_j$ be an indicator of measurement error, equal to $1$ when we incorrectly measure the binary outcome and $0$ when we observe the true binary outcome. We suppose this is a stochastic variable where $\pr(P_j = 1 \mid Y_j = 1) =: FN$ is the false-negative rate and $\pr(P_j = 1 \mid Y_j = 0) =: FP$ is the false-positive rate.  If individual $j$ is selected (i.e., $I_j = 1$) then the observed outcome can be written as $Y_j^{\star} = Y_j(1-P_j) + (1-Y_j) P_j$.  The attentive data analyst will recognize the estimator $\bar y_n$ is now biased for simple random samples.  In Appendix~\ref{app:memestimator}, assuming sensitivity and specificity are known a priori, a novel iterative procedure is used to construct the estimator $\tilde y_n = (\bar y_n - FP)/(1-(FP+FN))$, which is unbiased under simple random sampling (SRS); see Appendix~\ref{app:modelbased} for a discussion of the connection to model-based estimators. To understand the impact of selection bias and imperfect testing, we derive the following statistical decomposition of the error between $\tilde y_n$ and $\bar Y$:
 \begin{equation}
 \label{eq:statdecomp}
 \rho_{I,Y} \times \sqrt{\frac{1-f}{f}} \times \sigma_{Y}
 \times \underbrace{\left[ 1 - \Delta \times \frac{\bar Y}{1-\bar Y} \times \frac{FP(1-\bar Y) + FN \cdot \bar Y}{f_0 (1-\bar Y) + f_1 \bar Y} \right] \times \frac{1}{1-(FP+FN)}}_{D_M},
 \end{equation}
 where $f=n/N$ is the sampling fraction, $f_1$ and $f_0$ are sampling fractions for COVID-19 positive and negative individuals respectively, and $\Delta = f_1 - f_0$ is the sampling rate differential.  See Appendix~\ref{app:memestimator} for the derivation. This extends work by \cite{Meng2018} to account for imperfect testing. The first three terms continue to represent \emph{data quality}, \emph{data quantity}, and \emph{problem difficulty} respectively.  The new term~$D_M$ represents the \emph{imperfect testing adjustment} which is a complex function of the sampling rate differential, the odds ratio, and the ratio of measurement error interaction with prevalence and sampling rates interaction with prevalence, which is scaled by the factor $(1 - FP - FN)^{-1}$.

 Figure~\ref{fig:heatmap} shows that $D_M$, as a function of the relative frequency ($f_1/f_0$) and log odds ratio, can be both positive and negative as well as a range of magnitudes~\citep{Beesley2020,Beesley2019,Smeden2019}. Assuming no measurement error, $D_M = 1$ so the relation between estimation and selection bias is simple, e.g., if COVID-19 positive individuals were more likely to receive test then this implies upward bias in prevalence estimates. Under random testing (i.e., $f_0 = f_1$), $D_M = (1-FP-FN)^{-1}$ so measurement error simply magnifies this error. When tests are imperfect and selection bias exists, this simple relationship no longer holds.

 \begin{figure}[!th]
 \centering
 \includegraphics[width = 0.7\textwidth]{../figs/mem_heatmap.png}
 \caption{Imperfect testing adjustment: relative frequency $f_1/f_0$ (x-axis) against odds ratio (y-axis) for $FP=0.024$ and $FN=0.13$. Color scaled so blue = $-10$, white = $0$, and red = $10$.}
 %\textcolor{red}{NEED TO DO FOR FP = 0.05 and FN = 0.3}}
 \label{fig:heatmap}
 \vspace{-0.3cm}
 \end{figure}

 Comparing the mean-squared error (MSE) under a selection mechanism $\I$ with imperfect testing and SRS with perfect testing, we see that
 $$
 \frac{E_{\I} \left[ (\bar y_n - \bar Y)^2 \right]}{\sqrt{V_{SRS} (\bar Y)}} = (N-1) E_{\I} \left[ \rho_{I,Y}^2 D_M^2 \right]
 $$
 implying the error relative to SRS increases as a function of population size~\citep{Meng2018}. The paradox that two countries with the same testing strategy can yield wildly different estimates due to population size is too often ignored in current communications of case-count data.  Thus, conclusions drawn from observed case-counts may not just be wrong, but very wrong.

 A key question is ``What is the (effective) sample size from a SRS with perfect testing that would yield equivalent MSE to the current testing strategy?'' In Appendix~\ref{app:effss}, we show the effective sample size $n_{eff}$ can be bounded by
 $$
 \frac{f}{1-f} \times \frac{1}{E_{\I} \left[ \rho_{I,Y}^2 D_M^2 \right]}.
 $$
 Between April 25th to 29th 2020, Indiana performed $95,879$ tests.  Indiana's population is roughly $6.732$ million, so $f = 0.003$.  The active infection rate was estimated to be~$1.81\%$~\citep{Yiannoutsos2021} in this time interval. Recent studies have suggested 87\% sensitivity~\citep{Arevalo2020} and 97.6\% specificity~\citep{Cohen2020} are reasonable measurement error rates for RT-PCR tests. Supposing COVID-19 positive individuals are $1.5$ times more likely to get tested, then the effective sample size is $168$. Recent proposals~\citep{Siddarth2020} have argued for increased testing capacity, which may likely reduce the relative sampling rate.  Even if the relative sampling rate drops to $1.2$ and $f$ increases to $0.01$ then the effective sample size will increase to $1025$.  Thus the effective sample size even in optimistic scenarios is equivalent to a moderate random sample from the population.  Moreover, increased testing capacity may also come at the cost of increases in false positive and negative rates.  Consider the case where $f$ increasing from $0.003$ to $0.01$ and is associated with the false negative rate rising from $13\%$ to $20$\%.  Then the effective sample size is $863$, representing a $5.1$ factor increase rather than the expected $6.1$ factor increase. See Section~\ref{section:effss} in the Supplementary Materials for additional effective sample size calculations.
 % \begin{wrapfigure}{r}{0.5\textwidth}

 % {\bf Figure of individual going through the states? Can discuss exposure length as selection more likely.}

 \subsection{Regrettable rates: complex biases resulting from self-selection}
 \label{section:rates}

 The prior analysis demonstrates the limited information regarding COVID-19 prevalence in observational case-count data.  Scientists, however, may claim that daily observed case-counts simply undercount daily total cases by a constant multiple over time.  If true then the ratio of case-counts at consecutive times may be a good estimate of the true change in prevalence, helping scientists understand the disease trajectory.  We next demonstrate how selection bias and imperfect testing impact such estimates.

 Let $\bar Y_{t-1}$ and $\bar Y_{t}$ denote the prevalence on two consecutive days and consider the estimator $r = \tilde y_t / \tilde y_{t-1}$.  Using a second-order Taylor series approximation, the error between ${\tilde y_t}/{\tilde y_{t-1}}$ and ${\bar Y_{t}}/{\bar Y_{t-1}}$ can be expressed approximately as
 $$
 \begin{aligned}
 \frac{\bar Y_t}{\bar Y_{t-1}} &\times \bigg[ \rho_{I_t,Y_t} D_{M_t} \sqrt{\frac{1-f_t}{f_t}} CV (Y_t)  -\rho_{I_{t-1},Y_{t-1}} D_{M_{t-1}} \sqrt{\frac{1-f_{t-1}}{f_{t-1}}} CV (Y_{t-1}) \bigg] \\
 &\times \left[ 1 - \rho_{I_{t-1},Y_{t-1}} D_{M_{t-1}} \sqrt{\frac{1-f_{t-1}}{f_{t-1}}} CV (Y_{t-1}) \right]
 \end{aligned}
 $$
 where $\rho_{I_j, Y_j}$ is the data quality, $f_j$ is the sampling fraction, $D_{M_j}$ is the measurement error adjustment, and $CV(Y_j) = \sigma_{Y_j}/ \bar Y_j$ is the coefficient of variation on day $j$.  See Appendix~\ref{app:ratio} for the derivation. The error magnitude depends on the true rate $\bar Y_{t} / \bar Y_{t-1}$ so a large decrease will have a small error relative to a large increase. The second term represents potential \emph{cancellation} which can  occur when data quality, sampling fraction, measurement error, and prevalence are constant across time.

 Figure~\ref{fig:ratio-bias} displays the trajectory of the true ratio and the potential biased estimators under a susceptible-infected-recovered (SIR) model~\citep{Pastor2001,Newman2002,Parshani2010} for the epidemic dynamics, with state evolution given by
 \begin{equation}
 \label{eq:sir}
 \frac{\partial s_t}{\partial t} = - \beta s_t i_t; \quad
 \frac{\partial i_t}{\partial t} = \beta s_t i_t - \gamma i_t; \quad
 \frac{\partial r_t}{\partial t} = - \gamma i_t
 \end{equation}
 where $s_t$ is the fraction of susceptible individuals, $i_t$ is the fraction of infected individuals, and $r_t$ is the fraction of removed (recovered or deceased) individuals in the population at time $t$.  In terms of bias, the rate is overestimated prior to the peak and underestimated afterwards; the bias increases dramatically when the relative fraction exceeds $2$.  Such biases may have a substantial impact on policy making.  First, overestimation may give policy makers more leverage in proposing aggressive actions to reduce prevalence.  Of course, the analysis provides evidence that estimates based on available data may be pessimistic. What is missing in the current discourse is that the direction of bias is non-constant over time.   Underestimation post-peak puts pressure on policy makers to prematurely relax social distancing measures.  Moreover, it appears that the peak time is the easiest to estimate in terms of having minimal error using available data.  Note, however, that the standard errors will be large due to small effective sample sizes.

 \subsection{Estimation of effective reproduction number}
 \label{section:r0-estimation}
 Many epidemiologists argue that tracking the effective reproduction number $R_t$ is the only way to manage through the crisis~\citep{Gabriel2020}.  Under a Poisson likelihood, a simple relation between the trajectory of new cases and the effective reproduction number can be derived \citep{Bettencourt2008}.  In particular, under an SIR model the number of case counts on day $t$, denoted $K_t$, is Poisson distributed with rate $K_{t-1} \exp \left( \gamma (R_t - 1) \right)$ where $K_{t-1} = Y_{t-1}-Y_{t-2}$ is the number of new cases on day $t-1$ and $\gamma$ is the serial interval, which is approximately $7$ days for COVID-19~\citep{Sanche2020}.  Using this relation, a moment-based estimator is given by
 $$
 R_t = 1 + \frac{1}{\gamma} \log \left( \frac{K_t}{K_{t-1}} \right).
 $$
 Of course, we do not observe $K_t$ and $K_{t-1}$.  Under SRS of new cases among those susceptible on day $t$, the natural estimator is $S_t \tilde y_t$.  Unfortunately the number of susceptible individuals on day $t$ is unknown. Here, we study the estimator $\hat R_t = 1 + \frac{1}{\gamma} \log \left( \tilde y_t / \tilde y_{t-1} \right)$. We can again express the statistical error of $\hat R_t - R_t$ in useful terms as follows
 $$
 \begin{aligned}
 \frac{1}{\gamma}\log &\bigg( 1 + \bigg[ \rho_{I_t,K_t} D_{M_t} \sqrt{\frac{1-f_t}{f_t}} CV (K_t)  -\rho_{I_{t-1},K_{t-1}} D_{M_{t-1}} \sqrt{\frac{1-f_{t-1}}{f_{t-1}}} CV (K_{t-1}) \bigg] \\
 &\times \left[ 1 - \rho_{I_{t-1},K_{t-1}} D_{M_{t-1}} \sqrt{\frac{1-f_{t-1}}{f_{t-1}}} CV (K_{t-1}) \right] \bigg) - \frac{1}{\gamma} \log \left( \frac{S_t}{S_{t-1}} \right).
 \end{aligned}
 $$
 This implies a similar trade-off as before but on the logarithmic scale.  The error is no longer scaled by $\bar Y_t/\bar Y_{t-1}$ but by the serial interval and does not depend on prevalence but on the fraction of new cases out of those susceptible. This leads to differences in when the bias is most pronounced. Figure~\ref{fig:r0-bias} displays the bias as a function of the relative sampling fraction assuming the population fraction that is susceptible remains large (i.e., $S_t / S_{t-1} \approx 1$).

 \begin{figure}
 \centering
 \begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{../figs/sir_ratio.png}
  \caption{Ratio estimator}
  \label{fig:ratio-bias}
 \end{subfigure}%
 \begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{../figs/sir_rt.png}
  \caption{Effective reproductive rate estimator}
  \label{fig:r0-bias}
 \end{subfigure}
 \caption{Potential bias in ratio and effective reproductive rate estimators under an SIR model with $\beta = 1.2$ (black) and $\gamma = 0.15$.  Here, $f = 0.02$, $FP = 0.024$, $FN = 0.13$, and a range of relative sampling fractions $M = f_1/f_0$ are considered.}
 \label{fig:rates}
 \end{figure}

 \subsection{Rate comparisons}

 So far we have focused on understanding the limitations of using case-count data to understand population quantities of interest for a \emph{single} population.  Many are interested in cross-population comparisons to contrast the impact of countries' mitigation policies.  Here, for simplicity, we focus on comparing the estimated effective reproductive rate.  We assume the two time-series are aligned so that $t=0$ is the time of first case in each population respectively.
 % This negates alignment issues and is common in practice.

 % While the above issues on $Z$-scores and effective sample size are important, here we highlight a separate issue.
 %Here we demonstrate the interaction of biases in estimation of the two trajectories when the peak infection times differ slightly.
 %First, we can write the difference in the rates as $R_{t1} - R_{t2} = \frac{1}{\gamma} \log \left( \frac{1 + e_{1t}}{1+e_{2t}} \right)$ where $e_{jt}$ is the error associated given in Section \ref{section:r0-estimation}.
 %Recall that the error allowed for over-estimation prior to the peak and then under-estimation post-peak.  Here, these errors can mingle in interesting ways.
 Consider two countries (A and B) in which the peak occurs 2 weeks prior for country A than country B.  Figure~\ref{fig:comparison} presents such a comparison where each country's disease trajectory follows an SIR model (A=black and B=grey). Figure~\ref{fig:eff} shows how biases interact in complex ways.  At first, the difference is correctly estimated; then the gap is over-estimated as country A sees a rapid rise in cases; then the magnitude of over-estimation increases as country A sees declining case-count while country B sees rapidly increasing case-count; then country A's rate is correctly estimated while country B's rate is under-estimated as it sees declining case-count; finally, the gap disappears.  While this may not always be the case, the analysis demonstrates how estimates can tell a more complex story than the truth (i.e., country A's peak is 2 weeks prior to country B's peak).

 \begin{figure}
 \centering
 \begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{../figs/sir.png}
  \caption{Fraction of new cases in population}
  \label{fig:fracpop}
 \end{subfigure}%
 \begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{../figs/sir_rt_comparison.png}
  \caption{Effective reproduction rate estimators}
  \label{fig:eff}
 \end{subfigure}
 \caption{Left: fraction infected in two SIR models with $\beta = 1.2$ and $0.9$ respectively and $\gamma = 0.15$ with same initial conditions. Right: comparison of $\hat R_t$ across time with $FN = 0.30$, $FP = 0.024$, and $M = 4$.}
 \label{fig:comparison}
 \end{figure}

 \section{Potential improvements to prevalence estimation}
 \label{section:improvedcasecount}

 The prior section presented negative consequences of selection-bias and measurement error when estimating infection prevalence, rates of change, and the effective reproduction number from observed case-count data.  This raises the immediate question whether we can can improve upon these estimators.  In this section, we consider two directions.
 The proposed methods and statistical error decompositions guide our recommendations in Section~\ref{section:discussion}.

 \subsection{Selection propensity estimation}

 With non-probability samples, one can reduce bias by modelling the self-selection propensity and use inverse probability weighting (IPW)~\citep{Elliott2017} techniques to adjust for selection bias.  In causal inference~\citep{Hernan2020}, under certain assumptions, the observational data permits direct estimation of treatment propensity.  In the COVID-19 context, one does not observe those who decide not to be tested.  In order to estimate selection propensities, one needs auxiliary information.

 To see this mathematically, let $X_j$ denote a vector of covariates for the $j$th individual in the population, $j=1,\ldots,N$.  For simplicity, the dynamic nature of the outbreak and recoverability of individuals is ignored for now. Suppose the selection indicator $I_j$ is a Bernoulli random variable that depends only on these covariates, i.e., $P(I_j = 1 \mid X_j = x) = \pi (x; \theta)$. Here, we focus on logistic regression models, i.e., $\pi(x; \theta) = \text{expit} \left( x^\top \theta \right)$. Maximum likelihood estimation follows by maximizing
 \begin{equation}
 \label{eq:propensity}
 \sum_{j=1}^N I_j \log \left( \frac{\pi (X_j; \theta)}{1-\pi(X_j; \theta)} \right) + \sum_{j=1}^N \log \left( 1 - \pi (X_j; \theta) \right).
 \end{equation}
 The first sum only involves individuals observed in the nonprobability sample.
 The second sum is over the entire population.  Maximum likelihood estimation therefore requires knowledge of covariate information for every individual in the population.  Typically, this is not possible.  Here we present a method that uses auxiliary information obtained from probability samples.

 \subsubsection{Auxiliary information through probability samples}
 \label{subsec:auxprob}

 Here we assume access to a probability sample measuring the same set of covariates.  Let $\tilde I_j$ be an indicator that the individual was included in the probability sample and $\tilde W_j$ be the probability of inclusion for the $j$th individual (i.e., equal to $n/N$ for a simple random sample of size $n$). In recent work,~\cite{Chen2019} use a probability sample to construct a design-unbiased estimator of the second term
 \begin{equation}
 \label{eq:auxinfoprob}
 \sum_{j=1}^N I_j \log \left( \frac{\pi (X_j; \theta)}{1-\pi(X_j; \theta)} \right)  + \sum_{i=1}^N \tilde I_j \tilde W_j \log ( 1 - \pi (X_j; \theta)).
 \end{equation}
 Taking an expectation with respect to the sampling design yields~\eqref{eq:propensity}. Solving~\eqref{eq:auxinfoprob} can be done by iteratively re-weighted least squares (IRLS); see Section~\ref{app:irls} in the supplementary materials for algorithmic details.

 % Under simple random sampling and non-parametric selection propensity on a stratification variable, the weight is given by estimator of the $\hat \pi_k = (d_k / \tilde d_k) (n / N)$ where $\tilde d_k$ is the number of observations in strata $k$ in the random sample.

 \subsubsection{An IPW estimator and statistical error decomposition}
 \label{section:IPWerrordecomp}

 Given a selection propensity, define the inverse probability weight~$w(x) = \pi (x; \hat \theta)^{-1}$. Then the IPW estimator adjusted for measurement error is given by
 \begin{equation}
 \label{eq:ipwest}
 \bar y_n^\star
 = \frac{1}{1-FP-FN} \cdot \frac{\sum_{i=1}^n w(x_i) (y_i - FP)}{\sum_{i=1}^n w (x_i)}
 \stackrel{(2)}{=} \frac{1}{1-FP-FN} \sum_{k=1}^K \frac{d_k w_k}{w} (\bar y_k - FP),
 \end{equation}
 where equality (2) is under the assumption that $x_i$ is a stratification variable with $k$ indexing the strata, $w_k$ is the weight and $d_k$ is the number of samples in strata $k$, and $w = \sum_{k=1}^K d_k w_k$.
 % Under simple random sampling with one strata~$\hat \pi_k = n/N$ and we recover the estimator $\tilde y_n$ from Section~\ref{section:imperfecttesting}.

 Let $\tilde I_j (X_j) = I_j  \cdot W(X_j)$ for $j=1,\ldots,N$.  Then the error when comparing weighted estimator~$\bar y_n^\star$ to the true prevalence $\bar Y$ can be expressed as:
 \begin{equation}
 \label{eq:statdecomp2}
 \rho_{\tilde I (X), Y} \times \sqrt{\frac{1-f+ CV^2_W}{f}} \times \sigma_{Y} \times \underbrace{\left[ 1 - \tilde \Delta \times \frac{\bar Y}{1-\bar Y} \times \frac{FP(1-\bar Y) + FN \cdot \bar Y}{\tilde f_0 (1-\bar Y) + \tilde f_1 \bar Y} \right] \times \frac{1}{1-(FP+FN)}}_{\tilde D_M}
 \end{equation}
 where $CV_W$ is the coefficient of variation (i.e., standard deviation/mean) of $W_J$ given $I_J = 1$, $\rho_{\tilde I(X), Y}$ is the empirical correlation which here depends on covariate distribution, $\tilde f_k = E[ W_J I_J \mid Y_j = k]$ for $k=0,1$, and $\tilde \Delta = \tilde f_1 - \tilde f_0$.  See Appendix~\ref{app:ipwderivation} for the derivation.

 Comparing~\eqref{eq:statdecomp2} to~\eqref{eq:statdecomp} shows that weighting impacts the estimation error in three ways.  First, there is a negative impact on the data quantity component; taking the ratio of these quantities yields
 $\sqrt{1 + \frac{CV_W^2}{1-f}} \geq 1$.  Hence, if the data quality does not increase (i.e., $| \rho_{\tilde I (X), Y} | = | \rho_{I,Y}|$ ) then weighting increases the error magnitude.  Second, the relative error increase depends on the fraction of population sampled $f$, implying that for large samples there is a larger potential increase in the error if the weights do not improve data quality. Third, the impact of measurement-error on data quality is changed when considering a weighted estimand. In particular, weighting may result in $\text{sgn}(\Delta) \neq \text{sgn} (\tilde \Delta)$ which implies the impact of measurement-error may be in a different direction.

 As demonstrated below in Lemma~\ref{lemma:ipw}, if the propensity model is correctly specified then the $E_{\bf R} [ \rho_{\tilde I (X), Y} ] = 0$ and therefore $\rho_{\tilde I(X), Y} = O(N^{-1})$; however, if the weights are not correctly specified then the data quality index is unlikely to inversely scale with population size.
 % Indeed, Horvitz-Thompson estimators are sensitive to errors in the estimated weights.  Low selection propensity imply large weights which can have deleterious impacts by certain observations dominating estimation.
 Similar to~\cite{Meng2018}, if the data quality is not at the level of $N^{-1}$, then confidence intervals constructed from an IPW estimator are likely to put too much confidence in the sheer data size.

 \subsubsection{Time-varying propensities}

 Here we extend the IPW approach to the temporal setting to account for the dynamic nature of the outbreak by considering the joint likelihood
 \begin{equation}
 \label{eq:tvpropensity}
 \sum_{t=1}^T \left[ \sum_{j=1}^N I_{j,t} \log \left( \frac{\pi_t (X_{j,t}; \theta)}{1-\pi_t(X_{j,t}; \theta)} \right) + \sum_{j=1}^N \log \left( 1 - \pi_t (X_{j,t}; \theta) \right) \right]
 \end{equation}
 where $t=1,\ldots,T$ are the days when case-count data is reported.  Here, $I_{j,t}$ denotes self-selection into testing on day $t$, which is highly correlated with prior testing and results.  For example, an individual who tests positive may be unlikely to seek testing in the subsequent few days/weeks.  Moreover, an individual in a high prevalence area may be more likely to seek out testing.  Here, we assume that the covariate vector $X_{j,t}$ contains all features of the past relevant for selection.

 If sufficiently large random samples are collected at each time $t =1,\ldots,T$, then the pseudo-likelihood can be re-written as in~\eqref{eq:auxinfoprob} and propensities estimated separately per time point.  Unfortunately, large probabilistic samples are not available at every time within a given region.  To address this, here we consider a non-parametric kernel-based approach where the selection propensity at time~$t$, denoted $\hat \theta_t$, maximizes the smoothed pseudo-likelihood
 $$
 \sum_{t^\prime=1}^T K_h(|t^\prime - t|) \left[ \sum_{j=1}^N I_{j,t^\prime} \log \left( \frac{\pi_t (X_{j,t^\prime}; \theta)}{1-\pi_t(X_{j,t^\prime}; \theta)} \right) + \sum_{j=1}^N \tilde W_{j,t^\prime} \tilde I_{j,t^\prime} \log \left( 1 - \pi_t (X_{j,t^\prime}; \theta) \right) \right]
 $$
 where $K_h$ is a kernel function with tunable parameter $h$. Given $\pi (x;\hat \theta_t)$, the prevalence estimator $\bar y_{n,t}^\star$ is given by~\ref{eq:ipwest} using case-count data observed on day~$t$.

 \subsubsection{Asymptotics}
 \label{section:asymptotics}

 Suppose there is a sequence of finite populations of size $N_{\nu}$ indexed by $\nu$.  For each finite population, we assume there are $L_\nu$ non-probability samples and probability samples of size $n$ drawn at equally spaced times $\{ t^\prime_l \}_{l=1}^{L_{\nu}}$ over the study window~$[0,T]$. Assuming the selection propensity model is correctly specified, then under the design of the probability sample and design of the subsample from the nonprobability sample, Lemma~\ref{lemma:ipw} shows that the IPW estimator is consistent as $\nu$ goes to infinity (i.e., $N_{\nu}$ and $L_{\nu} \to \infty$) and calculates the estimator's variance.  In Lemma~\ref{lemma:ipw}, sensitivity and specificity are not assumed known.  Instead, we suppose that these are estimated using a pseudo-likelihood on two random samples from the population whose size (denoted~$n_{FP}$ and $n_{FN}$ respectively) is chosen so the resulting estimated sensitivity and specificity have desired levels of uncertainty.  See Appendix~\ref{app:asympderivations} for additional details.

 \begin{lemma}[Variance of IPW estimator] \normalfont
 \label{lemma:ipw}
 % \textcolor{red}{Must check.}
 The estimates~$\hat \mu_t := \bar y_{n,t}^\star$, $\hat \pi_{j,t}$, $\hat FP$, and $\hat FN$ are solutions to the following set of estimating equations
 $$
 \Phi_n (\eta_t) =
 \left (
 \begin{array}{c}
 \frac{1}{N} \sum_{j=1}^N I_{j,t} \frac{Y_{j,t} - FP - (1-FP-FN) \cdot \mu_t}{\pi_{j,t}} \\
 \frac{1}{N \times \sum_{l = 1}^L K_h \left(|t - t_l^\prime| \right)} \sum_{l = 1}^L K_h \left(|t - t_l^\prime| \right) \left[ \sum_{j=1}^N I_{j,t_l^\prime} X_{j,t_l^\prime} - \sum_{j = 1}^N \tilde I_{j,t_l^\prime} \tilde W_{j,t_l^\prime}  \pi_{j,t_l^\prime} X_{j,t_l^\prime}  \right] \\
 \frac{1}{n_{FP}} \sum_{i=1}^{n_{FP}} \frac{Z_j}{FP} - \frac{1-Z_j}{1-FP} \\
 \frac{1}{n_{FN}} \sum_{i=1}^{n_{FN}} \frac{\tilde Z_j}{FN} - \frac{1-\tilde Z_j}{1-FN} \\
 \end{array}
 \right ) = {\bf 0}
 $$
 where $\eta_t = (\mu_t, \pi_t, FP, FN)$. Under certain regularity assumptions (see Appendix~\ref{app:asympderivations}), we have $\bar y_{n,t}^\star - \bar Y_{t} = O_p (\bar n^{-1/2})$  and $\text{var} (\hat y_{n,t}) = V_{t}^{(IPW)} + o (\bar n^{-1})$ where $\bar n = n \sum_{l=1}^L K_{h} (|t - t_l^\prime|)$ and $V_t^{(IPW)}$ is the first diagonal element of $E [\phi_n(\eta_0)]^{-1} \text{Var}(\phi_n(\eta_0))E [\phi_n(\eta_0)]^{-1}$ where $\phi_n = \frac{\partial \Phi (\eta)}{\partial \eta}$ with $E[\cdot]$ and $\text{Var} (\cdot)$ are under the joint randomization of propensity score and sampling designs.
 \end{lemma}

 \subsection{Model-based estimation}
 \label{section:modelbased}

 Up to this point, the primary focus has been selection bias in coronavirus case-count data from a survey sampling perspective.  Here, we consider compartmental model approaches from infectious disease epidemiology.  Our primary objective is a model-based forecast of strata-level prevalence $\bar y_k$. To do this, a probabilistic extension of a standard epidemiological state-space model -- the susceptible, exposed, infected, and removed (recovered and death) model, or SEIR model -- is presented. A probabilistic SIR model was originally proposed by~\cite{Osthus2017} with only one-dimensional time series of infected proportions; this formulation was extended by~\cite{Song2020} to model coronavirus case-counts.

 Let $s_t$, $e_t$, $i_t$, and $r_t$ denote the proportion of survivors, exposed, infected, and removed cases (i.e., including both recovered cases and deaths) at time $t$.  The population-level SEIR dynamics are then given by the following set of differential equations:
 \begin{align}
 \frac{\partial s_t}{\partial t} &= - \beta s_t i_t; \quad \frac{\partial e_t}{\partial t} = \beta s_t i_t - \sigma e_t; \quad \label{eq:seir} \\
 \frac{\partial i_t}{\partial t} &= \sigma e_t - \gamma i_t; \quad
 \frac{\partial r_t}{\partial t} = - \gamma i_t. \nonumber
 \end{align}
 Following~\cite{Song2020}, we employ Runge-Kutta (RK4) approximations for discretization. Here, we consider covariate information that takes the form of a stratification variable with $K$ strata.  Let $I_{k,t}$ to denote the number of new infections on day $t$ in the $k$th strata.  As there are a total of $N \cdot i_t := I_t$ new infections on day $t$, we assume the joint distribution of strata-specific new infection totals takes a multinomial distribution with
 $$
 ( I_{1,t}, \ldots, I_{K,t} ) \sim \text{Multinomial} \left( I_t, (p_{1,t}, \ldots, p_{K, t}) \right).
 $$
 where $( {\bf p}_t )_{t=1}^T = ( (p_{1,t}, \ldots, p_{K, t}) )_{t=1}^T$ are a sequence of parameters on the simplex, i.e., $\sum_{k=1}^K p_{k,t} = 1$.

 Selection bias is addressed by analyzing COVID-19 death data rather than case counts.  Let $D_{k,r}$ denote the number of individuals who pass away on day $r$ from strata $k$.  Then
 $$
 D_{k,r} \mid p, \theta, \nu \sim \text{Poisson} \left( \sum_{t=1}^r \text{IFR}_{k} \cdot I_{k,t} \theta_{(r-t)} \right)
 $$
 where $\text{IFR}_{k}$ is the infection fatality rate for the $k$th strata and $\theta_{j}$ is a discrete-time distribution for time from infection to death.  This extends prior analysis of death data~\cite{Johndrow2020} by allowing the infection fatality rate to depend on a stratification variable, which is important as IFR depends heavily on age~\citep{Levin2020}. See Section~\ref{section:application_of_modelbased} for discussion of parameter choices.  %Additional details on inference are discussed in Appendix~\ref{app:modelbased}.

 % 0-17: 20/1000000 * 100 -> 0.002 (at age 9)
 % 18-49: 500/1000000 * 100 -> 0.05 (at age 33)
 % 50–64 years old: 6,000/1000000 * 100 -> 0.6
 % 65+ years old: 90,000/1000000 * 100 -> 9.0
 %

 \subsection{Doubly robust estimation.}
 Rather than relying solely on IPW or epidemiological forecasts, here we combine forecasting and inverse-probability weighting by extending recent work by~\cite{Chen2019} to account time-varying propensities and measurement-error. In this context, the doubly-robust estimator is given by
 $$
 \bar y_{n}^{(DR)} = \frac{1}{N} \sum_{j=1}^N \hat \mu (x_j) + \frac{1}{\sum_{j=1}^N I_j w (x_j)} \sum_{j=1}^N I_j w(x_j) \left( \frac{Y_j - FP}{1 - FP - FN} - \hat \mu(x_j) \right).
 $$
 where the estimates $\hat \mu(x_j)$ are not corrected for measurement-error as they are built to estimate true case count from death-only data. This estimator is called ``doubly-robust'' because it is consistent if either the model-based forecasts or the time-varying propensities are correctly specified. A statistical error decomposition can be derived
 \begin{align}
 \label{eq:statdecomp3}
 \begin{split}
 \rho_{\tilde I (X), Y-\mu(X)} &\times \sqrt{\frac{1-f+ CV^2_W}{f}} \times \sigma_{Y-\mu(X)}  \\
 &\times \underbrace{\left[ 1 - \frac{\rho_{\tilde I(X),Y} \sigma_{Y}}{\rho_{\tilde I(X),Y-\mu(X)} \sigma_{Y-\mu(X)}} \times \tilde \Delta \times \frac{\bar Y}{1-\bar Y} \times \frac{FP(1-\bar Y) + FN \cdot \bar Y}{\tilde f_0 (1-\bar Y) + \tilde f_1 \bar Y} \right]}_{\tilde D_M}.
 \end{split}
 \end{align}
 See Appendix~\ref{app:DRderivation} for the derivation.
 % Comparing~\eqref{eq:statdecomp3} to~\eqref{eq:statdecomp2} shows that the addition of the model-based forecast impacts the estimation error in two ways.
 If the model is adequate, then one may expect a reduction in the problem difficulty and (potentially) in the data quality components.
 % Note that the model-based forecast does not impact the data quantity component.
% Second, if the model-based forecast is correct then $E_{\bf R} [ \rho_{\tilde I (X), Y-\mu(X)} ] = 0$  and therefore $\rho_{\tilde I(X), Y-\mu(X)} = O(N^{-1})$; however, if the forecasts are not correctly specified then the data quality index is unlikely to inversely scale with population size.
 Interestingly, the impact of measurement-error on data quality now depends on a relative comparison of the data quality and problem difficulty of the weighted estimator and the doubly-robust estimator.
 % For decent models, one would expect this ratio to be greater than one.  Therefore, $\tilde \Delta < 0$ implies an expected increase in impact of measurement-error and $\tilde \Delta > 0$ implies the magnitude could shrink toward zero or change sign and have an even large magnitude in the opposite direction.  Again, if the data quality is not at the level of $N^{-1}$, then confidence intervals constructed from a DR estimator are likely to put too much confidence in the sheer size of the data.

 Unfortunately, the first term of the doubly robust estimator cannot be computed as covariate information is not collected on the entire population.  Here, we suppose the same asymptotic regime as in Section~\ref{section:asymptotics} and use the probability samples and consider the following estimator at time~$t$
 \begin{align*}
 &\frac{1}{\sum_{l=1}^L K_{h} (|t - t_l^\prime|) \sum_{j=1}^N \tilde I_{j,t_l^\prime} \tilde W_{j,t_l^\prime}} \sum_{l=1}^L K_{h} (|t - t_l^\prime|) \sum_{j=1}^N  \tilde I_{j,t_l^\prime} \tilde W_{j,t_l^\prime} \hat \mu_{t^\prime_l} (X_{j,t^\prime}) \\
 + &\quad \frac{1}{\sum_{j=1}^N I_{j,t} w (X_{j,t})} \sum_{j=1}^N I_{j,t} w(X_{j,t}) \left( \frac{Y_{j,t} - FP}{1 - FP - FN} - \mu_t(X_{j,t}) \right).
 \end{align*}
 where $\tilde I_{j,t_l^\prime}$ indicates an individual is included in the random sample at time~$t_l^\prime$ and $\tilde W_{j,t_l^\prime}$ is the associated weight for that unit. Assuming the selection propensity model is correctly specified, then under the design of the probability sample and design of the subsample from the nonprobability sample, Lemma~\ref{lemma:dr} shows that the doubly robust estimator is consistent and calculates the estimator's variance. See Appendix~\ref{app:asympderivations} for additional details.

 \begin{lemma}[Variance of doubly-robust estimator] \normalfont
 \label{lemma:dr}
 The estimates~$\mu_t := \bar y_{n,t}^\star$ and $\hat \pi_{j,t}$
 are solutions to the following set of estimating equations
 $$
 \Phi_n (\eta_t) =
 \left (
 \begin{array}{c}
 \frac{1}{N} \left[ \sum_{j=1}^N \frac{I_{j,t}}{\pi_{j,t}} \left( \left( \frac{Y_{j,t} - FP}{1-FP-FN} - \hat \mu_{j,t}  \right)  - \mu_t \right) + \frac{1}{\sum_{j=1}^N \tilde I_j \tilde W_j} \sum_{j=1}^N  \tilde I_j \tilde W_j \hat \mu_t (X_{j,t}) \right] \\
 \frac{1}{N \times \sum_{l=1}^L K_{h} (|t - t_l^\prime|)} \sum_{l=1}^L K_{h} (|t - t_l^\prime|) \sum_{j=1}^N I_{j,t^\prime} X_{j,t^\prime} - \frac{1}{N} \sum_{j = 1}^N \tilde I_{j,t^\prime} \tilde W_{j,t^\prime}  \pi_{j,t^\prime} X_{j,t^\prime}  \\
 \frac{1}{n_{FP}} \sum_{j=1}^{n_{FP}} \frac{Z_j}{FP} - \frac{1-Z_j}{1-FP} \\
 \frac{1}{n_{FN}} \sum_{j=1}^{n_{FN}} \frac{\tilde Z_j}{FN} - \frac{1-\tilde Z_j}{1-FN} \\
 \end{array}
 \right ) = {\bf 0}
 $$
 where $\eta_t = (\mu_t, \pi_t, FP, FN)$. Under regularity assumptions (see Appendix~\ref{app:asympderivations}), we have $\bar y_{n,t}^\star - \bar Y_{t} = O_p (\bar n^{-1/2})$ and $\text{var} (\hat y_{n,t}) = V_{t}^{(DR)} + o (\bar n^{-1})$ where $\bar n = n \sum_{l=1}^L K_h(|t-t_l|)$ and $V_t^{(DR)}$ is the first diagonal element of $E [\phi_n(\eta_0)]^{-1} \text{Var}(\Phi_n(\eta_0))E [\phi_n(\eta_0)]^{-1}$ where $\phi_n = \frac{\partial \Phi (\eta)}{\partial \eta}$ with $E[\cdot]$ and $\text{Var} (\cdot)$ are under the joint randomization of propensity score and sampling designs.
 \end{lemma}

 \section{COVID-19 active infection prevalence in Indiana}
 \label{section:applications}

 % For COVID-19 case-count data, the only covariate information provided by state and local governments are summary statistics for those factors discussed in Section~\ref{section:testinginfo} -- age, sex,  and race.  These are discrete factors and therefore define a stratification of the population.  Currently, US testing does not collect additional covariate information such as symptom information; however, this need not be the case moving forward.  Recommendation XX below states that additional covariate information should be collected from nonprobabilistic samples.

 We next consider estimation of the active infection rate in Indiana using unweighted, IPW, model-based, and doubly robust estimates. To start, we recap the data sources used in estimation:
 \begin{itemize}[leftmargin=*]
 \item {\bf Testing data}:  The number of daily tests performed and number of daily positive tests are reported.  These counts are broken out jointly by age, gender, and racial demographic information.  Figure~\ref{fig:in-cases} and~\ref{fig:in-tests} plots daily COVID-19 positive cases and tests by age strata.

 \begin{figure}
 \centering
 \begin{subfigure}{.45\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{../figs/indianacasecounts_byage.png}
  \caption{Case-counts per day}
  \label{fig:in-cases}
 \end{subfigure}
 \begin{subfigure}{.45\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{../figs/indianacovidtests_byage.png}
 \caption{Tests per day}
 \label{fig:in-tests}
 \end{subfigure} \\ [1ex]
 \begin{subfigure}{\linewidth}
 \centering
 \includegraphics[width=.45\linewidth]{../figs/indianadeaths_byage.png}
  \caption{Deaths per day}
  \label{fig:in-deaths}
 \end{subfigure}
 \caption{Indiana daily COVID-19 case count, testing, and death data by age strata}
 \label{fig:in-data}
 \end{figure}

 \item {\bf Random/Nonrandom statewide sample}: Table~\ref{tab:indiana} summarizes data from a random sample from April 25--29th as well as a nonrandom sample obtained between May 2--3 in racial/ethnic minority communities~\cite{Yiannoutsos2021}. The nonrandom sample is not a random subsample of the overall case-counts; however, this data provides important supplementary covariate information.
 % Namely, the samples collected included covariate information on (1) prior symptoms, (2) COVID-19 positive individual in household, and (3) prior COVID-19 positive test.
 To account for the nonrandom sample being from high risk areas, we adjust estimates to match the statewide positivity rate of 11.7\% on August 30th. These adjusted rates are presented in parentheses in Table~\ref{tab:indiana} along with the relevant covariate information.

 \item {\bf Facebook symptom survey}: From the daily symptom surveys completed on Facebook (see Section~\ref{subsection:fbsymptom} for details), we extract survey responses for individuals who identify as living in Indiana.  We collect age, sex, and demographic information as well as COVID-19 related symptoms, which Facebook defines as having (1) a fever and cough, (2) shortness of breath, \emph{or} (3) difficulty breathing. To assess potential bias, we compare the random sample from Indiana to the Facebook symptom survey data collected from April 25th to April 29th.  Table~\ref{tab:comparison} demonstrates minimal bias in the Facebook symptom survey with respect to symptom distributions.

 \begin{table}[!th]
 \centering
 \begin{tabular}{c | c c c | c}
 & \multicolumn{3}{c}{Facebook} & Random \\ \cline{2-4}
 & Lower CI & Estimate & Upper CI & Estimate \\ \hline
 Fever &  0.017 & 0.020 & 0.024 & 0.018 \\
 Cough &  0.142 & 0.148 & 0.154 & 0.149 \\
 Shortness & 0.058 & 0.060 & 0.062 & 0.062 \\ \hline
 \end{tabular}
 \caption{Comparison of Facebook survey and random sample}
 \label{tab:comparison}
 \end{table}

 % \item {\bf Hospitalization data}: On each day, we observe the number of individuals who are COVID-19 positive that visited a hospital on that day.  Unfortunately, hospitalization data is not broken out by age, gender, and racial demographic information.  Instead, we rely on combining Facebook data to calculate the fraction of those who sought the hospital per week in each strata. We then use this to figure out how many individuals per strata were hospitalized and then back out how many

 \item {\bf Death data}: Daily COVID-19 related deaths are observed by age, gender, and racial demographics.  Figure~\ref{fig:in-deaths} plots the cumulative number of COVID-19 related deaths over time by age group. While most COVID-19 tests and positive cases are within the younger age strata, most deaths are within the 70+ age strata.
 \end{itemize}

 \begin{table}[th]
 \begin{tabular}{c | c | r r | r | r r }
 & & \multicolumn{2}{c}{Total Tests (\%)} & US & \multicolumn{2}{c}{Positive Test Rate (\%)}\\
 \cline{3-4} \cline{6-7}
 & & NonRandom & Random & Census & NonRandom & Random \\ \hline
 \multirow{2}{*}{Sex} & Female & 58.2 & 55.0 & 50.7 & 21.7 (11.2) & 1.4 \\
 & Male & 41.8 & 45.0 & 49.3 & 24.2 (12.4) & 2.1 \\ \hline
 \multirow{3}{*}{Age} & $<40$ & 39.4 & 28.0 & 52.7 & 29.7 (15.0) & 1.7 \\
 & $40-59$ & 41.1 & 36.0 & 25.2 & 24.9 (12.5) & 2.1 \\
 & $\geq 60$ & 19.5 & 36.0 & 22.1 & 6.7 (3.4) & 0.9 \\ \hline
 \multirow{2}{*}{Race} & White & 23.1 & 92.0 & 86.9 & 19.5 (9.6) & 1.5 \\
 & Nonwhite & 76.9 & 8.0 & 13.1 & 25.0 (12.3) & 3.4 \\ \hline
 % Hispanic & Hispanic & 44.1 & 2.0 & 7.1 & 37.6 & 6.9 \\
 % Origin & Non-Hispanic & 55.9 & 98.0 & 92.9 & 13.0 & 1.3 \\ \hline
 \multirow{2}{*}{Fever} & Yes & 17.0 & 1.8 & - & 66.4 (32.1) & 4.5 \\
 & No & 83.0 & 98.2 & - & 15.6 (7.5) & 1.3 \\ \hline
 Household & Yes & 10.8 & 1.4 & - & 46.1 (22.4) & 29.4 \\
 $+$ Case & No & 89.2 & 98.6 & - & 21.6 (10.4) & 1.3 \\ \hline
 Prior $+$ & Yes & 6.1 & 1.4 & - & 39.2 (20.2) & 24.4 \\
 Test & No & 93.9 & 98.6 & - & 21.6 (11.1) & 1.3 \\ \hline
 \end{tabular}
 \caption{Estimated point prevalence of active infection with SARS-CoV-2 by demographics in Indiana~\citep{Yiannoutsos2021}.}
 \label{tab:indiana}
 \end{table}

 \subsection{Inverse-probability weighting approach}
 \label{section:ipwapproach}

 % FP: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7307014/
 % FN: https://www.medrxiv.org/content/10.1101/2020.04.16.20066787v2.full.pdf

 We start by using Table~\ref{tab:indiana} to compute IPW weights for end of April, early May.  The random sample is $n=3,658$ and IPW weights are computed per strata which depend on Gender, Age, Race, Fever, Positive Case in Household, and Prior Positive Test.  The weight for the strata defined as Male, 40--59, White, has a Fever, no household cases, and no prior tests, for example, is proportional to:
 $$
 \frac{\left( 0.493 \times 0.252 \times 0.869 \times 0.018 \times 0.986 \times 0.986 \right)}{\left(0.418 \times 0.411 \times 0.231 \times 0.17 \times 0.892 \times 0.939 \right)} \approx 0.334
 $$
 Using the constructed weights and the adjusted strata-level prevalence estimates, the IPW estimate under no measurement-error is $7.7\%$, a decrease of four percentage points from the observed prevalence of $11.7\%$. We also make use of Facebook's symptom survey over that window of time to construct a second IPW estimate conditional on demographics and fever but ignoring the household and prior testing information, which under no measurement-error is $6.5$\%.

 The Indiana study did not report sensitivity and specificity; therefore, we take the suggested measures from~\cite{Arevalo2020} which report 87\% sensitivity is a reasonable estimates and~\cite{Cohen2020} which report 97.6\% specificity.  This corresponds to a false negative rate of $13$\% and false positive rate of $2.4\%$.  Therefore, the IPW estimates accounting for measurement-error are $6.2\%$ and $4.9\%$ respectively.

 \subsubsection{Disease prevalence by April 2020}

 Indiana's population as of 2019 was $6.732$ million.  A total of $19,649$ tests were administered in Indiana between April 25th to 29th. Subtracting off the $95,879$ tests that had already performed yields a sampling fraction of $f = 2.96 \times 10^{-3}$.  Here we consider estimation of data quality $E_{\I} \left[ \rho_{\tilde I(X),Y} \right]$. In \cite{Meng2018}, estimation relied on observing the true outcome (i.e., election vote totals), which is not possible in the current crisis.  Here, we use the random sample which estimates the true prevalence of active COVID-19 disease at 1.81\%~\cite{Yiannoutsos2021}.  Our goal is not inferential. Instead, we aim to build sensitivity analyses that can aid in understanding the amount of information in observational COVID case-count data.

 The unweighted estimate for the true prevalence of active COVID-19 disease between April 25th and April 29th is 11.7\%.  Assuming a false negative rate of $13$\% and false positive rate of $2.4$\%, the unweighted estimate is $11.0$\%, leading to an error of $9.2$\%.  Using~\eqref{eq:statdecomp}, we construct an estimate of the relative sampling rate as follows
 \begin{equation*}
 \rho D_M = \sqrt{\frac{f}{1-f}} \frac{\text{0.092}}{\sigma_Y} = 3.75 \times 10^{-2} \Rightarrow \Delta = 1.39 \times 10^{-2} \Rightarrow M = 6.1.
 \end{equation*}
 Using inverse-probability weights, the IPW estimates for the true prevalence of active COVID-19 disease between April 25th and April 29th are $7.7$\% and $6.5\%$ respectively.  Under the same sensitivity and specificity, the weighted estimates adjusted for measurement-error are $6.2$\% and $4.9$\%, leading to errors $4.4$\% and $3.1$\% respectively.  Using~\eqref{eq:statdecomp2}, we construct an estimate of the relative sampling rate as follows
 \begin{equation*}
 \tilde \rho D_M = \sqrt{\frac{f}{1-f+CV_W^2}} \frac{\text{0.044}}{\sigma_Y} = 8.14 \times 10^{-3} \Rightarrow \tilde \Delta = 8.22 \times 10^{-3}  \Rightarrow \tilde M = 3.9,
 \end{equation*}
 using the Indiana non-random sample.  For the IPW estimate using Facebook data, we have $\tilde M = 3.3$. Therefore, the impact of self selection on bias of $\hat R_t$ appears strong using unweighted data, remaining moderate even using the weighted estimates.

 A sensitivity analysis is performed by considering the range of false negative and false positive rates.  Here, sensitivity is assumed to fall between $81$\% and $91$\% and specificity between $95$\% and $98.8$\%~\cite{Katz2020}, which leads to a range for $M$ of $(5.60, 6.44)$ for the unweighted analysis, and $(3.63,4.07)$ and $(3.10, 3.45)$ for the weighted analyses respectively.

 \subsubsection{Time-varying IPW estimator}
 \label{section:tvipw}

 The IPW analysis is next extended to the time-varying setting.  To do so, we make use of the Facebook symptom survey, hospitalization, and testing data.  Recall the COVID-19 testing and positive case counts are known by age, gender, and racial demographics. Symptom status (e.g., fever, coughing, shortness of breath) and related covariate information (COVID-19 positive contact), however, are currently unavailable\footnote{Whether this is due to data privacy considerations or lack of data collection efforts is unclear.}.  As COVID-19 symptom status is likely to alter the testing propensity and likelihood of active infection, for illustrative purposes Facebook symptom survey and hospitalization data are used in two ways to impute symptom status per strata. We outline the imputation methods below; see Section~\ref{section:discussion} for additional discussion.

 In the first approach, we estimate two logistic regressions using weighted pseudo-likelihoods with the Facebook survey symptom data.  The first estimates the probability of contact with a COVID-19 positive individual given demographic and COVID-19 test status (positive or negative COVID-19 test). Figure~\ref{fig:contactlik1} and~\ref{fig:contactlik2} in Section~\ref{app:in_add_details} in the supplementary materials presents the estimated likelihood of contact with a COVID-19 positive individual given a negative and positive COVID-19 test in the past 24 hours respectively.  The second estimates the probability of fever given demographic, COVID-19 test status, \emph{and} whether the individual has had contact with a COVID-19 positive individual.  Figure~\ref{fig:symptomlik1} and~\ref{fig:symptomlik2} in Section~\ref{app:in_add_details} in the supplementary materials presents the estimated likelihood of reporting a fever with a COVID-19 positive individual given a negative and positive COVID-19 test in the past 24 hours respectively.  We see that likelihood of fever varies greatly based on whether the individual also reported contact with a COVID-19 positive individual. Using these two models, mean imputation is used to calculate the number of positive and negative tests given demographic strata, fever status, and COVID-19 contact status. Inverse probability weights are computed based on the resulting dataset (termed IPW1).

 \begin{figure}[!th]
 \centering
 \begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{../figs/tvprop_alt_fig1_mainpaper.png}
  \caption{Testing Likelihood Given Symptoms}
  \label{fig:testinglik1_mainpaper}
 \end{subfigure}%
 \begin{subfigure}{.5\textwidth}
  \centering
 \includegraphics[width=.9\linewidth]{../figs/tvprop_alt_fig2_mainpaper.png}
  \caption{Testing Propensity Given No Symptoms}
  \label{fig:testinglik2_mainpaper}
 \end{subfigure}
 \caption{Testing Propensity across age strata and COVID-19 symptom status}
 \label{fig:testinglik_mainpaper}
 \end{figure}

 In the second approach, we leverage hospitalization information. We first estimate the probability of fever given demographic and COVID-19 test status.  When the COVID-19 test status is positive, the probability of fever is allowed to depend on whether the individual was hospitalized.  Figure~\ref{fig:symptomlik1_model2} and~\ref{fig:symptomlik2_model2} in Section~\ref{app:in_add_details} in the supplementary materials presents the estimated likelihood of reporting a fever with a COVID-19 positive individual given a negative and positive COVID-19 test in the past 24 hours respectively.  We see that hospitalization significantly increases the likelihood of fever as expected.  We use these models and the case-to-hospitalization rates by demographic strata to perform mean imputation of the number of positive and negative tests given demographic strata, fever status, and COVID-19 contact status. Inverse probability weights are then computed based on the resulting dataset (termed IPW2).

 Figure~\ref{fig:testinglik_mainpaper} presents the testing propensity for non-Hispanic, white males across age strata and fever status using the second approach.  Note that the ratio of testing propensity within strata across fever status is time-varying, starting 50 times more likely in early April and dropping to ten times more likely by end of 2020.

 \subsection{Application of SEIR model}
 \label{section:application_of_modelbased}

 Here we consider the multinomial SEIR model using observed death data as presented in Figure~\ref{fig:in-deaths}. The model, presented in Section~\ref{section:modelbased}, requires infection fatality rates to be specified per-strata.  In this paper, similar to~\cite{Johndrow2020}, we specify fixed per-strata infection fatality rates.  Based on published age-specific IFRs~\citep{Levin2020}, IFR closely follows a log-linear relationship with age. \cite{Ironse2103272118} uses Indiana's random survey and death data to estimate a marginal IFR of $0.84\%$.  Combining across published age-specific IFRs and anchoring our analysis to the estimated marginal IFR of $0.84\%$ for Indiana, Table~\ref{tab:ifrperage} presents the age-specific IFRs used in the analysis. Note the model-based approach collapses strata for ages less than $40$ given the limited number of COVID-19 related deaths in this age range.
 \begin{table}[!th]
 \begin{tabular}{c | c c c c c c c c}
 Age Strata & 0-40 & 50-59 & 60-69 & 70-79 & 80+ \\ \hline
 IFR & $0.014\%$ & $0.120\%$ & $1.206\%$ & $3.815\%$ & $12.920\%$
 \end{tabular}
 \caption{IFR by age-strata}
 \label{tab:ifrperage}
 \vspace{-1cm}
 \end{table}
 The discrete-time distribution~$\{ \theta_{j} \}$ is a discretized version of a truncated normal distribution with mean $25$, standard deviation~$5$, minimum value $0$, and maximum value $44$.  This was chosen to closely mirror the distribution from~\cite{Johndrow2020}.

  \begin{figure}
 \centering
 \begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{../figs/deaths_ppc.png}
 \caption{PPC on aggregate daily death counts.}
 \label{fig:deathsppc}
 \end{subfigure}%
 \begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{../figs/cases_ppc.png}
 \caption{Posterior on daily case counts.}
 \label{fig:casesppc}
 \end{subfigure}
 \caption{Posterior predictive check (PPC) on death counts and posterior distribution of case counts.}
 \label{fig:ppcs}
 \end{figure}

 The SEIR model in equation~\eqref{eq:seir} is extended to allow for time-varying transmission rates~$\beta_t$ to account for public policy changes discussed in Section~\ref{section:publicpolicyindiana} over the time period considered.  Here, we insert three change points $(t_1,t_2,t_3)$ on March 23rd, June 15th, and October 1st of 2020 respectively.  Given sensitivity to these choices, instead of considering a simple multiplicative structure, we allow for delayed implementation and slow change in the transmission rate by using weights:
 $$
 \beta_t = \sum_{k=1}^{3} \left(\beta_{k-1} (1-w^{(k)}_t ) + \beta_{k} w_t\right),\quad w^{(k)}_t = (1 + \exp \left( -\xi (t - t_k - \nu) \right))^{-1} {\bf 1}[t \geq t_k]
 $$
 where~$(\beta_0,\beta_1, \beta_2, \beta_3)$ is the vector of transmission rate parameters, $\nu \geq 0$ is the potential delay, $\xi$ is the rate of change, and ${\bf 1}[\cdot]$ is an indicator function.  Each weight~$w_t^{(k)} = 0$ for $t < t_k$ and $w_t^{(k)} \to 1$ as $t \to \infty$. See Section~\ref{app:prior_modelbased} in the supplementary materials for details on prior specification.

 To check how well our model fits the observed data, a posterior predictive check on aggregate death counts was performed. Figure~\ref{fig:deathsppc} shows the results which suggests chosen SEIR model fits the observed data well. Figure~\ref{fig:casesppc} presents the posterior distribution of total cases which can be used to estimate the active infection rate.  Figure~\ref{fig:undercounting} in the supplementary materials computes the weekly undercount factor of observed case cases.  Posterior distributions for the age-specific case counts will be used as part of the doubly-robust estimator of the active infection rate.

 \subsection{Time-varying prevalence estimates}

 Here, we construct active infection rate estimates using the unweighted, inverse-probability weighted, model-based, and doubly robust estimators.  Figure~\ref{fig:tvestimates} presents the results. The inverse-probability weighting methods does reduce bias compared to the unweighted estimates. Bias, however, does remain which we conjecture is due to the limited availability of symptom and other important covariate information to estimate strata-specific infection rates. This suggests better data collection in the future may strongly improve performance of the IPW methodology.

 While the model-based estimates appear more reasonable, note that the estimate on the last week of April is 0.25\% which is a severe underestimate compared to the estimated prevalence using the stratified random sample. Caution is therefore warranted when interpreting these estimates.  Infection fatality rates are likely time-varying and may vary by other factors such as quality of healthcare.  Therefore, improvements could be made by using more accurate IFR estimates, but there is currently minimal publicly available data to do so.  Moreover, the (thankfully) low number of deaths per strata make uncertainty in these estimates quite high.  Finally, the doubly robust estimate appears similar to the IPW estimate.

Figure~\ref{fig:air_cis} in the supplementary materials presents the confidence intervals per time point for the IPW2 estimator.  The confidence interval length decreases substantially over time, reflecting the increased testing capacity.  Due to the number of surveys per week, there is minimal uncertainty in the parameter estimates.  As the number of tests per week increases to over one hundred thousand, there is therefore minimal uncertainty in the active infection rate estimates. This points to the importance of the statistical decomposition~\eqref{eq:statdecomp2} and the discussion in Section~\ref{section:IPWerrordecomp}.

  % We also construct ratio estimates as discussed in Section~\ref{section:rates}. Figure~\ref{fig:air_ratio} presents the results.  Unweighted, IPW, and doubly robust estimates appear similar.  Recall the severe bias during the growth phase due to selection bias dicussed in Section~\ref{section:rates} may imply the observed rates are  overestimated between October 2020 and January 2021.  The model-based methods may be as well.

 \begin{figure}[!th]
 % \centering
 %\begin{subfigure}{.5\textwidth}
 \centering
 \includegraphics[width=.7\linewidth]{../figs/tv_air.png}
 % \caption{Active Infection Rate Estimates}
 % \label{fig:air}
 %\end{subfigure}%
 %\begin{subfigure}{.5\textwidth}
 % \centering
 % \includegraphics[width=.9\linewidth]{../figs/tv_air_cis.png}
 % \caption{IPW2 estimate with confidence intervals}
 % \label{fig:air_cis}
 % \end{subfigure}
 \caption{Time-varying active infectionrate estimates based on unweighted, model-based, IPW and doubly robust methods.}
 \label{fig:tvestimates}
 \end{figure}

 \section{Discussion}
 \label{section:discussion}

 There is nothing routine about COVID-19, including the corresponding statistical questions.  The goal of this paper was to point out questionable statistical routines.  Precision in reported case-count data gives the illusion of information when what what is needed is quantification of uncertainty. Extensions of recent statistical error decompositions~\citep{Meng2018} demonstrate how selection bias leads data analysts to feel certain about incorrect conclusions.  As case-count data is routinely used for public health policy making, we presented an inverse-probability weighting method and a doubly-robust estimation method that leverage auxiliary information collected through random samples to overcome these issues. We end with a brief discussion of important related topics.

 \subsection*{Data quantity versus quality}

 Governments and policy makers often implicitly argue that increased testing capacity will alleviate selection bias.  Without complete compliance, however, our understanding of future outbreaks may be plagued by self-selection bias, compounded by changing sensitivity and specificity rates of RT-PCR testing. Random testing removes these effect modifiers, giving governments more information to fight the disease.  This paper emphasizes that data quantity is secondary to both data quality and methodological considerations to account for selection bias.

% \subsection*{Higher quality information on prevalence.}

 % Governments are pushing for increased testing capacity and robust contact tracing.  Contact tracing is incredibly useful for identifying carriers early and preventing spread of the disease.  We argue that understanding active infections is also key.  We know that the probability of an outbreak is a function of current prevalence and network connectivity.  Epidemic critical thresholds have been derived for many infectious disease models~\citep{Pastor2001,Newman2002,Parshani2010}.  Knowing prevalence will help governments determine long-term community-level risk and allow for more targeted interventions -- shutting down only certain locations when necessary.

 \subsection*{Model-based solutions.}

 A common argument is that the SEIR model can be extended to account for selection bias and measurement error directly; therefore, there is no need for auxiliary random sampling.  Without strong assumptions on the selection mechanism, however, the estimates are often not identifiable.  When an issue ``cannot be resolved nonparametrically then it is usually dangerous to resolve it parametrically''\citep{CoxHink74}. Absent some type of random sampling, the best route forward for all data analyses is careful associated sensitivity analyses and humility in data-driven conclusions.  \cite{Ironse2103272118}, for example, use the Indiana seroprevalence random survey and to anchor their analysis; however, infrequent random surveys imply time-varying aspects such as the IFR may not be resolved by use of a single anchor.

 %\textcolor{red}{Underascertainmnet as pre-processed or modelled but often is an assumption that must be resolved parametrically. }

 \subsection*{Real-world implementation of the proposed method}

 Application of the proposed methods relied on access to two critical datasets: (1) COVID-19 testing, case, hospitalization, and death data by demographic strata and (2) Facebook symptom survey.  Public access to (1) required the author to submit a data request directly to the state of Indiana.  Through conversations with the state of Indiana, the author's request was one of only five data requests to be approved by Indiana for public release.  Most states do not release such granular data, making selection bias adjustments difficult. Coordinated, systematic data collection and reporting is critical.
 Lack of covariate information including symptom and COVID-19 contacts on individuals seeking COVID-19 tests is unacceptable.  Random samples over time to supplement this data with auxiliary information is desperately needed.  We acknowledge that this proposal will lead to data privacy concenrs that will need to be addressed.

 % While we are eternally grateful to Indiana for the release, symptom information was not made available.  This severely limits our ability to estimate correct selection propensity models and strata-specific infection rates.  A more sustained effort to collect relevant measures of an individual's past history that may impact their likelihood of seeking testing and being COVID-19 positive is required.  Such data collection and dissemination is critical to address selection bias.

 A valid criticism of the proposed approach is that recording covariates on every tested individual is too time-consuming and costly, i.e., the proposed data collection protocol does not scale well.  We argue that the list of relevant factors is not too long.  For example, symptom status and COVID-19 contact are clearly relevant factors in testing selection.  Future work may consider how to include/exclude various factors over time that are not significant to limit citizen burden in reporting.  Moreover, equation~\eqref{eq:auxinfoprob} presupposes covariate information is collected for all individuals in the nonprobability sample; however, this is often not feasible for state and local governments where rapid sample collection is prioritized over additional collection of covariate information.  We can easily extend~\eqref{eq:auxinfoprob} by randomly sampling a subset of the nonprobability sample on which to collect the covariate information.
 %yielding a new pseudo-likelihood
 % \begin{equation*}
 % % \label{eq:auxinfoprob2}
 % \sum_{j=1}^N I_j I_j^{(2)} W_j^{(2)} \log \left( \frac{\pi (X_j; \theta)}{1-\pi(X_j; \theta)} \right)  + \sum_{i=1}^N \tilde I_j \tilde W_j \log ( 1 - \pi (X_j; \theta)).
 % \end{equation*}
 % Here, $I_j^{(2)}$ is an indicator that the individual is in the probability subsample and $W_j^{(2)}$ is the probability of inclusion. Estimation proceeds by maximizing the pseudo-likelihood which now depends on random sampling in both the first and second term.
 This would allow for rapid testing and minimize data collection costs.

\newpage
% \textcolor{red}{Fix references!}
\bibliographystyle{plainnat}
\bibliography{covid-refs}

% \newpage
% % \appendix

% % \section{Reproducible code}

% % All relevant code can be found at \url{https://github.com/wdempsey/covid-umich}.

% % \section{Technical details}

% % Recall that $P_j$ is  an indicator of measurement error, equal to $1$ when we incorrectly measure the outcome and $0$ when we observe the true outcome. We suppose this is a stochastic variable where $\pr(P_j = 1 \mid Y_j = 1) =: FN$ is the false-negative rate and $\pr(P_j = 1 \mid Y_j = 0) =: FP$ is the false-positive rate.  If individual $j$ is selected (i.e., $I_j = 1$) then the observed outcome can be written as $Y_j^{\star} = Y_j(1-P_j) + (1-Y_j) P_j$.

% % \subsection{Derivations for imperfect testing framework.}
% % \label{app:imperfect}
% % We start by considering the empirical mean estimator under imperfect testing,
% % $$
% % \bar y_n^\star = \frac{\sum_{j=1}^N Y_j^\star I_j}{\sum_{j=1}^N I_j} = \frac{\sum_{i=1}^N  I_j Y_j^\star }{\sum_{j=1}^N  I_j } = \frac{\sum_{i=1}^N  I_j \left[ Y_j (1-P_j) + (1-Y_j) P_j \right]}{\sum_{j=1}^N  I_j }
% % $$
% % For any set of numbers $\{ A_1, \ldots, A_N \}$ we can view it as the support of a random variable $A_J$ induced by the random index $J$ defined on $\{1,\ldots, N\}$.  When $J$ is uniformly distributed $E_J (A_J) = \sum_{j=1}^N A_j / N \equiv \bar A_N$. Then
% % $$
% % \begin{aligned}
% % \bar y_n^\star  - \bar Y_N &= \frac{E_J \left[ I_J \left[ Y_J (1-P_J) + (1-Y_J) P_J \right] \right]}{E_J [ I_J ] } - E_J[Y_J] \\
% % &= \frac{E_J \left[ I_J P_J (1-2Y_J) \right]}{E_J [ I_J ] } + \left( \frac{E_J [I_J Y_J]}{E_J [ I_J ] } - \frac{E_J[Y_J] E_J[I_J]}{E_J[I_J]} \right) \\
% % \end{aligned}
% % $$
% % The term in parentheses can be re-written as
% % $$
% % \begin{aligned}
% % \frac{E_J [I_J Y_J]- E_J[Y_J] E_J[I_J]}{E_J[I_J]} &=  \frac{E_J [I_J Y_J]- E_J[Y_J] E_J[I_J]}{\sqrt{V_J(I_J) V_J(Y_J)}} \frac{\sqrt{V_J(I_J)}}{E_J[I_J]} \times \sqrt{V_J(Y_J)} \\
% % &= \rho_{I,Y} \times \sqrt{\frac{(1-f)}{f}} \times \sigma_Y
% % \end{aligned}
% % $$
% % which agrees with Meng's (2019) decomposition. For the other term, first we define $Z_j := 1 - 2 Y_j $. Then $Z_j = 1$ if $Y_j = 0$ and $Z_j = -1$ if $Y_j = 1$. Then the term can be re-written as
% % $$
% % \begin{aligned}
% % \frac{E_J \left[ I_J P_J (1-2Y_J) \right]}{E_J [ I_J ] } &= \left( \frac{E_J \left[ I_J P_J Z_J \right]}{E_J [ I_J ] } -  \frac{E_J \left[ P_J Z_J \right] E_J[ I_J]}{E_J [ I_J ] } \right) +  \frac{E_J \left[ P_J Z_J \right] E_J[ I_J]}{E_J [ I_J ] } \\
% % \end{aligned}
% % $$
% % The term in parentheses can be re-expressed using the previous technique as:
% % $$
% % \rho_{I, PZ} \times \sqrt{\frac{1-f}{f}} \times \sigma_{PZ}
% % $$
% % where now the ``data defect'' and ``problem difficulty'' are with respect to $PZ$ rather than $Y$. The final term is equal to
% % $$
% % \begin{aligned}
% % E_J [P_J Z_J ] &= E_J [ E_J [ P_J Z_J \mid Y_J ] ] \\
% % &= \pr (P = 1 \mid Y = 0) (1-\bar Y) - \pr(P=1 \mid Y = 1) \bar Y \\
% % &= FP - (FP + FN) \cdot \bar Y
% % \end{aligned}
% % $$
% % Combining these yields:
% % $$
% % \bar y_n^\star - \bar Y = \sqrt{\frac{1-f}{f}} \left(\rho_{I,Y} \sigma_Y + \rho_{I, PZ} \sigma_{PZ} \right) + \left( FP - (FP+FN) \bar Y \right)
% % $$

% % \subsubsection{Derivation of an estimator unbiased under SRS}
% % \label{app:memestimator}
% % We see the final term is given by $FP (1-\bar Y) - FN \bar Y$ is the bias associated with using the unadjusted prevalence estimate $\bar y_n^\star$.  This motivates an adjusted estimate
% % $$
% % \tilde y_n^{(0)} = \bar y_n^\star - FP (1- \bar y_n^\star ) + FN \bar y_n^\star
% % = \bar y_n^\star (1 + FN + FP) - FP.
% % $$
% % Now considering the error for the adjusted estimate, $\tilde y_n^{(0)} - \bar Y$, we have
% % $$
% % \begin{aligned}
% %  &\bar y_n^\star (1 + FN + FP) - FP - \bar Y \\
% %  =&(\bar y_n^\star - \bar Y) +  (FN + FP) \bar y_n^\star - FP  \\
% %  =&\underbrace{\sqrt{\frac{1-f}{f}} \left[\rho_{I,Y} \sigma_Y + \rho_{I, PZ} \sigma_{PZ} \right]}_{\Psi} + (  FN + FP ) (\bar y_n^\star -  \bar Y) \\
% %  =& \Psi + (FN + FP) \Psi + (FN + FP) (FP - (FP+FN) \bar Y).
% %  \end{aligned}
% % $$
% % The final term is a (smaller) bias term and so we propose another adjusted estimator $\tilde y_n^{(1)} = \tilde y_n^{(0)} + (FN+FP)  ( (FN+FP) \bar y_n^\star - FP)$, with associated error $\tilde y_n^{(1)} - \bar Y$ given by
% % $$
% % \begin{aligned}
% %  &(\bar y_n^{(0)} - \bar Y) + (FN+FP)  ( (FN+FP) \bar y_n^\star - FP)\\
% %  =&\Psi + (FN + FP) \Psi + (FN + FP) (FP - (FP+FN) \bar Y) + (FN + FP) ((FP+FN) \bar y_n^\star - FP)  \\
% %   =& \Psi + (FN + FP) \Psi + (FN + FP)^2 \Psi + (FN+FP)^2 (FP - (FP+FN) \bar Y).
% %  \end{aligned}
% % $$
% % This motivates recursively defining estimators $\tilde y_n^{(t)} = \tilde y_n^{(t-1)} + (FN+FP)  ( (FN+FP) \bar y_n - FP)$ for $t=1,2,\ldots$ where $\tilde y_n^{(0)} = \bar y_n^\star$.  Then
% % $$
% % \tilde y_n^{(t)} = \bar y_n^\star \sum_{s=0}^{t+1} (FP+FN)^s - FP \sum_{s=0}^{t} (FP+FN)^s
% % $$
% % and the associated error at iteration $t$ given by
% % $$
% % \Psi \sum_{s=0}^t (FN+FP)^s = \Psi \frac{1 - (FN+FP)^t}{1 - (FN+FP)}.
% % $$
% % We can then get an estimator with no residual bias term by taking the limit as $t$ goes to infinity; that is, define
% % $$
% % \tilde y_n = \lim_{t \to \infty} \tilde y_n^{(t)} = \frac{\bar y_n^\star - FP}{1 -(FN+FP)}.
% % $$
% % Then the associated error $\tilde y_n - \bar Y$ can be expressed as $\frac{\Psi}{1-(FN+FP)}$.

% % \subsection{Model-based derivation of the estimator.}
% % \label{app:modelbased}
% % The estimator $\tilde y$ was derived as a limit of a process that removes the residual bias term at each step.  Here we consider a model-based explanation.  Let $\theta = \pr( Y = 1)$ and $\phi = \pr( \text{test is positive})$.  Given a known false negative (FN) and false positive (FP) rates, we have
% % $$
% % \begin{aligned}
% % \phi &= \theta \cdot (1-FN) + (1-\theta) \cdot FP = \theta (1-FN-FP) + FP \\
% % \Rightarrow \theta &= \frac{\phi - FP}{1-FN-FP}.
% % \end{aligned}
% % $$
% % Thus, the estimator $\tilde y$ is also the appropriate estimator under a model-based approach.  While the derivation here is more straightforward, the derivation in the prior section provides a simple formula for the associated error $\tilde y_n - \bar Y$ and gives a novel connection between the empirical estimator $\bar y_n^\star$ and the adjusted estimator $\tilde y_n$ without reference to the model-based approach.

% % \subsection{Further simplification.}
% % For the binary outcome $Y$, we have $\sigma_Y = \sqrt{\bar Y (1-\bar Y)}$. Moreover,
% % $$
% % \begin{aligned}
% % V_J(P_J Z_J) &= E_J[(P_J Z_J)^2] - E_J[P_J] E_J[Z_J] \\
% % &= E_J[P_J] - E_J[P_J] (1 - 2 \bar Y) = 2 \bar Y E_J [ P_J ] \\
% % &= 2 \bar Y \left( FP (1-\bar Y) + FN \bar Y \right) \\
% % \Rightarrow \sigma_{PZ} &= \sqrt{ 2 \bar Y \left( FP (1-\bar Y) + FN \cdot  \bar Y \right) }
% % \end{aligned}
% % $$
% % Then the formula for the error is given by:
% % \begin{equation}
% % \label{eq:finalstep}
% % \sqrt{\frac{1-f}{f}} \left[\rho_{I,Y} \sqrt{\bar Y (1-\bar Y)} + \rho_{I, PZ} \sqrt{ 2 \bar Y \left( FP (1-\bar Y) + FN \cdot \bar Y \right )} \right] \times \frac{1}{1- (FN+FP)}
% % \end{equation}
% % By definition, we have
% % $$
% % \begin{aligned}
% % \rho_{I,PZ} &= \frac{C(I, PZ)}{\sqrt{V(PZ) V(I)}} \\
% % &= \frac{C(I, PZ)}{\sqrt{V(Y) V(I)}} \sqrt{\frac{V(Y)}{V(PZ)}} \\
% % &= \rho_{I,Y} \frac{C(I,PZ)}{C(I,Y)} \sqrt{ \frac{(1-\bar Y)}{2 ( FP (1-\bar Y) + FN \cdot \bar Y)} }
% % \end{aligned}
% % $$

% % $$
% % \begin{aligned}
% % C(I, PZ) &= E[ I P Z ] - E[I] E[PZ] \\
% % &=  [FP f_0 - (FP f_0 + FN f_1) \bar Y] - f [ FP - (FP+FN) \bar Y ] \\
% % &=  - FP \Delta \bar Y + FP \bar Y^2 \Delta - FN \bar Y^2 \Delta \\
% % &=  - \Delta \bar Y (FP \cdot (1-\bar Y) + FN \cdot \bar Y) \\
% % \end{aligned}
% % $$
% % where $f = f_1 \bar Y + f_0 (1-\bar Y)$ so $f_0 - f = -\Delta \bar Y$ and $f_1 - f = \Delta (1-\bar Y)$.
% % $$
% % \begin{aligned}
% % C(I, Y) &= E[ I Y ] - f \bar Y \\
% % &=  f_1 \bar Y + f_0 (1-\bar Y) - f \bar Y \\
% % &=  f_0 (1-\bar Y) + \Delta (1-\bar Y) \bar Y \\
% % &= (1-\bar Y) (f_0 + \Delta \bar Y)
% % \end{aligned}
% % $$
% % Combining yields
% % $$
% % \begin{aligned}
% % \rho_{I,PZ} &= \rho_{I,Y} \times \frac{- \Delta \bar Y (FP \cdot (1-\bar Y) + FN \cdot \bar Y) }{(1-\bar Y) (f_0 + \Delta \bar Y)} \times \sqrt{ \frac{(1-\bar Y)}{2 ( FP (1-\bar Y) + FN \cdot \bar Y)} } \\
% % &= - \rho_{I, Y} \times \Delta \times \sqrt{\frac{\bar Y}{1-\bar Y}} \frac{\sqrt{FP(1-\bar Y) + FN \cdot \bar Y}}{f_0 (1-\bar Y) + f_1 \bar Y} \times \sqrt{\frac{\bar Y}{2}}
% % \end{aligned}
% % $$
% % We can then re-write $\rho_{I,Y} \sigma_Y + \rho_{I,PZ} \sigma_{PZ}$ as
% % $$
% % \rho_{I,Y} \sigma_Y \left( 1 - \Delta \times \frac{\bar Y}{1-\bar Y} \times \frac{FP(1-\bar Y) + FN \cdot \bar Y}{f_0 (1-\bar Y) + f_1 \bar Y} \right).
% % $$
% % Inserting into equation~\eqref{eq:finalstep} yields the desired result.

% % \subsection{Derivation of effective sample size}
% % \label{app:effss}

% % Let $S_Y^2 = (N-1)^{-1} \sum_{j=1}^N (Y_j - \bar Y)^2$ be the population variance as defined in survey sampling~\citep{Cochran77}.  Then $\sigma_Y^2 = (N-1)/N \cdot S_Y^2$.  Under SRS, the MSE is the variance as the estimate is unbiased and the variance is given by $(1-f)/n S_Y^2$.  Then setting the MSE under general selection and SRS equal we have
% % $$
% % \begin{aligned}
% % \underbrace{\frac{1-f}{f} \times E_{\I} \left[ \rho_{I, Y}^2 \times D_M^2 \right]}_{1/n_{eff}^\star} \times \sigma_Y^2 &= \frac{1-f}{n_{eff}} S_Y^2 \\
% % \frac{1}{n_{eff}^\star} \times \frac{N-1}{N} S_Y^2 &= \frac{1-f}{n_{eff}} S_Y^2 \\
% % \frac{1}{n_{eff}^\star} &=  \left( \frac{1}{n_{eff}} - \frac{1}{N} \right) \left( \frac{N}{N-1} \right) \\
% % \frac{1}{n_{eff}^\star} \left[ 1 - \frac{1}{N} + \frac{n_{eff}^\star}{N} \right]  &=  \frac{1}{n_{eff}} \\
% % n_{eff}^\star \left[ 1 - \frac{1}{N} + \frac{n_{eff}^\star}{N} \right]^{-1}  &=  n_{eff} \\
% % \frac{n_{eff}^\star}{ 1 + (n_{eff}^\star -1) N^{-1}} &=  n_{eff} \\
% % \end{aligned}
% % $$
% % Then if $n_{eff}^\star \geq 1$, we have that
% % $$
% % n_{eff} \leq n_{eff}^\star = \frac{f}{1-f} \times \frac{1}{E_{\I} \left[ \rho^2_{I, Y} \times D_M^2 \right]}
% % $$

% % \subsection{Ratio estimator}
% % \label{app:ratio}

% % Let ${\bf u} = (u_{t-1},u_t) \in \mathbb{R}^2$ and $g({\bf u}) = \frac{u_t}{u_{t-1}}$, i.e., a differentiable function $g:\mathbb{R}^2 \to \mathbb{R}$. Centering a Taylor series expansion of second-order around coordinates $(U_2, U_1) \in \mathbb{R}^2$ yields
% % $$
% % \begin{aligned}
% % g({\bf u}) =& g(U_{t-1}, U_t) - \frac{U_t}{U_{t-1}^2} (u_{t-1} - U_{t-1}) + \frac{1}{U_{t-1}} (u_t - U_t) \\
% % &+ \frac{1}{2} \left[ \frac{2 U_t}{U_{t-1}^3} (u_{t-1} - U_{t-1})^2 + 0 \times (u_t - U_t)^2 - 2 \times (u_{t-1} - U_{t-1}) (u_t - U_t) \frac{1}{U_t^2} \right]
% % \end{aligned}
% % $$
% % Plugging in $(\bar y_{t-1}, \bar y_t)$ for $(u_{t-1}, u_t)$ and $(\bar Y_{t-1}, \bar Y_t)$ for $(U_{t-1}, U_t)$ yields the $\frac{\bar y_t}{\bar y_{t-1}} - \frac{\bar Y_t}{\bar Y_{t-1}} $ is equal to
% % $$
% % \begin{aligned}
% % =&
% % - \frac{\bar Y_t}{\bar Y_{t-1}^2} (\bar y_{t-1} - \bar Y_{t-1}) + \frac{1}{\bar Y_{t-1}} (\bar y_t - \bar Y_t) \\
% % &+ \frac{\bar Y_t}{\bar Y_{t-1}^3} (\bar y_{t-1} - \bar Y_{t-1})^2 -  (\bar y_{t-1} - \bar Y_{t-1}) (\bar y_t - \bar Y_t) \frac{1}{\bar Y_{t-1}^2} \\
% % &= \frac{\bar Y_t}{\bar Y_{t-1}} \bigg[  \rho_{I_t,Y_t} \sqrt{\frac{1-f_t}{f_t}} CV (Y_t)  -\rho_{I_{t-1},Y_{t-1}} \sqrt{\frac{1-f_{t-1}}{f_{t-1}}} CV (Y_{t-1}) \\
% % &+ \rho^2_{I_{t-1},Y_{t-1}} \frac{1-f_{t-1}}{f_{t-1}} CV^2 (Y_{t-1}) -  \rho_{I_{t-1},Y_{t-1}} \sqrt{\frac{1-f_{t-1}}{f_{t-1}}} CV (Y_{t-1}) \times
% % \rho_{I_t,Y_t} \sqrt{\frac{1-f_t}{f_t}} CV (Y_t)   \bigg] \\
% % &= \frac{\bar Y_t}{\bar Y_{t-1}} \bigg[ \rho_{I_t,Y_t} \sqrt{\frac{1-f_t}{f_t}} CV (Y_t)  -\rho_{I_{t-1},Y_{t-1}} \sqrt{\frac{1-f_{t-1}}{f_{t-1}}} CV (Y_{t-1}) \bigg] \left[ 1 - \rho_{I_{t-1},Y_{t-1}} \sqrt{\frac{1-f_{t-1}}{f_{t-1}}} CV (Y_{t-1}) \right]
% % \end{aligned}
% % $$
% % where the second equality is obtained by plugging in the statistical decomposition of the error for both time points and the coefficient of variation being defined as $CV(Y) := \sigma_Y/\mu_Y$.  Under measurement error, the extra terms $D_{t}$ and $D_{t-1}$ can be inserted in the correct locations.

% % \subsection{Estimation of effective reproduction number}

% % Let
% % $$
% % \delta_t := \bigg[ \rho_{I_t,K_t} D_{M_t} \sqrt{\frac{1-f_t}{f_t}} CV (K_t)  -\rho_{I_{t-1},K_{t-1}} D_{M_{t-1}} \sqrt{\frac{1-f_t}{f_t}} CV (K_{t-1}) \bigg].
% % $$
% % Then the previous sections derivation shows that the estimate of the number of new cases on day t is given by
% % $$
% % \frac{S_t \cdot \bar y_t}{S_{t-1} \cdot \bar y_{t-1}} =
% % \frac{K_t}{K_{t-1}} \left( 1 + \delta_t \times \left[ 1 - \rho_{I_{t-1},K_{t-1}} D_{M_{t-1}} \sqrt{\frac{1-f_t}{f_t}} CV (K_{t-1}) \right] \right)
% % $$
% % Then setting $e_t = \delta_t \times [1 - \rho_{I_{t-1},K_{t-1}} D_{M_{t-1}} \sqrt{\frac{1-f_t}{f_t}} CV (K_{t-1}) ]$, we have
% % $$
% % \begin{aligned}
% % \log \left( \frac{S_t \bar y_t}{S_{t-1} \bar y_{t-1}} \right) - \log \left( \frac{K_t}{K_{t-1}} \right) &= \log (1 + e_t) \\
% % \log \left( \frac{\bar y_t}{\bar y_{t-1}} \right) - \log \left( \frac{K_t}{K_{t-1}} \right) &= 1 + e_t - \log \left( \frac{S_t}{S_{t-1}} \right) \\
% % 1 + \frac{1}{\gamma} \log \left( \frac{\bar y_t}{\bar y_{t-1}} \right) - \left[ 1 + \frac{1}{\gamma} \log \left( \frac{K_t}{K_{t-1}} \right) \right] &= \frac{1}{\gamma} \left[ \log \left( 1 + e_t \right) - \log \left( \frac{S_{t}}{S_{t-1}} \right) \right] \\
% % \Rightarrow \hat R_t - R_t &= \frac{1}{\gamma} \left[ \log \left( 1 + e_t \right) - \log \left( \frac{S_{t}}{S_{t-1}} \right) \right]
% % \end{aligned}
% % $$

% % \subsection{Computing the effective sample size}
% % \label{section:effss}

% % For binary outcomes, we have
% % \begin{equation} \label{eq:binaryrho}
% % \rho_{I,Y} = \Delta \sqrt{\frac{\bar Y (1 - \bar Y)}{f (1-f)} }
% % \end{equation}
% % where $\Delta = P_J (I_J = 1 \mid Y_J = 1) - P(I_J = 1 \mid Y_J = 0) = f_1 - f_0$.  Suppose that $M = f_1/f_0$; then $f_0 = f / (\bar Y \cdot (M-1) + 1)$.  Using the upper bound $E_{\I} [ \rho_{I,Y}^2 ] \leq E_{\I} [\rho_{I,Y} ]^2$, we compute effective sample size under a range of prevalences $\bary$, and relative sample rates $M$ given $f = 0.003$ (i.e., current sampling fraction).

% % \begin{table}[ht]
% % \centering
% % \begin{tabular}{rrrrrrrr}
% % & \multicolumn{7}{c}{$M$} \\ \cline{2-8}
% % $\bar y$ & 1.05 & 1.15 & 1.25 & 1.35 & 1.45 & 1.55 & 1.65 \\
% %   \hline
% % 0.01 & 40444 & 4503 & 1624 & 830 & 503 & 338 & 242 \\
% % 0.03 & 13787 & 1541 & 558 & 286 & 174 & 117 & 85 \\
% % 0.05 & 8463 & 950 & 345 & 178 & 109 & 73 & 53 \\
% % 0.07 & 6187 & 697 & 254 & 132 & 81 & 55 & 40 \\
% % 0.09 & 4928 & 557 & 204 & 106 & 65 & 44 & 32 \\
% % 0.11 & 4131 & 469 & 173 & 90 & 56 & 38 & 28 \\
% %    \hline
% % \end{tabular}
% % \end{table}

% % We also present the same plot under $FP = 0.024$ and $FN = 0.13$ to show the impact of measurement error on effective sample size.

% % \begin{table}[ht]
% % \centering
% % \begin{tabular}{rrrrrrrr}
% % & \multicolumn{7}{c}{$M$} \\ \cline{2-8}
% %  & 1.05 & 1.15 & 1.25 & 1.35 & 1.45 & 1.55 & 1.65 \\
% %   \hline
% % 0.01 & 29019 & 3247 & 1177 & 605 & 368 & 248 & 179 \\
% % 0.03 & 9894 & 1112 & 405 & 209 & 128 & 87 & 63 \\
% % 0.05 & 6075 & 686 & 251 & 130 & 80 & 54 & 39 \\
% % 0.07 & 4442 & 504 & 185 & 96 & 59 & 41 & 30 \\
% % 0.09 & 3539 & 403 & 149 & 78 & 48 & 33 & 24 \\
% % 0.11 & 2967 & 339 & 126 & 66 & 41 & 28 & 21 \\
% %    \hline
% % \end{tabular}
% % \end{table}

% % For binary outcomes, we have
% % \begin{equation} \label{eq:weightedbinaryrho}
% % \rho_{\tilde I (X),Y} = \tilde \Delta \sqrt{\frac{\bar Y (1 - \bar Y)}{f (1-f) \text{E}(W_J \mid I_J = 1)^2 + f \text{Var}(W_J \mid I_J = 1)} }
% % \end{equation}

% % % \section{Effect modifiers in COVID-19 clinical trials}

% % % Here we show how selection bias and measurement error can creep into clinical trial analysis. The key concern is whether clinical trials on COVID-19 recruit from the pool of individuals who have tested positive for COVID-19 or whether they sample randomly from the population, test, and then recruit from this subset of tested individuals.  To see this issue, suppose we have an outcome $Y$, a treatment $A$ and an \emph{unobserved} variable $U$.  Suppose treatment is assigned at random, i.e., $A =1$ with probability $50\%$ and $A=0$ with probability $50\%$.  Suppose the conditional mean of the outcome satisfies  $E(Y \mid U, A ) = \beta_0 + \beta_1 A + \beta_2 U + \beta_3 A U$.  Typically, we are interested in the \emph{causal effect} of $A$ on $Y$.  In counterfactual language,
% % % $$
% % % E( Y(A=1) - Y(A=0) ) = E_U ( E_A ( Y(A=1) - Y(A=0) \mid U=u ))
% % % = \beta_1 + \beta_3 E(U).
% % % $$
% % % Even in a randomized experiment, the question is what the correct value for $E(U)$ is. For COVID-19, the value of interest is likely the expected value of $U$ in the population of COVID-19 positive individuals, i.e., $E_J [ U_J Y_J]/ E_J[Y_J] = \sum_{j=1}^N U_j Y_j / \sum_{j=1}^N Y_j$.  Assuming no measurement error, the estimator is given by the ratio $\sum_{j=1}^n y_i u_i / \sum_{j=1} y_i$.  Due to selection bias, the bias in the marginal treatment effect compared to the true marginal effect of interest is approximately equal to
% % % $$
% % % \beta_3 \times \sqrt{\frac{1-f}{f}} \times \frac{\bar U^\prime}{\bar Y} \bigg[ \rho_{I,U^\prime} \times CV (U^\prime)  - \rho_{I,Y} \times CV (Y) \bigg] \left[ 1 - \rho_{I,Y} \sqrt{\frac{1-f}{f}} CV (Y) \right]
% % % $$
% % % where $U^\prime = U Y$.  The derivation again uses a second-order Taylor series approximation and follows exactly the same as in Appendix~\ref{app:ratio}. Selection bias then leads to the estimated effect being biased if $\beta_1 \neq 0$.  Directionality of the bias will depend on the relationship between $U$ and $Y$.  If $U$ is a measure of disease severity, then it could be expected to see $\beta_3 > 0$ and $\rho_{I,U^\prime} \times CV (U^\prime)  > \rho_{I,Y} CV (Y)$.  In such situations, the treatment effect estimates will tend to be overly optimistic as most individuals recruited are potentially higher on the severity index than the population average.  Not only that, but for fixed data quality, the error compared to SRS in terms of effect estimation scales with the population size (i.e., the law of large populations).

% % % Randomization of treatment assignment in clinical trials negates unobserved confounders.  This, however, does not negate effect modifiers.  Therefore, for marginal treatment effects to be interpretable, there must be a well-defined population.  Most often, our main interest is in causal effects on the population of COVID-19 positive.  Randomization within the clinical trial design yields internal validity, but we also need external validity to estimate the correct marginal effect of interest~\citep{Keiding2016}.

% % \section{IPW statistical error decomposition derivation}
% % \label{app:ipwderivation}
% % We start by considering the empirical weighted mean estimator under imperfect testing,
% % $$
% % \bar y_n^\star = \frac{\sum_{j=1}^N W_j Y_j^\star I_j}{\sum_{j=1}^N W_j I_j} = \frac{\sum_{i=1}^N  I_j W_j Y_j^\star }{\sum_{j=1}^N  I_j } = \frac{\sum_{i=1}^N  I_j \left[ Y_j (1-P_j) + (1-Y_j) P_j \right]}{\sum_{j=1}^N  I_j }
% % $$
% % Then
% % $$
% % \begin{aligned}
% % \bar y_n^\star  - \bar Y_N &= \frac{E_J \left[ I_J W_J \left[ Y_J (1-P_J) + (1-Y_J) P_J \right] \right]}{E_J [ I_J W_J ] } - E_J[Y_J] \\
% % &= \frac{E_J \left[ I_J W_J P_J (1-2Y_J) \right]}{E_J [ I_J W_J ] } + \left( \frac{E_J [I_J W_J Y_J]}{E_J [ I_J W_J ] } - \frac{E_J[Y_J] E_J[I_J W_J]}{E_J[I_J W_J]} \right) \\
% % \end{aligned}
% % $$
% % The term in parentheses can be re-written as
% % $$
% % \begin{aligned}
% % \frac{E_J [I_J W_J Y_J]- E_J[Y_J] E_J[I_J W_J]}{E_J[I_J W_J]} &=  \frac{E_J [I_J W_J Y_J]- E_J[Y_J] E_J[I_J W_J]}{\sqrt{V_J(I_J W_J) V_J(Y_J)}} \frac{\sqrt{V_J(I_J W_J)}}{E_J[I_J]} \times \sqrt{V_J(Y_J)} \\
% % &= \rho_{\tilde I (X),Y} \times \frac{\sqrt{V_J(I_J W_J)}}{E_J[I_J]} \times \sigma_Y
% % \end{aligned}
% % $$
% % Then
% % \begin{align*}
% % E_J (I_J W_J) &= P(I_J = 1) E_J [W_J \mid I_J = 1] = f \times E_J [W_J \mid I_J = 1] \\
% % V_J(I_J W_J) &= E[ I_J W_J^2] - E[I_J W_J]^2 \\
% % &= f \cdot \left( E [W_J^2 \mid I_J = 1] - f E[W_J \mid I_J = 1 ]^2 \right) \\
% % &= f \cdot \left( E [W_J^2 \mid I_J = 1] \pm E[W_J \mid I_j = 1]^2 - f E[W_J \mid I_J = 1 ]^2 \right) \\
% % &= f \cdot \left( V (W_J \mid I_J = 1) + E[W_J \mid I_J = 1]^2 (1-f) \right)
% % \end{align*}
% % Taking the ratio:
% % $$
% % \sqrt{\frac{f \cdot \left( V (W_J \mid I_J = 1) + E[W_J \mid I_J = 1]^2 (1-f) \right)}{f^2 E_J [W_J \mid I_J = 1]^2}} = \sqrt{\frac{1 - f + CV(W)^2}{f}}
% % $$
% % which agrees with Meng's (2019) decomposition of weighted outcome. For the other term, first we define $Z_j := 1 - 2 Y_j $. Then $Z_j = 1$ if $Y_j = 0$ and $Z_j = -1$ if $Y_j = 1$. Then the term can be re-written as
% % $$
% % \begin{aligned}
% % \frac{E_J \left[ I_J W_J P_J (1-2Y_J) \right]}{E_J [ I_J W_J ] } &= \left( \frac{E_J \left[ I_J W_J P_J Z_J \right]}{E_J [ I_J W_J ] } -  \frac{E_J \left[ P_J Z_J \right] E_J[ I_J W_J]}{E_J [ I_J W_J ] } \right) +  \frac{E_J \left[ P_J Z_J \right] E_J[ I_J W_J]}{E_J [ I_J W_J ] } \\
% % \end{aligned}
% % $$
% % The term in parentheses can be re-expressed using the previous technique as:
% % $$
% % \rho_{\tilde I, PZ} \times \sqrt{\frac{1-f+CV(W)^2}{f}} \times \sigma_{PZ}
% % $$
% % where now the ``data defect'' and ``problem difficulty'' are with respect to $PZ$ rather than $Y$. The final term is equal to
% % $$
% % \begin{aligned}
% % E_J [P_J Z_J ] &= E_J [ E_J [ P_J Z_J \mid Y_J ] ] \\
% % &= \pr (P = 1 \mid Y = 0) (1-\bar Y) - \pr(P=1 \mid Y = 1) \bar Y \\
% % &= FP - (FP + FN) \cdot \bar Y
% % \end{aligned}
% % $$
% % Combining these yields:
% % $$
% % \bar y_n^\star - \bar Y = \sqrt{\frac{1-f+CV(W)^2}{f}} \left(\rho_{\tilde I(X),Y} \sigma_Y + \rho_{\tilde I(X), PZ} \sigma_{PZ} \right) + \left( FP - (FP+FN) \bar Y \right)
% % $$
% % By previous arguments, we have
% % $$
% % \rho_{\tilde I (X), PZ} = \rho_{\tilde I(X), Y} \times \frac{C(\tilde I(X), PZ)}{C(\tilde I(X), Y)} \sqrt{ \frac{1-\bar Y}{ 2 (FP(1-\bar Y) + FN \bar Y)} }
% % $$
% % Then
% % \begin{align*}
% % C(IW, PZ) &= E[IWPZ] - E[IW] E[PZ] \\
% % &=  [FP \tilde f_0  - (FP \tilde f_0 + FN \tilde f_1) \bar Y]  - \tilde f \cdot ( FP - (FP+FN) \bar Y )\\
% % &= -FP \tilde \Delta \bar Y + FP \bar Y^2 \Delta - FN \bar Y^2 \Delta \\
% % &= -\tilde \Delta \bar Y ( FP \cdot (1-\bar Y) + FN \cdot \bar Y)
% % \end{align*}
% % where $\tilde f_i = E[ I_J W_J \mid Y_J = i]$ for $i \in \{0,1 \}$ and $f = \tilde f_1 \bar Y + \tilde f_0 (1-\bar Y)$. Moreover,
% % \begin{align*}
% % C(\tilde I,Y) &= E[ I W Y ] - \tilde f \bar Y \\
% % &= \tilde f_1 \bar Y + \tilde f_0 (1-\bar Y) - \tilde f \bar Y \\
% % &= \tilde f_0 (1-\bar Y) + \tilde \Delta (1-\bar Y) \bar Y \\
% % &= (1- \bar Y) (\tilde f_0 + \tilde \Delta \bar Y)
% % \end{align*}
% % Combining yields
% % $$
% % \begin{aligned}
% % \rho_{\tilde I (X),PZ} &= \rho_{\tilde I(X),Y} \times \frac{- \tilde \Delta \bar Y (FP \cdot (1-\bar Y) + FN \cdot \bar Y) }{(1-\bar Y) (\tilde f_0 + \tilde \Delta \bar Y)} \times \sqrt{ \frac{(1-\bar Y)}{2 ( FP (1-\bar Y) + FN \cdot \bar Y)} } \\
% % &= - \rho_{\tilde I(X), Y} \times \Delta \times \sqrt{\frac{\bar Y}{1-\bar Y}} \frac{\sqrt{FP(1-\bar Y) + FN \cdot \bar Y}}{\tilde f_0 (1-\bar Y) + \tilde f_1 \bar Y} \times \sqrt{\frac{\bar Y}{2}}
% % \end{aligned}
% % $$
% % We can then re-write $\rho_{\tilde I (X),Y} \sigma_Y + \rho_{\tilde I(X),PZ} \sigma_{PZ}$ as
% % $$
% % \rho_{\tilde I(X),Y} \sigma_Y \left( 1 - \tilde \Delta \times \frac{\bar Y}{1-\bar Y} \times \frac{FP(1-\bar Y) + FN \cdot \bar Y}{\tilde f_0 (1-\bar Y) + \tilde f_1 \bar Y} \right).
% % $$
% % Thus yielding the desired result.

% % \section{Doubly robust statistical error decomposition derivation}
% % \label{app:DRderivation}
% % We start by considering the empirical weighted mean estimator under imperfect testing,
% % \begin{align*}
% % &\frac{1}{N} \sum_{j=1}^N \mu(X_j) - \frac{\sum_{j=1}^N I_j W_j \mu(X_j)}{\sum_{j=1}^N I_j W_j} + \frac{\sum_{j=1}^N I_j W_j Y_j^\star}{\sum_{j=1}^N I_j W_j} \\
% % =& \frac{\sum_{i=1}^N  I_j W_j \left( \left[ Y_j (1-P_j) + (1-Y_j) P_j \right] - \mu (X_j) \right)}{\sum_{j=1}^N  I_j W_j } + \frac{1}{N} \sum_{j=1}^N \mu(X_j
% % )
% % \end{align*}
% % Then
% % $$
% % \begin{aligned}
% % &\frac{E_J \left[ I_J W_J \left( \left[ Y_J (1-P_J) + (1-Y_J) P_J \right] - \mu(X_J) \right) \right]}{E_J [ I_J W_J ] } - E_J[Y_J -\mu (X_J)] \\
% % = &\frac{E_J \left[ I_J W_J P_J (1-2Y_J) \right]}{E_J [ I_J W_J ] } + \left( \frac{E_J [I_J W_J (Y_J -\mu (X_J))]}{E_J [ I_J W_J ] } - \frac{E_J[Y_J-\mu(X_J)] E_J[I_J W_J]}{E_J[I_J W_J]} \right) \\
% % \end{aligned}
% % $$
% % The term in parentheses can be re-written as
% % $$
% % \begin{aligned}
% % &\frac{E_J [I_J W_J (Y_J-\mu(X_J))]- E_J[(Y_J-\mu(X_J))] E_J[I_J W_J]}{E_J[I_J W_J]} \\
% % =  &\frac{E_J [I_J W_J (Y_J-\mu(X_J))]- E_J[(Y_J-\mu(X_J))] E_J[I_J W_J]}{\sqrt{V_J(I_J W_J) V_J(Y_J-\mu(X_J))}} \frac{\sqrt{V_J(I_J W_J)}}{E_J[I_J]} \times \sqrt{V_J(Y_J-\mu(X_J))} \\
% % = &\rho_{\tilde I (X),Y-\mu(X)} \times \frac{\sqrt{V_J(I_J W_J)}}{E_J[I_J]} \times \sigma_{Y-\mu(X)}
% % \end{aligned}
% % $$
% % From prior derivations, we then have the desired result.

% % \section{Asymptotic derivations}
% % \label{app:asympderivations}

% % % https://www.thelancet.com/journals/lanres/article/PIIS2213-2600%2820%2930453-7/fulltext
% % % https://www.medrxiv.org/content/10.1101/2020.04.26.20080911v1.full.pdf

% % Here, we take the suggested measures from~\cite{Arevalo2020} which report 87\% sensitivity is a reasonable estimates and~\cite{Cohen2020} which report 97.6\% specificity.  This corresponds to a false negative rate of $13\%$ and false positive rate of $2.4\%$.  To allow for potential uncertainty, we assume two \emph{synthetic} simple random samples.  The first is a simple random sample of true negative cases denoted $S_{FP}$; the second is simple random sample of true positive cases denoted $S_{FN}$.  As $\mu_t$ and $\theta_{t}$ are estimated independently at each time $t$, we can focus on the estimating equations per time point~$t$ separately.

% % Suppose there is a sequence of finite populations of size $N_{\nu}$ indexed by $\nu$.  Each finite population has an associated non-probability at time $t$ of  size~$\tilde n$ and $L_\nu$ probability samples of fixed size $n$ drawn at equally spaced times $\{ t^\prime_l \}_{l=1}^L$ over the study window~$[0,T]$. Let~$\Delta_\nu$ denote the gap times between consecutive sample times. Finally, let~$n_{FP}$ and $n_{FN}$ denote the sample sizes of $S_{FP}$ and $S_{FN}$ respectively.  For simplicity, dependency on $\nu$ is suppressed and the limiting process is represented by $N \to \infty$. The asymptotic argument below relies on regularity conditions C1-C6 from~\cite{Chen2019}, smoothness of the population-level distribution of the time-varying covariates~$\{ X_{j,t} \}$, and smoothness of the propensity function~$\pi(x; \theta)$ as a function of $x \in \mathbb{R}^p$.

% % For each~$\nu$ and $t$, the estimates of $\eta = (\mu_t, \theta_t, FP, FN)$ are given by:
% % $$
% % \Phi (\eta) = \left (
% % \begin{array}{c}
% % \frac{1}{N} \sum_{j=1}^N I_{j,t} \frac{Y_{j,t} - FP - (1-FP-FN) \cdot \mu_t}{\pi_{j,t}} \\
% % \frac{1}{N \sum_{l=1}^L K_{t,t_l^\prime}} \sum_{l=1}^L K_{t,t_l^\prime} \left[ \sum_{j=1}^N I_{j,t^\prime} X_{j,t^\prime_l} - \sum_{j = 1}^N \tilde I_{j,t_l^\prime} \tilde W_{j,t_l^\prime}  \pi_{j,t_l^\prime} X_{j,t_l^\prime} \right] \\
% % \frac{1}{n_{FP}} \sum_{j \in S_{FP}} \frac{Z_j}{FP} - \frac{1-Z_j}{1-FP} \\
% % \frac{1}{n_{FN}} \sum_{j \in S_{FN}} \frac{\tilde Z_j}{FN} - \frac{1-\tilde Z_j}{1-FN} \\
% % \end{array}
% % \right ) = {\bf 0}
% % $$
% % where $\pi_{j,t^\prime} = \pi (X_{j,t^\prime}; \theta_t)$, $Z_j$ is an indicator of a false positive in the sample $S_{FP}$  and $\tilde Z_j$ is an indicator of a false negative in the sample $S_{FN}$.  We assume the $|S_{FP}|^{-1} \sum_{j \in S_{FP}} Z_j = 0.024$ and $|S_{FP}|^{-1} \sum_{j \in S_{FP}} Z_j = 0.30$.  Sample sizes, denoted $n_{FP}$ and $n_{FN}$ respectively, are chosen to achieve desired estimator variance, i.e., $\frac{0.024 \cdot (0.976)}{n_{FP}} = \sigma^2_{FP}$ and $\frac{0.13 \cdot 0.87}{n_{FN}} = \sigma^2_{FN}$. In this paper, we set $\sigma^2_{FP} = 0.02^2$ and $\sigma_{FN}^2 = 0.05^2$ which means $n_{FP} \approx 59$ and $n_{FN} \approx 45$ respectively.

% % To prove consistency, first consider the setting where a single random sample of size $n$ is observed at time $t$.  Let $\tilde \Phi (\eta)$ denote the corresponding estimating equations, which is the same as $\Phi(\eta)$ except for the second component which is now given by
% % $$
% % \frac{1}{N} \left[ \sum_{j=1}^N I_{j,t} X_{j,t} - \frac{1}{N} \sum_{j = 1}^N \tilde I_{j,t} \tilde W_{j,t}  \pi_{j,t} X_{j,t} \right].
% % $$
% % Then under joint randomization of propensity score model and sampling designs, we have $E [ \tilde \Phi (\eta_0) ] = {\bf 0}$.  Consistency then follows by arguments in Section 3.2 of Tsiatis (2006). Under conditions C1-C6 from~\cite{Chen2019} and as $n$ increases, we have $\tilde \Phi (\hat \eta) = 0$ and $\tilde \Phi(\eta_0) = O_p ( n^{-1/2} )$.

% % To prove consistency under the sequence of sampling regimes, we need to show that $E[ \Phi (\eta_0) ] \to 0$ as $N \to \infty$.  First, we study the difference in the 2nd term:
% % \begin{align*}
% % &\frac{1}{N}  E \left[ \sum_{j = 1}^N \tilde I_{j,t} \tilde W_{j,t}  \pi_{j,t} X_{j,t} - \sum_{l=1}^L  \sum_{j = 1}^N \tilde K_{t,t^\prime_l} \tilde I_{j,t^\prime_l} \tilde W_{j,t^\prime_l}  \pi_{j,t^\prime_l} X_{j,t^\prime_l} \right] \\
% % = &\frac{1}{N} \sum_{j = 1}^N \left[ \pi_{j,t} X_{j,t} - \sum_{l=1}^L \tilde K_{t,t^\prime_l} \pi_{j,t^\prime_l} X_{j,t^\prime_l} \right], \\
% % = &\frac{1}{N} \sum_{j = 1}^N \left[ \sum_{l=1}^L \tilde K_{t,t^\prime_l} \pi_{j,t} X_{j,t} \pm \sum_{l=1}^L \tilde K_{t,t^\prime_l} \pi_{j,t_l^\prime} X_{j,t}  - \sum_{l=1}^L \tilde K_{t,t^\prime_l} \pi_{j,t^\prime_l} X_{j,t^\prime_l} \right], \\
% % = &\frac{1}{N} \sum_{j = 1}^N \left[ \left( \sum_{l=1}^L \tilde K_{t,t^\prime_l}  \left( \pi (X_{j,t}; \theta) - \pi (X_{j,t^\prime_l}; \theta) \right) \right) X_{j,t} + \sum_{l=1}^L \tilde K_{t,t^\prime_l} \pi_{j,t^\prime_l} (X_{j,t} - X_{j,t^\prime_l} ) \right],
% % \end{align*}
% % where $\tilde K_{t, t^\prime_l}$ is the normalized kernel such that $\sum_{l=1}^L \tilde K_{t, t^\prime_l} = 1$. The difference in the 1st term is similarly defined.

% % Let~$X_{j,t,l}$ denote the $l$th component of the vector $X_{j,t}$.  Then consider the population-level distribution $\{ X_{j,t,l} \}_{j=1}^N$.  We make the additional assumption that
% % $$
% % \lim_{\epsilon \to 0} \lim_{N \to \infty} \frac{1}{N} \sum_{j=1}^N X_{j,t+\epsilon, l}  = \lim_{N \to \infty} \frac{1}{N} \sum_{j=1}^N X_{j,t, l}.
% % $$
% % That is, the limiting distributions are continuous as a function of $t$. If $\pi(x; \theta)$ is a continuous function in~$x$ and under suitable conditions on the kernel density~$\tilde K_{t, \cdot}$ with the bandwidth $h_\nu \to \infty$ such that $\sum_{l=1}^L \tilde K_{t,t_l} \to \infty$ as $\nu \to \infty$, we have
% % $$
% % \left | \sum_{l=1}^L \tilde K_{t, t_l^\prime} \pi (X_{j, t_l^\prime}; \theta) - \pi (X_{j, t}; \theta) \right|  \to 0
% % $$
% % Combining this, $E \left[ \Phi(\eta_0)  - \tilde \Phi (\eta_0) \right] \to 0$ which implies consistency as desired.

% % Then by a first-order Taylor expansion we have $\hat \eta - \eta_0 = \left[ \phi (\eta) \right]^{-1} \Phi(\eta_0) + o_p (\tilde n^{-1/2})$ where $\phi(\eta) = \partial \Phi (\eta)/\partial \eta$ and $\tilde n = n \sum_{l=1}^L \tilde K_{t,t_l}$.  To calculate
% % $$
% % \frac{\partial \Phi (\eta)}{\partial \mu_t} =
% % \left (
% % \begin{array}{c}
% % -\frac{(1-FP-FN)}{N} \sum_{j=1}^N \frac{I_{j,t}}{\pi_{j,t}} \\
% % 0 \\
% % 0 \\
% % 0
% % \end{array}
% % \right )
% % $$

% % $$
% % \frac{\partial \Phi (\eta)}{\partial \theta_t} =
% % \left (
% % \begin{array}{c}
% % -\frac{1}{N} \sum_{j=1}^N I_{j,t} (Y_{j,t} - FP - (1-FP-FN) \cdot \mu_t) \frac{1-\pi_{j,t}}{\pi_{j,t}} X_{j,t}^\top \\
% % - \frac{1}{N} \sum_{t^\prime = 1}^T K_{t,t^\prime} \sum_{j = 1}^N \tilde I_{j,t^\prime} \tilde W_{j,t^\prime}  \pi_{j,t^\prime} (1-\pi_{j,t^\prime}) X_{j,t^\prime} X_{j,t^\prime}^\top\\
% % 0 \\
% % 0
% % \end{array}
% % \right )
% % $$

% % $$
% % \frac{\partial \Phi (\eta)}{\partial FP} =
% % \left (
% % \begin{array}{c}
% % -\frac{1}{N} \sum_{j=1}^N I_{j,t} \frac{(1-\mu_t)}{\pi_{j,t}} \\
% % 0\\
% % -\frac{1}{n_{FP}} \sum_{j \in S_{FP}} \left( \frac{Z_j}{FP^2} + \frac{1-Z_j}{(1-FP)^2} \right)  \\
% % 0
% % \end{array}
% % \right )
% % $$

% % $$
% % \frac{\partial \Phi (\eta)}{\partial FN} =
% % \left (
% % \begin{array}{c}
% % - \frac{1}{N} \sum_{j=1}^N I_{j,t} \frac{\mu_t}{\pi_{j,t}} \\
% % 0\\
% % 0 \\
% % -\frac{1}{n_{FN}}  \sum_{j \in S_{FN}} \left( \frac{\tilde Z_j}{FN^2} + \frac{1-\tilde Z_j}{(1-FN)^2} \right)
% % \end{array}
% % \right )
% % $$

% % We are interested in computing $- E \left[ \frac{\partial \Phi (\eta)}{\partial \mu_t} \right]^{-1}$ where the expectation is with respect to the random indicators. First, $- E \left[ \frac{\partial \Phi (\eta)}{\partial \mu_t} \right]$ is equal to
% % $$
% % \left (
% % \begin{array}{c c c c}
% % 1-FP-FN & \frac{1}{N} \sum_{j=1}^N \zeta_{t,j} (1-\pi_{j,t}) X_{j,t}^\top & 1 - \mu_t & \mu_t \\
% % 0 & \frac{1}{N} \sum_{t^\prime=1}^T K_{t,t^\prime} \sum_{j=1}^N \pi_{j,t} (1-\pi_{j,t}) X_{j,t} X_{j,t}^\top & 0 & 0\\
% % 0 & 0 & \left( FP (1-FP) \right)^{-1}  & 0\\
% % 0 & 0 & 0 & \left( FN (1-FN) \right)^{-1}
% % \end{array}
% % \right )
% % $$
% % where $\zeta_{t,j} = (Y_{j,t} - FP - (1-FP-FN) \cdot \mu_t)$. This matrix can be written as a diagonal matrix equal to the diagonal of $\partial \Phi_n (\eta)/\partial \eta$ and then a rank-three matrix that is zero except for the first row on the off-diagonal. Let $A + B$ denote the sum broken into these 2 components.  Then Woodbury matrix identity gives us
% % $$
% % (A+B)^{-1} = A^{-1} - A^{-1} B A^{-1}
% % $$
% % In particular, the resulting inverse is of the form
% % $$
% % \left (
% % \begin{array}{c c c c}
% % \frac{1}{1-FP-FN} & b_1 & b_2  & b_3 \\
% % 0 & \left[\frac{1}{N} \sum_{t^\prime=1}^T K_{t,t^\prime} \sum_{j=1}^N \pi_{j,t} (1-\pi_{j,t}) X_{j,t} X_{j,t}^\top\right]^{-1} & 0 & 0\\
% % 0 & 0 & FP (1-FP)   & 0\\
% % 0 & 0 & 0 & FN (1-FN)
% % \end{array}
% % \right)
% % $$
% % where $b_1, b_2,$ and $b_3$ can be calculated using the Woodbury identity.

% % Next, note that $\text{Var}(\Phi(\eta))$ can be decomposed into $A_{1,t} + A_{2,t} + A_{3,t} + A_{4,t}$

% % $$
% % A_{1,t} = \frac{1}{N} \left( \begin{array}{c}
% % \sum_{j=1}^N \frac{I_{j,t} (Y_{j,t} - FP - (1-FP-FN) \mu_t)}{\pi_{j,t}} \\
% % \sum_{t^\prime = 1}^T K_{t,t^\prime} \sum_{j=1}^N I_{j,t^\prime} X_{j,t^\prime} - \pi_{j,t^\prime} X_{j,t^\prime} \\
% % 0 \\
% % 0
% % \end{array}
% % \right)
% % $$
% % and
% % $$
% % A_{2,t} = \frac{1}{N} \left( \begin{array}{c}
% % 0 \\
% % \sum_{t^\prime = 1}^T K_{t,t^\prime} \sum_{j=1}^N \pi_{j,t^\prime} X_{j,t^\prime} - \sum_{j=1}^N \tilde I_{j,t^\prime} \pi_{j,t^\prime} X_{j,t} \\
% % 0 \\
% % 0
% % \end{array}
% % \right),
% % $$
% % and
% % $$
% % A_{3,t} = \left( \begin{array}{c}
% % 0 \\
% % 0 \\
% % \frac{1}{n_{FP}} \sum_{j \in S_{FP}} \frac{Z_j}{FP} - \frac{1-Z_j}{1-FP} \\
% % 0
% % \end{array}
% % \right),
% % $$
% % and
% % $$
% % A_{4,t} = \left( \begin{array}{c}
% % 0 \\
% % 0 \\
% % 0 \\
% % \frac{1}{n_{FN}} \sum_{j \in S_{FN}} \frac{\tilde Z_j}{FN} - \frac{1-\tilde Z_j}{1-FN} \\
% % \end{array}
% % \right),
% % $$
% % By independence of these components, the variance can be calculated by summation over variance of each term individually yielding. First, $V_1$
% % {
% % \tiny
% % $$
% %  N^{-2} \sum_{i=1}^N \left(
% % \begin{array}{c c c c}
% % \frac{(1-\pi_{j,t})}{\pi_{j,t}} (Y_{j,t} - (1-FP-FN) (\mu_{j,t} +FP))^2 & \sum_{t^\prime} K_{t,t^\prime} (1-\pi_{j,t^\prime}) \zeta_{t^\prime, j} X_{j,t^\prime}^\top & 0 & 0 \\
% % \sum_{t^\prime} K_{t,t^\prime} (1-\pi_{j,t^\prime}) \zeta_{t^\prime, j} X_{j,t}&
% % \sum_{t^\prime} K_{t,t^\prime} \pi_{j,t^\prime} (1-\pi_{j,t^\prime}) X_{j,t} X_{j,t}^\top
% % & 0 & 0 \\
% % 0 & 0 & 0 & 0 \\
% % 0 & 0 & 0 & 0 \\
% % \end{array}
% % \right)
% % $$
% % }
% % and $V_{j,t}$ are block-diagonal matrices with only one non-zero block that is equal to $D_{j,t}$ which is the design-based variance-covariance matrix under each probability sampling design.
% % $$
% % V_{3,t} = \left( \begin{array}{c}
% % 0 \\
% % 0 \\
% % \frac{1}{n_{FP}} \frac{1}{FP(1-FP)} \\
% % 0
% % \end{array}
% % \right),
% % $$
% % and
% % $$
% % V_{4,t} = \left( \begin{array}{c}
% % 0 \\
% % 0 \\
% % 0 \\
% % \frac{1}{n_{FN}} \frac{1}{FN(1-FN)}
% % \end{array}
% % \right),
% % $$

% % The asymptotic variance for the IPW estimator is then the first diagonal element of the matrix $E \left[ \phi_n (\eta_0) \right]^{-1} \times \left[ V_{1,t} + V_{2,t} + V_{3,t} + V_{4,t} \right] E \left[ \phi_n (\eta_0) \right]^{-1}$.
% % \section{IRLS}

% % First derivative with respect to $\theta$
% % $$
% % \nabla E(\theta) = \sum_{t^\prime=1}^T K_h(|t^\prime - t|) \left[ \sum_{j=1}^N I_{j,t^\prime} X_{j,t^\prime} + \sum_{j=1}^N \tilde W_{j,t^\prime} \tilde I_{j,t^\prime} \pi_t (X_{j,t^\prime}; \theta) X_{j,t^\prime} \right]
% % $$
% % Second derivative with respect to $\theta$
% % $$
% % H(\theta) = \sum_{t^\prime=1}^T K_h(|t^\prime - t|) \sum_{j=1}^N \tilde W_{j,t^\prime} \tilde I_{j,t^\prime} \pi_t (X_{j,t^\prime}; \theta) \left( 1- \pi_t (X_{j,t^\prime}; \theta) \right) X_{j,t^\prime} X_{j,t^\prime}^\top
% % $$
% % Then the Fisher scoring method sets
% % $$
% % \theta^{(t+1)} = \theta^{(t)} + H \left(\theta^{(t)} \right)^{-1} \nabla E \left(\theta^{(t)} \right)
% % $$

% % \section{Indiana COVID-19 Analysis: Additional details}
% % \label{app:in_add_details}

% % In this section, we describe the two mean-imputation methods.
% % \subsection{Model 1}

% % First, we use the pseudo-likelihood to compute the likelihood of contact given age, gender, and test result.  Figure~\ref{fig:contactlik1} and~\ref{fig:contactlik2} plots the likelihood for 25-34 year olds given a negative test and positive test respectfully.  We see that the likelihood of COVID-19 is much higher for those with a positive test and the likelihood is time-varying.

% % \begin{figure}[!th]
% % \centering
% % \begin{subfigure}{.5\textwidth}
% %  \centering
% %  \includegraphics[width=.9\linewidth]{../figs/tvprop_contact_fig1.png}
% %  \caption{Contact Likelihood Given Negative Test}
% %  \label{fig:contactlik1}
% % \end{subfigure}%
% % \begin{subfigure}{.5\textwidth}
% %  \centering
% % \includegraphics[width=.9\linewidth]{../figs/tvprop_contact_fig2.png}
% %  \caption{Contact Likelihood Given Positive Test}
% %  \label{fig:contactlik2}
% % \end{subfigure}
% % \caption{Likelihood of COVID-19 contact}
% % \label{fig:contactlik}
% % \end{figure}

% % We then use the pseudo-likelihood to compute the likelihood of fever given age, gender, COVID-19 contact status and test result.  Figure~\ref{fig:symptomlik1} and~\ref{fig:symptomlik2} plots the likelihood for 35-44 year olds given a negative test and positive test respectfully.  We see that the likelihood of fever depends heavily on whether they had a COVID-19 contact and their test results. Again, the likelihood is time-varying for most configurations.

% % \begin{figure}[!th]
% % \centering
% % \begin{subfigure}{.5\textwidth}
% %  \centering
% %  \includegraphics[width=.9\linewidth]{../figs/tvprop_symptom_fig1.png}
% %  \caption{Fever Likelihood Given Negative Test \textcolor{red}{FIX!!}}
% %  \label{fig:symptomlik1}
% % \end{subfigure}%
% % \begin{subfigure}{.5\textwidth}
% %  \centering
% % \includegraphics[width=.9\linewidth]{../figs/tvprop_symptom_fig2.png}
% %  \caption{Fever Likelihood Given Positive Test}
% %  \label{fig:symptomlik2}
% % \end{subfigure}
% % \caption{Likelihood of reported fever}
% % \label{fig:symptomlik}
% % \end{figure}

% % Based on mean-imputation of COVID-19 contact and fever indicators, we can use the proposed pseudo-likelihood approach to compute the likelihood of getting tested for COVID-19 given age, ethnicity, race, gender, fever status, and COVID-19 contact indicator.  Figure~\ref{fig:testinglik1} and~\ref{fig:testinglik2} plots the likelihood of testing given fever and COVID-19 contact and no fever nor COVID-19 contact respectfully.  We see that the likelihood is time-varying and depends heavily on both fever and COVID-19 contact indicators.

% % \begin{figure}[!th]
% % \centering
% % \begin{subfigure}{.5\textwidth}
% %  \centering
% %  \includegraphics[width=.9\linewidth]{../figs/tvprop_fig1.png}
% %  \caption{Testing Likelihood Given Fever and COVID-19 Contact}
% %  \label{fig:testinglik1}
% % \end{subfigure}%
% % \begin{subfigure}{.5\textwidth}
% %  \centering
% % \includegraphics[width=.9\linewidth]{../figs/tvprop_fig2.png}
% %  \caption{Testing Propensity}
% %  \label{fig:testinglik2}
% % \end{subfigure}
% % \caption{Testing Likelihood Given no Fever nor COVID-19 Contact}
% % \label{fig:testinglik}
% % \end{figure}

% % \newpage

% % \subsection{Model 2 (Hospitalization)}

% % This model uses hospitalization records to try and improve our mean-imputation strategy.  First, we use the pseudo-likelihood to compute the likelihood of fever given age, gender, and test result.  The model only depends on hospitalization status for positive tests.  Figure~\ref{fig:symptomlik1_model2} and~\ref{fig:symptomlik2_model2} plots the likelihood for 35-44 year olds given a negative test and positive test respectfully.  We see that the likelihood of fever is time-varying.  For a negative test, the likelihood of ever fever is very high in early April, which accounts for testing restrictions. We see that hospitalization significantly increases the risk of fever given a positive test.

% % \begin{figure}[!th]
% % \centering
% % \begin{subfigure}{.5\textwidth}
% %  \centering
% %  \includegraphics[width=.9\linewidth]{../figs/tvprop_symptom_alt_fig1.png}
% %  \caption{Symptom Likelihood Given Negative Test}
% %  \label{fig:symptomlik1_model2}
% % \end{subfigure}%
% % \begin{subfigure}{.5\textwidth}
% %  \centering
% % \includegraphics[width=.9\linewidth]{../figs/tvprop_symptom_alt_fig2.png}
% %  \caption{Symptom Likelihood Given Positive Test}
% %  \label{fig:symptomlik2_model2}
% % \end{subfigure}
% % \caption{Likelihood of COVID-19 contact}
% % \label{fig:symptomlik_model2}
% % \end{figure}

% % First, we computed the probability of fever given COVID-19 positive test and hospitalization as well as COVID-19 positive test and no hospitalization.  We then computed the lieklihood of symptom given a COVID-19 positive test by weighting these two likelihoods by the fraction of COVID-19 positive hospitalizations to COVID-19 total positive cases per week.  Mean imputation of fever status for COVID-19 negative tests was performed based on the above model without use of hospitalization data.

% % Based on mean-imputation of COVID-19 fever using hospitalization records indicators, we can use the proposed pseudo-likelihood approach to compute the likelihood of getting tested for COVID-19 given age, ethnicity, race, gender, and fever status.  Figure~\ref{fig:testinglik1_mainpaper} and~\ref{fig:testinglik1_mainpaper} in Section~\ref{section:tvipw} of the manuscript plots the likelihood of testing given fever and no fever respectfully for all age ranges.  We see that the likelihood is time-varying and depends heavily on both fever status.

% % % \begin{figure}[!th]
% % % \centering
% % % \begin{subfigure}{.5\textwidth}
% % %  \centering
% % %  \includegraphics[width=.9\linewidth]{../figs/tvprop_alt_fig1.png}
% % %  \caption{Testing Likelihood Given Fever}
% % %  \label{fig:tvprop_alt_fig1.png}
% % % \end{subfigure}%
% % % \begin{subfigure}{.5\textwidth}
% % %  \centering
% % % \includegraphics[width=.9\linewidth]{../figs/tvprop_alt_fig2.png}
% % %  \caption{Testing Propensity Given No Fever}
% % %  \label{fig:tvprop_alt_fig2.png}
% % % \end{subfigure}
% % % \caption{COVID-10 Testing Likelihood Given no Fever nor COVID-19 Contact}
% % % \label{fig:testinglik}
% % % \end{figure}

% % \section{Model-based: Prior specification}
% % \label{app:prior_modelbased}

% % For simplicity, we list the priors used in STAN below:
% % \begin{itemize}
% %   \item \code{beta $\sim$ Normal(1.5, 1) T[0,];}
% %   \item \code{gamma $\sim$ normal(0.3, 0.5) T[0,];}
% %   \item \code{sigma $\sim$ normal(0.4, 0.5) T[0,];}
% %   \item \code{phi\_inv $\sim$ exponential(5);}
% %   \item \code{i0 $\sim$ normal(0, 10);}
% %   \item \code{e0 $\sim$ normal(0, 10);}
% %   \item \code{eta $\sim$ normal(1, 1) T[0,];}
% %   \item \code{eta\_two $\sim$ normal(1, 1) T[0,];}
% %   \item \code{eta\_three $\sim$ normal(1, 1) T[0,];}
% %   \item \code{nu $\sim$ exponential(1./5);}
% %   \item \code{nu\_two $\sim$ exponential(1./5);}
% %   \item \code{nu\_three $\sim$ exponential(1./5);}
% %   \item \code{xi\_raw $\sim$ beta(1, 1);}
% %   \item \code{phi = 1/phi\_inv;}
% %   \item \code{xi = xi\_raw + 0.5;}
% % \end{itemize}
% % where \code{beta, eta, eta\_two, eta\_three} refer to the four values that form the time-varying parameter $\beta_t$ when combined with \code{xi, eta, eta\_two, eta\_three} in the SEIR model as described in Section~\ref{section:modelbased}.

\end{document}