### 2022-03-30 write-up/letters/cover letter-covid.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Plain Cover Letter
% LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute (http://www.rpi.edu/dept/arc/training/latex/resumes/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt]{letter} % Default font size of the document, change to 10pt to fit more text

\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font
%\usepackage{helvet} % Uncomment this (while commenting the above line) to use the Helvetica font

% Margins
\topmargin=-0.5in % Moves the top of the document 1 inch above the default
\textheight=8.75in % Total height of the text on the page before text goes on to the next page, this can be increased in a longer letter
\oddsidemargin=0in % Position of the left margin, can be negative or positive if you want more or less room
\textwidth=6.5in % Total width of the text, increase this if the left margin was decreased and vice-versa

\let\raggedleft\raggedright % Pushes the date (at the top) to the left, comment this line to have the date on the right

%\address{Rutgers University and University of Michigan}

\begin{document}

%----------------------------------------------------------------------------------------
%	ADDRESSEE SECTION
%----------------------------------------------------------------------------------------

\begin{letter}{Professor Karen Kafadar \\
    Editor-In-Chief, {\em Annals of Applied Statistics}
         }
%----------------------------------------------------------------------------------------
%	YOUR NAME & ADDRESS SECTION
%----------------------------------------------------------------------------------------

%\begin{center}
%\large
%\end{center}
%\vfill

\signature{Walter Dempsey\\
University of Michigan \\
Department of Biostatistics\\
M4057 SPH II \\
1415 Washington Heights \\
wdem@umich.edu} % Your name for the signature at the bottom

%----------------------------------------------------------------------------------------
%	LETTER CONTENT SECTION
%----------------------------------------------------------------------------------------

\date\today

\opening{Dear Dr. Kafadar:}

Kindly find enclosed the revised version of the manuscript number AOAS2006-048 titled ``Statistical paradoxes in coronavirus case-counts: selection bias, measurement error, and the COVID-19 pandemic'' now titled ``Addressing selection bias and measurement error in COVID-19 case count data using auxiliary information.''  We have revised the article to address the referees’ requests about exposition and other points of clarification.  The reports called for a major revision to make the manuscript ``forward compatible'', to re-structure the manuscript to present figures and tables near their corresponding discussion, to discuss potential non-response bias in the motivating studies, to include a more careful discussion of COVID-19 testing and disease dynamics, to include a more precise definition of the active infection rate and how it impacts model-based estimation, and to more carefully tie the proposal to related parts of the epidemiological literature. The new version complies with these requests.

We believe the work presents an important advance in how researchers should think about addressing time-varying selection bias and measurement error to better understanding the COVID-19 pandemic trajectory. Attached please find a detailed response to the referees outlining how the changes to the manuscript address the concerns raised in their reports.

Thank you for your consideration. We look forward to hearing back from you.

\closing{Sincerely,}

\end{letter}

\end{document}