### 2022-09-12 write-up/letters/response-ref-1-v1.tex
updating letters and sensitivity analysis

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Plain Cover Letter
% LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute (http://www.rpi.edu/dept/arc/training/latex/resumes/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt]{letter} % Default font size of the document, change to 10pt to fit more text

\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font
%\usepackage{helvet} % Uncomment this (while commenting the above line) to use the Helvetica font

% Margins
\topmargin=-1in % Moves the top of the document 1 inch above the default
\textheight=8.5in % Total height of the text on the page before text goes on to the next page, this can be increased in a longer letter
\oddsidemargin=-10pt % Position of the left margin, can be negative or positive if you want more or less room
\textwidth=6.5in % Total width of the text, increase this if the left margin was decreased and vice-versa

\let\raggedleft\raggedright % Pushes the date (at the top) to the left, comment this line to have the date on the right
\usepackage{amsthm, amsfonts}
\usepackage{xcolor}
\usepackage{hyperref}

\def\Y{{\bf Y}}

\begin{document}

%----------------------------------------------------------------------------------------
%	ADDRESSEE SECTION
%----------------------------------------------------------------------------------------

\begin{letter}{Professor
	Peter Glynn\\
	Editor, {\em Journal of Applied Probability}}

%----------------------------------------------------------------------------------------
%	YOUR NAME & ADDRESS SECTION
%----------------------------------------------------------------------------------------

%\begin{center}
%\large
%\end{center}
%\vfill

\signature{Walter Dempsey\\
University of Michigan\\
Department of Biostatistics\\
1415 Washington Heights\\
Ann Arbor, MI 48103} % Your name for the signature at the bottom

%----------------------------------------------------------------------------------------
%	LETTER CONTENT SECTION
%----------------------------------------------------------------------------------------

\vspace{5mm}

\newpage

{\bf Response to Referee}

We appreciate the very helpful comments from the referee which has helped improve the paper substantially. We have addressed the concerns as best as possible in this revision. We provide a further point-by-point explanation below. Any comments in {\it italics} indicate a statement from the referee report.
\begin{enumerate}
\item {\it I appreciate the addition of the remark that states that the dynamics mentioned in this paper refer to the original strain. There are still lines (for example: “Upon infection with SARS-CoV-2, an incubation period (time to symptom onset) starts and lasts approximately five days”) that are more definitive than I would prefer. Perhaps adding in “Upon infection with the original SARS-CoV-2 variant” would be preferable.}

\vspace{5mm}
We agree with this point and have gone through the manuscript to reduce the chances a reader miscontrues our explanations. We aim for clarity in these edits by avoiding definitive statements while not making statements that are too vauge or ambiguous.
\vspace{5mm}

\item {\it The reliance on the random sample with such a low response rate may be subject to unmeasured confounding - perhaps a sensitivity analysis showing the impact of a potential unmeasured confounder could strengthen this point.}

\vspace{5mm}
We have added a sensitivity analysis to the supplementary materials where we consider an unmeasured binary confounder~$Z_{j,t} \in \{0,1\}$.  In this setting, {\it prevalence}~$pr(Z=1|X,Y)$ and {\it propensity} $\pi(X, Z; (\theta,\theta_z))$ both impact the sensitivity analysis.  Imputation-based sensitivity analyses will thus depend heavily on the choice of prevalence and $\theta_z$.    Therefore, we consider a Taylor series expansion of the IPW and DR-estimators at the estimated propesity which is equivalent to~$\theta_z = 0$.denote the coefficient in the selection propensity model and
$$
\hat \mu_{IPW} + \theta_z \frac{1}{\sum_{j=1}^N I^{NR}_j w_j (X_{j,t}; \hat \theta_t)} \left[ \sum_{j=1}^N I^{NR}_j \frac{1-\pi_j(X_{j,t}; \hat \theta_t)}{\pi_j(X_{j,t}; \hat \theta_t)} E[Z_j] (Y_j - \hat \mu_{IPW}) \right]
$$
Thus bias depends on the correlation of the error $Y_j - \hat \mu_{IPW}$ to the odds ratio times the expected value of the unmeasured binary confounder.  We use this to calculate the

\vspace{5mm}

\item {\it On the plots that show the active infection rate estimates, perhaps including a line that is the case counts divided by the population as a floor would be worthwhile (at least in the supplement) to demonstrate that the model-based estimates are at least above this.}

\vspace{5mm}
We have added this this floor .  We clarify that ``case counts'' in the comment above should be a measure of recent case counts to be a relevant  active infection rate
\vspace{5mm}

\item {\it I found the grey hard to differentiate for the figures that have more than two categories – perhaps varying the shapes would help if trying to keep these grey-scale?}

\vspace{5mm}
We have added shapes to help differentiate figures while keeping the grey-scale.
\vspace{5mm}

\end{enumerate}

\end{letter}

\end{document}