### 2022-09-15 write-up/letters/response-ref-1-v1.tex
updating sensitivity model and making all relevant changes for final referee comments

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Plain Cover Letter
% LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute (http://www.rpi.edu/dept/arc/training/latex/resumes/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt]{letter} % Default font size of the document, change to 10pt to fit more text

\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font
%\usepackage{helvet} % Uncomment this (while commenting the above line) to use the Helvetica font

% Margins
\topmargin=-1in % Moves the top of the document 1 inch above the default
\textheight=8.5in % Total height of the text on the page before text goes on to the next page, this can be increased in a longer letter
\oddsidemargin=-10pt % Position of the left margin, can be negative or positive if you want more or less room
\textwidth=6.5in % Total width of the text, increase this if the left margin was decreased and vice-versa

\let\raggedleft\raggedright % Pushes the date (at the top) to the left, comment this line to have the date on the right
\usepackage{amsthm, amsfonts}
\usepackage{xcolor}
\usepackage{hyperref}

\def\Y{{\bf Y}}

\begin{document}

%----------------------------------------------------------------------------------------
%	ADDRESSEE SECTION
%----------------------------------------------------------------------------------------

\begin{letter}{Professor
	Peter Glynn\\
	Editor, {\em Journal of Applied Probability}}

%----------------------------------------------------------------------------------------
%	YOUR NAME & ADDRESS SECTION
%----------------------------------------------------------------------------------------

%\begin{center}
%\large
%\end{center}
%\vfill

\signature{Walter Dempsey\\
University of Michigan\\
Department of Biostatistics\\
1415 Washington Heights\\
Ann Arbor, MI 48103} % Your name for the signature at the bottom

%----------------------------------------------------------------------------------------
%	LETTER CONTENT SECTION
%----------------------------------------------------------------------------------------

\vspace{5mm}

\newpage

{\bf Response to Referee}

We appreciate the comments from the referee which has helped improve the paper. We have addressed the concerns as best as possible in this revision. We provide a further point-by-point explanation below. Any comments in {\it italics} indicate a statement from the referee report.

\begin{enumerate}
\item {\it I appreciate the addition of the remark that states that the dynamics mentioned in this paper refer to the original strain. There are still lines (for example: “Upon infection with SARS-CoV-2, an incubation period (time to symptom onset) starts and lasts approximately five days”) that are more definitive than I would prefer. Perhaps adding in “Upon infection with the original SARS-CoV-2 variant” would be preferable.}

\vspace{5mm}
We agree with this point and have gone through the manuscript to reduce the chances a reader miscontrues our explanations. We aim for clarity in these edits by avoiding definitive statements while not making statements that are too vauge or ambiguous.
\vspace{5mm}

\item {\it The reliance on the random sample with such a low response rate may be subject to unmeasured confounding - perhaps a sensitivity analysis showing the impact of a potential unmeasured confounder could strengthen this point.}

\vspace{5mm}
We have added a sensitivity analysis in Section K of the supplementary materials. Note the difficulty of such an analysis comes from needing to specify a joint distribution of the measured covariates~$X_t$, the potential unmeasured and \emph{time-varying} confounder~$U_t$, the self-selection indicator~$I_t$ and the outcome~$Y_t$.  Here, we extend recent work in the causal inference literature on sensitivity analysis to the self-selection setting in order to avoid direct specification of these distributions.  Instead, we specify a conditional distribution of the self-selection propensity~$\pi_t(X_t, U_t)$ given the estimated propensity~$\pi_t (X_t)$ and use this as the basis for a sensitivity analysis to derive a simple expression for potential bias.  The bias depends on two sensitivity parameters: (1) $\alpha_t$ which measures the confounder-selection relationship strength, and (2) $\delta_t$ which measures the confounder-outcome relationship strength.  We calibrate the selected sensitivity parameter values by analyzing the observed data as if a particular observed covariate were an unobserved confounder.  A complete set of derivations and associated discussion are presented in Section K of the supplementary materials, which we briefly discuss in Section 5.3 of the main manuscript.
\vspace{5mm}

\item {\it On the plots that show the active infection rate estimates, perhaps including a line that is the case counts divided by the population as a floor would be worthwhile (at least in the supplement) to demonstrate that the model-based estimates are at least above this.}

\vspace{5mm}
We have added the floor of weekly case counts divided by population size to the figure of active infection rate estimates.  We interpret ``case counts'' in the referee comment above as ``recent case counts'' so that the floor calculation is relevant for estimating the active infection rate.
\vspace{5mm}

\item {\it I found the grey hard to differentiate for the figures that have more than two categories – perhaps varying the shapes would help if trying to keep these grey-scale?}

\vspace{5mm}
We have added shapes in each figure with more than two categories to help differentiate each curve while keeping them grey-scale to match AOAS figure guidelines.
\vspace{5mm}

\end{enumerate}

\end{letter}

\end{document}