### 2022-09-20 write-up/letters/cover letter-covid.tex
Updated papers - almost there

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Plain Cover Letter
% LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute (http://www.rpi.edu/dept/arc/training/latex/resumes/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt]{letter} % Default font size of the document, change to 10pt to fit more text

\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font
%\usepackage{helvet} % Uncomment this (while commenting the above line) to use the Helvetica font

% Margins
\topmargin=-0.5in % Moves the top of the document 1 inch above the default
\textheight=8.75in % Total height of the text on the page before text goes on to the next page, this can be increased in a longer letter
\oddsidemargin=0in % Position of the left margin, can be negative or positive if you want more or less room
\textwidth=6.5in % Total width of the text, increase this if the left margin was decreased and vice-versa

\let\raggedleft\raggedright % Pushes the date (at the top) to the left, comment this line to have the date on the right

%\address{Rutgers University and University of Michigan}

\begin{document}

%----------------------------------------------------------------------------------------
%	ADDRESSEE SECTION
%----------------------------------------------------------------------------------------

\begin{letter}{Professor Ji Zhu \\
    Editor-In-Chief, {\em Annals of Applied Statistics}
         }
%----------------------------------------------------------------------------------------
%	YOUR NAME & ADDRESS SECTION
%----------------------------------------------------------------------------------------

%\begin{center}
%\large
%\end{center}
%\vfill

\signature{Walter Dempsey\\
University of Michigan \\
Department of Biostatistics\\
M4057 SPH II \\
1415 Washington Heights \\
wdem@umich.edu} % Your name for the signature at the bottom

%----------------------------------------------------------------------------------------
%	LETTER CONTENT SECTION
%----------------------------------------------------------------------------------------

\date\today

\opening{Dear Dr. Zhu:}

Kindly find enclosed the revised version of the manuscript number AOAS2006-048 titled ``Addressing selection bias and measurement error in COVID-19 case count data using auxiliary information.''  We are happy that the two reviewers, associate editor, and editor have all found the revision much improved. We have revised the article to address the few remaining concerns that one of the reviewers raised but are glad that they ``appreciate the paper and [are] very supportive of the pragmatic approach.''  We have edited the remark, performed a sensitivy analysis to assess sensitivity due to potential confounding, and adjusted the figures to assist in comprehension.

We continue to believe the work presents an important advance in how researchers should think about addressing time-varying selection bias and measurement error to better understanding the COVID-19 pandemic trajectory. Attached please find a detailed response to the referee outlining how the changes to the manuscript address the concerns raised in their report.

Thank you for your consideration. We look forward to hearing back from you.

\closing{Sincerely,}

\end{letter}

\end{document}